### 2020-07-14 text/data_set_generation.Rmd
made child docs, fixed mutation burden plot, worked on fig4 function - still needs some tweaking

<div style="text-align: right"> Generation of Somatic Mutation Data </div>
```{r data-set-creation, eval=FALSE}
## Data acquired from GDC
tcga_ssm <- fread("../data/simple_somatic_mutation.open.tsv.gz", header = T, sep = "\t")
## Somatic mutations focused in analysis
selected_mut <- c("3_prime_UTR_variant", "5_prime_UTR_premature_start_codon_gain_variant", "5_prime_UTR_variant", "frameshift_variant", "missense_variant", "splice_acceptor_variant", "splice_donor_variant", "stop_lost", "start_lost", "stop_gained")
## Filtering tcga_ssm for our selected mutations
tcga_ssm <- tcga_ssm[tcga_ssm$consequence_type %in% selected_mut, ]
## Selecting column that we are interested in
tcga_ssm <- data.frame(chr = paste0("chr",tcga_ssm$chromosome),
                   start = tcga_ssm$chromosome_start,
                   end = tcga_ssm$chromosome_end,
                   ref_allele = tcga_ssm$reference_genome_allele,
                   alt_allele = tcga_ssm$mutated_to_allele,
                   type = tcga_ssm$consequence_type,
                   mut_type = tcga_ssm$mutation_type,
                   ensembl_id = tcga_ssm$gene_affected,
                   donor_id = tcga_ssm$icgc_donor_id,
                   sample_id = tcga_ssm$submitted_sample_id,
                   stringsAsFactors = F)
## Removing duplicate rows
tcga_ssm <- tcga_ssm %>% distinct()
## Recategorizing the mutation type
tcga_ssm[which(tcga_ssm$type == "frameshift_variant"), "type"] <- tcga_ssm[which(tcga_ssm$type == "frameshift_variant"), "mut_type"] 
tcga_ssm <- tcga_ssm %>% mutate(type = case_when(type == "insertion of <=200bp" ~ "frameshins_variant", ## relabel values
                                                 type == "deletion of <=200bp" ~ "frameshdel_variant",
                                                 type == "splice_acceptor_variant" ~ "splice_variant",
                                                 type == "splice_donor_variant" ~ "splice_variant",
                                                 type == "start_lost" ~ "nonsense_variant",
                                                 type == "stop_gained" ~ "nonsense_variant",
                                                 type == "5_prime_UTR_premature_start_codon_gain_variant" ~ "UTR_5_prime_variant",
                                                 type == "5_prime_UTR_variant" ~ "UTR_5_prime_variant",
                                                 type == "3_prime_UTR_variant" ~ "UTR_3_prime_variant",
                                                 type == "missense_variant" ~ "missense_variant",
                                                 type == "stop_lost" ~ "nonstop_variant")) %>% 
  dplyr::select(-mut_type) ## rm mut_type col

## Preparing sample metadata ##

complete_sample <- tcga_ssm %>% .[,"sample_id"] %>% unique() %>% .[grep(pattern= "TCGA-", x = .)] # complete TCGA tumor samples

tcga_sample <- fread("../data/sample.tsv.gz", header = T, sep = "\t")
tcga_sample <- tcga_sample %>% dplyr::select(project_code, submitted_sample_id, submitted_specimen_id, icgc_donor_id, submitted_donor_id, study) %>% 
  .[.$submitted_sample_id %in% complete_sample,] %>%
  mutate(type = substring(.$submitted_sample_id, first = 14, last = 15)) %>% ## 01 is primary solid tumor
  filter(type == "01") ## filtering 01 samples
tcga_dup_id <- tcga_sample %>% .[.$icgc_donor_id %in% .[which(duplicated(.$submitted_donor_id)),]$icgc_donor_id,] %>% ## what are the dup donors
  filter(study == "PCAWG") ## selecting dup samples that are PCAWG
tcga_sample <- tcga_sample %>% .[!.$icgc_donor_id %in% tcga_dup_id$icgc_donor_id, ] %>% ## remove the dups from tcga_samples
  rbind(., tcga_dup_id) ## rbind the fixed dups

tcga_donor <- fread("../data/donor.tsv.gz", header = T, sep = "\t")
tcga_donor <- tcga_donor %>% dplyr::select(icgc_donor_id, project_code, submitted_donor_id, donor_sex, donor_vital_status, disease_status_last_followup, donor_age_at_diagnosis, donor_age_at_enrollment, donor_age_at_last_followup, donor_survival_time, donor_interval_of_last_followup) ## selecting columns that we are interested in

sample_sheet <- left_join(tcga_sample, tcga_donor, by = c("submitted_donor_id", "project_code", "icgc_donor_id"))

## Preparing gene metadata ##

# genelist <- unique(as.character(tcga_ssm$ensembl_id)) ## complete genelist
# ensembl_2_entrez_full_list <- bitr(genelist, 
#                                    fromType = "ENSEMBL", 
#                                    toType = c("ENTREZID"), OrgDb = org.Hs.eg.db)
load("../data/ensembl_2_entrez_full_list.Rda")
tcga_ssm <- tcga_ssm[tcga_ssm$ensembl_id %in% ensembl_2_entrez_full_list$ENSEMBL,]

## Now we only have 19067 genes
complete_ensembl <- as.data.frame(unique(as.character(tcga_ssm$ensembl_id)))
names(complete_ensembl) <- "ensembl_id"

selected_mut <- unique(as.character(tcga_ssm$type))
## For every mutation type.. select them.. and create a tsv file 
for(selected_mut in selected_mut) {
  print(selected_mut)
  df <-  tcga_ssm %>% filter(type == selected_mut) %>% ## filter data by selected_mut
    dplyr::select(sample_id, ensembl_id) %>% ## select column sample_id and ensembl_id
    mutate(mutation = 1) %>% ## create a new column with 1's
    distinct() %>% ## remove duplicated rows
    spread(., key = "sample_id", value = "mutation") %>% ## long to wide format
    remove_rownames() %>% ## move ensembl_id column to rownames
    column_to_rownames(var = "ensembl_id") %>% 
    replace(is.na(.), values = 0) %>% ## convert NA's to 0's
    .[, names(.) %in% sample_sheet$submitted_sample_id] ## rm any sample_id not included in complete_sample
  missing_gene <- as.data.frame(matrix(data = 0, nrow = length(complete_ensembl$ensembl_id[!complete_ensembl$ensembl_id %in% rownames(df)]), ncol = length(df)))
  rownames(missing_gene) <- complete_ensembl$ensembl_id[!complete_ensembl$ensembl_id %in% rownames(df)] ## insert missing ensembl_id into row names
  names(missing_gene) <- names(df) ## insert sample_id names into col names
  df <- rbind(df, missing_gene) ## rbind df and missing_gene
  rm(missing_gene) ## rm data
  gc()
  df <- df[match(complete_ensembl$ensembl_id, rownames(df)),] ## match df rownames with genelist
  missing_sample <- as.data.frame(matrix(data = 0, nrow = nrow(df), ncol = length(sample_sheet$submitted_sample_id[!sample_sheet$submitted_sample_id %in% names(df)])))
  rownames(missing_sample) <- rownames(df) ## insert ensembl_id into row names
  names(missing_sample) <- sample_sheet$submitted_sample_id[!sample_sheet$submitted_sample_id %in% names(df)] ## insert missing sample_id names into col names
  df <- cbind(df, missing_sample) ## rbind df and missing_gene
  rm(missing_sample) ## rm data
  gc()
  df <- df[, match(sample_sheet$submitted_sample_id, names(df))]
  assign(paste0(selected_mut), value = df) ## create df in r environment
  rm(df)
  gc()
  }

# ensembl <- useEnsembl(biomart="ensembl", dataset = "hsapiens_gene_ensembl")
 # genelist <- rownames(missense_variant)
 # gene_data <- getBM(
 #   attributes = c(
 #     'ensembl_gene_id',
 #     'external_gene_name',
 #     'gene_biotype',
 #     'chromosome_name', 
 #     'start_position',
 #     'end_position', 
 #     'strand'),
 #   filters = 'ensembl_gene_id',
 #   values = genelist,
 #   mart = ensembl)
 # save(gene_data, file = "gene_data.Rda")
load("../data/gene_data.Rda")
######################

missense_variant <- missense_variant[rownames(missense_variant) %in% gene_data$ensembl_gene_id,]
UTR_3_prime_variant <- UTR_3_prime_variant[rownames(UTR_3_prime_variant) %in% gene_data$ensembl_gene_id,]
UTR_5_prime_variant <- UTR_5_prime_variant[rownames(UTR_5_prime_variant) %in% gene_data$ensembl_gene_id,]
frameshdel_variant <- frameshdel_variant[rownames(frameshdel_variant) %in% gene_data$ensembl_gene_id,]
frameshins_variant <- frameshins_variant[rownames(frameshins_variant) %in% gene_data$ensembl_gene_id,]
splice_variant <- splice_variant[rownames(splice_variant) %in% gene_data$ensembl_gene_id,]
nonstop_variant <- nonstop_variant[rownames(nonstop_variant) %in% gene_data$ensembl_gene_id,]
nonsense_variant <- nonsense_variant[rownames(nonsense_variant) %in% gene_data$ensembl_gene_id,]

## Matching df to the ensembl_id orders in gene_data
match_ensembl <- function(x) {
  x <- x[match(gene_data$ensembl_gene_id, rownames(x)),]
}

missense_variant <- match_ensembl(missense_variant)
UTR_3_prime_variant <- match_ensembl(UTR_3_prime_variant)
UTR_5_prime_variant <- match_ensembl(UTR_5_prime_variant)
frameshdel_variant <- match_ensembl(frameshdel_variant)
frameshins_variant <- match_ensembl(frameshins_variant)
splice_variant <- match_ensembl(splice_variant)
nonstop_variant <- match_ensembl(nonstop_variant)
nonsense_variant <- match_ensembl(nonsense_variant)

genes <- GRanges(
  seqnames = paste("chr", gene_data$chromosome_name, sep=""),
  ranges = IRanges(
    start = gene_data$start_position,
    end = gene_data$end_position),
  strand = gene_data$strand,
  mcols = list(
    ensembl_gene_id = gene_data$ensembl_gene_name,
    hgnc_symbol = gene_data$hgnc_symbol))

gdc_tumor_coding_complete <- SummarizedExperiment(
  assays = list(
    missense   = missense_variant,
    splice     = splice_variant,
    frameshdel = frameshdel_variant,
    frameshins = frameshins_variant,
    utr3       = UTR_3_prime_variant,
    utr5       = UTR_5_prime_variant,
    nonsense   = nonsense_variant,
    nonstop    = nonstop_variant),
  rowRanges = genes,
  colData = sample_sheet)
save(gdc_tumor_coding_complete, file = "../data/gdc_tumor_coding_complete.Rdata")
rm(list = ls(all.names = T))
gc()
```
<div style="text-align: right"> Reactome Pathway Creation </div>
```{r reactome-pathway-creation, eval=FALSE}
load("../data/gdc_tumor_coding_complete.Rdata")
load("../data/gene_data.Rda")
## Preparing level_0 to level_1 pathway conversion table. level_0 is the master pathway. level_1 is the parent pathway
main2first <- fread(file = "../data/reactome_main2first_level.txt", header = T, sep = "\t")

main2first <- main2first %>% .[,2:3] %>%
  'colnames<-' (c("first_level", "main_level")) ## rename colnames

## Filtering out master pathways from first2second_level
first2second <- fread(file = "../data/reactome_first2second_level.txt", header = F, sep = "\t")
first2second <- first2second %>% 'colnames<-' (c("first_level", "second_level")) %>% ## rename colnames
  .[grep(pattern = "-HSA-", .$first_level)] %>% ## keep human pathways
  .[!.$first_level %in% unique(main2first$main_level),] %>% ## remove main level pathways from first level
  .[!.$first_level %in% unique(.$second_level),] ## remove second level pathways from first level

ensembl2reactome <- fread(file = "../data/Ensembl2Reactome_All_Levels.txt", header = F, sep = "\t") ## ensembl to pathway list acquired from reactome database
ensembl2reactome <- ensembl2reactome %>% .[,c(1:2,4)] %>% ## select columns ensembl, pathway_id, pathway_name columns
  'colnames<-' (c("ENSEMBL", "PATH_ID", "PATH_NAME")) %>% ## rename colnames
  .[grep(pattern = "-HSA-", .$PATH_ID),] %>% ## keep human pathways
  .[.$PATH_ID %in% first2second$second_level,] ## include only second level pathways

## Removing pathways that are not molecular functions such as pathways that related to disease, defect, bacteria, disorder, myosin and etc
rm_pathways <- c("R-HSA-1222499", "R-HSA-1226099", "R-HSA-1300642", "R-HSA-1300645", "R-HSA-162906",  "R-HSA-1643713", "R-HSA-168254",
                 "R-HSA-212436", "R-HSA-2160456", "R-HSA-2219528", "R-HSA-2474795", "R-HSA-2534343", "R-HSA-2644603", "R-HSA-2978092",
                 "R-HSA-3296482", "R-HSA-3304351", "R-HSA-3560782", "R-HSA-3781860", "R-HSA-388396",  "R-HSA-3906995", "R-HSA-4791275",
                 "R-HSA-5339562", "R-HSA-5387390", "R-HSA-5545483", "R-HSA-5576886", "R-HSA-5576890", "R-HSA-5576892", "R-HSA-5576893",
                 "R-HSA-5576894", "R-HSA-5578768", "R-HSA-5578775", "R-HSA-5579029", "R-HSA-5602358", "R-HSA-5609975", "R-HSA-5619084",
                 "R-HSA-5619102", "R-HSA-5632927", "R-HSA-5632928", "R-HSA-5632968", "R-HSA-5632987", "R-HSA-5663084", "R-HSA-5687613",
                 "R-HSA-6802957", "R-HSA-8862803", "R-HSA-8876384", "R-HSA-9005891", "R-HSA-9605310", "R-HSA-9616333", "R-HSA-9616334",
                 "R-HSA-9629232", "R-HSA-9630750", "R-HSA-9632693", "R-HSA-9645722", "R-HSA-9646303", "R-HSA-9646304")

## Ensembl to reactome pathway lookup table
gene2level2_lookup <- ensembl2reactome %>% dplyr::select(-PATH_NAME) %>% ## remove PATH_NAME col
  .[.$ENSEMBL %in% gene_data$ensembl_gene_id,] %>% ## only include genes that are in the gene_data 
  mutate(member = 1) %>% ## add a column for ensembl to pathway membership
  distinct() %>% ## keep unique rows
  spread(data = ., key = "PATH_ID", value = "member") %>% ## long to wide format
  column_to_rownames(var = "ENSEMBL") %>% ## ensembl_id as rownames
  .[,!names(.) %in% rm_pathways] %>% ## filter rm_pathways from lookup table
  replace(is.na(.), values = 0) ## replace NA's with 0's
#save(gene2level2_lookup, file = "gene2_level2_lookup.Rda")
## Pathway names to level 2 lookup table
pathname2level2 <- ensembl2reactome %>% dplyr::select(-ENSEMBL) %>% ## remove ENSEMBL col
  .[.$PATH_ID %in% names(gene2level2_lookup),] %>% ## only include pathways in gene2level2_lookup
  distinct(.) ## keep unique rows
#save(pathname2level2, file = "pathname2level2.Rda")
rm(rm_pathways, ensembl2reactome, main2first, first2second)
gc()
```
<div style="text-align: right"> Filter tissue-specific low expressed genes </div>
```{r filter-tissue-specific-low-expressed-genes, eval=FALSE}
gtf_file <- "../data/gencode.v30.annotation.gtf"
gencode_gtf <- rtracklayer::import(gtf_file)
gene_info <- as.data.frame(mcols(gencode_gtf)[, c("type", "gene_id", "transcript_id", "gene_type", "transcript_type", "gene_name")])#; rm(gencode_gtf)
## We only want transcript only 
gene_info <- gene_info[gene_info$type == "transcript",]
## We only want protein_coding only
gene_info <- gene_info[gene_info$gene_type == "protein_coding",]
gene_info <- gene_info[gene_info$transcript_type == "protein_coding",]
## Filtering for transcript_id, gene_id, gene_name columns
g2t <- gene_info %>% dplyr::filter(type == "transcript") %>% dplyr::select(transcript_id, gene_id, gene_name)
## Condense df to only having unique data. 
good.g2t <- as_tibble(unique(g2t))

## df with transcript_id and entrez id                      
gencode_entrez <- readr::read_tsv("../data/gencode.v30.metadata.EntrezGene.gz", col_names = FALSE)
head(gencode_entrez[1:10,])
colnames(gencode_entrez) <- c("transcript_id", "entrez_id")
## Joining data that are joined together by transcript_id
entrez_g2t <- left_join(good.g2t, gencode_entrez)
head(entrez_g2t[1:10,])
## Remove transcript_id column
entrez_g2t$transcript_id <- NULL
## Remove . and the numbers following from the ensembl id.
entrez_g2t$gene_id <- gsub("\\..*", "", entrez_g2t$gene_id)
## Removing data that does not have entrez id. 83091 out of 83647 had NA values
table(is.na(entrez_g2t$entrez_id))
entrez_g2t <- entrez_g2t[!is.na(entrez_g2t$entrez_id),]
entrez_g2t <- unique(entrez_g2t[, c("gene_id", "gene_name", "entrez_id")])
## Are there any duplicates in our df? Yes. Only one dupped ensembl id.
table(duplicated(entrez_g2t$gene_id))
## These are pseudogenes... We will remove them all.
dups_gene_id <- entrez_g2t[which(duplicated(entrez_g2t$gene_id)),]
## We're removing the rows that are associated with this pseudogene
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285607")),]
## It's gone
which(entrez_g2t$gene_id == "ENSG00000285607")
## Are there any duplicate gene names? Yes! 14 of them
table(duplicated(entrez_g2t$gene_name))
dups_gene_name <- entrez_g2t[which(duplicated(entrez_g2t$gene_name)),]
entrez_g2t[entrez_g2t$gene_name == "TBCE",]
## ENSG00000285053  is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285053")),]
entrez_g2t[entrez_g2t$gene_name == "PDE11A",]
## ENSG00000284741 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000284741")),]
entrez_g2t[entrez_g2t$gene_name == "ATXN7",]
## ENSG00000285258 is not in the database anywhere
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285258")),]
entrez_g2t[entrez_g2t$gene_name == "MATR3",]
## ENSG00000280987 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000280987")),]
entrez_g2t[entrez_g2t$gene_name == "SOD2",]
## ENSG00000285441 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285441")),]
entrez_g2t[entrez_g2t$gene_name == "ABCF2",]
## ENSG00000285292 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285292")),]
entrez_g2t[entrez_g2t$gene_name == "PINX1",]
## ENSG00000258724 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000258724")),]
entrez_g2t[entrez_g2t$gene_name == "HSPA14",]
## ENSG00000284024 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000284024")),]
entrez_g2t[entrez_g2t$gene_name == "IGF2",]
## ENSG00000284779 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000284779")),]
entrez_g2t[entrez_g2t$gene_name == "ATF7",]
## ENSG00000267281 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000267281")),]
entrez_g2t[entrez_g2t$gene_name == "DIABLO",]
## ENSG00000284934 is a homolog
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000284934")),]
entrez_g2t[entrez_g2t$gene_name == "GGT1",]
## ENSG00000286070 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000286070")),]
entrez_g2t[entrez_g2t$gene_name == "SCO2",]
## ENSG00000130489 no longer exist
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000130489")),]
entrez_g2t[entrez_g2t$gene_name == "TMSB15B",]
## ENSG00000269226 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000269226")),]

## Are there any dupped entrez id? 
dups_entrez <- entrez_g2t[which(duplicated(entrez_g2t$entrez_id)),]

entrez_g2t[entrez_g2t$entrez_id == "55486",]
## ENSG00000283765 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000283765")),]
entrez_g2t[entrez_g2t$entrez_id == "55300",]
## ENSG00000281028 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000281028")),]
entrez_g2t[entrez_g2t$entrez_id == "653082",]
## ENSG00000286094 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000286094")),]
entrez_g2t[entrez_g2t$entrez_id == "8622",]
## ENSG00000284762 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000284762")),]
entrez_g2t[entrez_g2t$entrez_id == "221468",]
## ENSG00000286105 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000286105")),]
entrez_g2t[entrez_g2t$entrez_id == "1124",]
## ENSG00000285162 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285162")),]
entrez_g2t[entrez_g2t$entrez_id == "102723849",]
## ENSG00000283765 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000273520")),]
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000262461")),]
entrez_g2t[entrez_g2t$entrez_id == "883",]
## ENSG00000286112 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000286112")),]
entrez_g2t[entrez_g2t$entrez_id == "22891",]
## ENSG00000261130 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285551")),]
entrez_g2t[entrez_g2t$entrez_id == "54816",]
## ENSG00000285253 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285253")),]
entrez_g2t[entrez_g2t$entrez_id == "23370",]
## ENSG00000268861 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000268861")),]
entrez_g2t[entrez_g2t$entrez_id == "284391",]
## ENSG00000286098 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000286098")),]
entrez_g2t[entrez_g2t$entrez_id == "440519",]
## ENSG00000283201 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000283201")),]
entrez_g2t[entrez_g2t$entrez_id == "149951",]
## ENSG00000285382 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285382")),]
entrez_g2t[entrez_g2t$entrez_id == "388820",]
## ENSG00000268861 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000278961")),]
entrez_g2t[entrez_g2t$entrez_id == "8471",]
## ENSG00000286098 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000260548")),]

## manual gene removal complete

load("../data/TCGA.RNA.Rda")

## Put entrez ids as rownames and shortening the TCGA ids since the last three portions of the TCGA barcode is coded to explain the test performed and etc. 
tumor_samples <- gdc_tumor_coding_complete %>% colData(.) %>% 
  as.data.frame(.) %>%
  dplyr::select(project_code, submitted_specimen_id) %>%
  mutate(submitted_specimen_id = substring(text = .$submitted_specimen_id, first = 1, last = 15))

TCGA.RNA <- TCGA.RNA %>% dplyr::select(-gene_symbol) %>%
  remove_rownames(.) %>% column_to_rownames(var = "gene_id") %>% 
  .[rownames(.) %in% entrez_g2t$entrez_id,] %>%
  'colnames<-' (substring(text = names(.), first = 1, last = 15)) %>% ## only need the first 15 character of barcode
  .[, names(.) %in% tumor_samples$submitted_specimen_id] ## only include tumor samples in our analysis

g2e_list <- entrez_g2t[entrez_g2t$entrez_id %in% rownames(TCGA.RNA),]

express_no_express_lookup <- list()
for(i in unique(tumor_samples$project_code)) {
  df <- which(tumor_samples$project_code == i) ## which rows are the spec cancer types
  df <- tumor_samples[df,] ## select only those rows
  gdc <- TCGA.RNA[, names(TCGA.RNA) %in% df$submitted_specimen_id] ## filter the norm_gdc_gene_exp dataset with those tumor samples
  mean_gdc <- as.data.frame(apply(gdc, MARGIN = 1, function(x) {
    mean(x)
  })) ## take the mean for each gene across the tumor samples
  names(mean_gdc) <- i
  express_no_express_lookup[[i]] <- mean_gdc ## store data in list
  rm(df, gdc, mean_gdc)
  gc
}
express_no_express_lookup <- do.call(cbind, express_no_express_lookup) ## cbind list
express_no_express_lookup[is.na(express_no_express_lookup)] <- 0

## 
gdc_ensembl_id <- gdc_tumor_coding_complete %>% rowRanges(.) %>% as.data.frame(.) %>% ## getting the ensembl_ids in our gdc dataset
  rownames(.)
g2e_list <- g2e_list[g2e_list$gene_id %in% gdc_ensembl_id,] ## subsetting the g2e_list to only have ensembl_ids in our gdc dataset

express_no_express_lookup <- express_no_express_lookup[rownames(express_no_express_lookup) %in% g2e_list$entrez_id, ]
express_no_express_lookup[express_no_express_lookup <= 10] <- 0 ## Set threshold to >10 for gene expression
express_no_express_lookup[express_no_express_lookup >10] <- 1

## convert entrez id to ensembl_id
express_no_express_lookup <- express_no_express_lookup %>% .[match(g2e_list$entrez_id, rownames(.)),] %>% ## match the entrez_id express_no_express_lookup with g2e_list
  'rownames<-' (g2e_list$gene_id) ## replace entrez_id with

## Booleanize mutation count for each tumor sample
missense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$missense) 
splice <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$splice)
frameshdel <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshdel)
frameshins <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshins)
utr3 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr3)
utr5 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr5)
nonstop <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonstop)
nonsense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonsense)

gdc_mut_count_patient <- missense + splice + frameshdel + frameshins + utr3 + utr5 + nonstop + nonsense ## add up all the mutation count
rm(missense, splice, frameshdel, frameshins, utr3, utr5, nonstop, nonsense)
gc()
gdc_mut_count_patient <- gdc_mut_count_patient[rownames(gdc_mut_count_patient) %in% rownames(express_no_express_lookup),] ## filter data with genes in the express_no_express_lookup
gdc_mut_count_patient[gdc_mut_count_patient > 0] <- 1 ## booleanize data, anything greater than 1 is equal to 1

gdc_gene_exp_adj <- list()
for(i in unique(tumor_samples$project_code)) {
  df <- which(tumor_samples$project_code == i) ## which rows are my spec cancer types
  df <- tumor_samples[df,] ## select only those rows
  gdc <- gdc_mut_count_patient[, names(gdc_mut_count_patient) %in% rownames(df)] ## filter the norm_gdc_gene_exp dataset with those tumor samples
  exp_data <- as.data.frame(express_no_express_lookup[, i]) ## select correct lookup column for the cancer type
  rownames(exp_data) <- rownames(express_no_express_lookup) ## Insert rownames to exp_data
  gdc <- gdc[match(rownames(exp_data), rownames(gdc)),] ## Match ensembl_id of exp_data with gdc data
  gdc_adj <- as.data.frame(apply(gdc, MARGIN = 2, function(x) {
    exp_data * x
  })) ## Function to multiply lookup to the data for the specific cancer type
  names(gdc_adj) <- names(gdc) ## Rename gdc_adj with tumor samples names
  gdc_gene_exp_adj[[i]] <- gdc_adj ## Store data to list
  rm(df, gdc, exp_data, gdc_adj)
  gc()
}
## Combining list into a dataframe.
gdc_gene_exp_adj <- do.call(cbind, gdc_gene_exp_adj) ## cbind list
names(gdc_gene_exp_adj) <- gsub("^.*\\.", "", names(gdc_gene_exp_adj)) ## tumor sample names have cancer type name in it. Removing it using gsub
rm(gdc_mut_count_patient, express_no_express_lookup, TCGA.RNA, gdc_tumor_coding_complete, gencode_gtf, dups_entrez, dups_gene_id, dups_gene_name, g2t, good.g2t, gene_info, g2e_list, entrez_g2t, gencode_entrez)
gc()
save(gdc_gene_exp_adj, file = "gdc_gene_exp_adj.Rda")
```
<div style="text-align: right"> Genes to pathway transformation </div>
```{r genes-to-pathways-transformation, eval=FALSE}
load("../data/gdc_gene_exp_adj.Rda")
load("../data/gene2_level2_lookup.Rda")
sample_list <- as.character(names(gdc_gene_exp_adj)) ## Create sample list
complete_path <- as.character(names(gene2level2_lookup)) ## Get names of pathways

df_gene <- as.data.frame(t(gdc_gene_exp_adj)) ## Transpose data so samples are rows and genes are columns
df_gene <- df_gene > 0 ## if 1 == TRUE if 0 == FALSE
df_gene <- as.data.frame(df_gene) ## turn matrix back to df

df_gene <- df_gene %>% rownames_to_column(var = "sample") %>% 
  gather(key = "gene", value = "mutation", -sample) %>% ## Wide to long format
  .[.$mutation,] %>% ## Keep TRUE's (which are 1's); Remove FALSE's (0's)
  dplyr::select(-mutation) ## Remove column named mutation

path_lookup <- as.data.frame(t(gene2level2_lookup)) ## Transpose data so paths are rows and genes are columns
path_lookup <- path_lookup > 0 ## if 1 == TRUE if 0 == FALSE
path_lookup <- as.data.frame(path_lookup) ## turn matrix back to df

path_lookup <- path_lookup %>% rownames_to_column(var = "path") %>%
  gather(key = "gene", value = "path_member", -path) %>% ## Wide to long format
  .[.$path_member,] %>% ## Keep TRUE's (which are 1's); Remove FALSE's (0's)
  dplyr::select(-path_member)

df_lookup <- df_gene %>% left_join(., path_lookup, by = "gene") %>% ## Match df_gene and lookup dataframes
  na.omit(.) ## Omit NA's that were generated due to no matches

## getting the genes involved in our analysis
genes_in_pathanalysis <- unique(df_lookup$gene)
write.table(genes_in_pathanalysis, file = "genes_in_pathanalysis.tsv", quote = FALSE, sep = "\t", row.names =  FALSE) ## save data

df_lookup <- as.data.frame(table(df_lookup$sample, df_lookup$path)) ## Generate 1's for the matches

df_lookup <- df_lookup %>% spread(key = "Var1", value = "Freq") %>% ## Long to wide format
  remove_rownames() %>%
  column_to_rownames(var = "Var2") ## path_id to rownames

## Insert missing tumor samples due to no mutations
df_missing <- sample_list[!(sample_list %in% names(df_lookup))] 
mx <- as.data.frame(matrix(data = 0, nrow = nrow(df_lookup), ncol = length(df_missing))) ## Create mx to store data
names(mx) <- df_missing ## Missing tumor sample names in col names
rownames(mx) <- rownames(df_lookup) ## Path names in rownames
df_lookup <- cbind(df_lookup, mx) ## cbind data

## Insert missing pathway due to no mutations
path_missing <- complete_path[!c(complete_path %in% rownames(df_lookup))]
mx <- as.data.frame(matrix(data = 0, nrow = length(path_missing), ncol = ncol(df_lookup))) ## Create mx to store data
names(mx) <- names(df_lookup) ## Insert tumor sample names
rownames(mx) <- path_missing ## Insert missing path names
df_lookup <- rbind(df_lookup, mx) ## cbind data

pathway_order <- rownames(df_lookup) ## need to perserve the order of pathway for the next operation
gdc_reactome_path_analysis <- df_lookup %>% mutate_if(is.numeric, ~1 * (. > 0)) %>% as.data.frame() %>% ## Mutate anything >0 to 1's (Booleanize data)
  'rownames<-' (pathway_order) %>%
  rownames_to_column(var = "PATH_ID") %>%
  left_join(., pathname2level2, by = "PATH_ID") %>%
  dplyr::select(-PATH_ID) %>% 
  remove_rownames() %>% column_to_rownames(var = "PATH_NAME")
rm(df_missing, path_missing, mx, df_gene, sample_list, complete_path, pathway_order, df_lookup)
gc()
save(gdc_reactome_path_analysis, file = "gdc_reactome_path_analysis.Rda")
```