### 2020-07-14 text/figure_4.Rmd

```{r figure-4-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_2d_3d_coors.tsv") ## load in data
load("../data/gene2_level2_lookup.Rda")
load("../data/pathname2level2.Rda")
load("../data/gdc_reactome_path_analysis.Rda")
load("../data/moderate_enrich_prob.Rda")
load("../data/gene_data.Rda")

## Parameters ##
filtered_class <- c(6,7) ## filtering for only class 6 and 7 samples
n_path <- 10 ## looking at the top10 path
enrich_cutoff <- 0.30
################

## Data Generation ##
class_spec <- tumor_clusters %>% dplyr::select(V1, knn_clust) %>% filter(knn_clust %in% filtered_class) %>% .[,.$V1] ## sample_id for class specific

toppath <- moderate_enrich_prob %>% filter(signif <= 2500 & difference >= enrich_cutoff) %>% ## signif is 95% CI with a specified cutoff
  filter(clust %in% filtered_class) %>% ## filter the cluster
  arrange(desc(difference)) %>% ## arrange enrichment score in descending order
  .[, "path"] %>% as.character() %>% unique() %>% .[1:n_path] ## getting the top10 enriched pathways

genes_in_path <- gene2level2_lookup %>% .[, match(pathname2level2$PATH_ID, names(.))] %>% ## match path ids
  'colnames<-' (pathname2level2$PATH_NAME) %>% ## replace with path_names
  dplyr::select(c(toppath)) %>% ## filter by top 10 pathways
  rownames_to_column(var = "ensembl_gene_id") %>% left_join(., gene_data[,c("ensembl_gene_id", "external_gene_name")], by = "ensembl_gene_id") %>% column_to_rownames(var = "external_gene_name") %>% dplyr::select(-ensembl_gene_id) %>% ## replace ensembl gene id with gene name
  .[rowSums(.[, -1])>0,]

df <- upset(genes_in_path, 
      sets = names(genes_in_path),
      mb.ratio = c(0.55, 0.45), 
      order.by = "freq", 
      point.size = 3.5,
      line.size = 2.0,
      mainbar.y.label = "Gene Intersection",
      sets.x.label = "Total Gene Per Pathway",
      text.scale = c(2.5, 2.4, 2, 2.1, 2.4, 1.5),
      number.angles = 0)

make_combinations <- function(x) {
  l <- length(x)
  mylist <- lapply(1:l, function(y) {
    combn(x, y, simplify = FALSE)
  })
  mylist
}

path_combo <- make_combinations(names(genes_in_path)) ## possible pathway combinations from 1 to 10

top10path <- genes_in_path %>% rowSums() %>% as.data.frame() %>% ## getting the sums of genes in pathway
  'colnames<-' ("sum_top10") %>% ## rename col
  rownames_to_column(var = "gene_name") ## make rownames a col

path_match_in_set <- list() ## making a list for matchings in sets

for(n in 1:10) {
  message(n)
  path_combo_set <- path_combo[[n]] ## selecting combination set (1 combo, 2 combo, 3 combo, etc)
  path_match_within_set <- list() ## making a list for matching within sets
  for(p in 1:length(path_combo_set)) {
    message(p)
    path_name <- path_combo_set[[p]] ## selecting specific combination pathway names
    spec_combo <- genes_in_path %>% dplyr::select(all_of(path_name)) ## select pathway combo
    if(length(spec_combo) > 1) {
      spec_combo <- spec_combo %>% rowSums() %>% as.data.frame() %>% ## rowSums the pathway to get the gene total membership
        'colnames<-' ("combo") %>% filter(combo == n) ## rename col and filter for combo that is n (n combo)
    }
    if(length(spec_combo) == 1) {
      names(spec_combo) <- "combo"
    }
    spec_combo <- spec_combo %>% rownames_to_column(var = "gene_name") %>% left_join(., top10path, by = "gene_name")
    spec_combo$match <- spec_combo$combo == spec_combo$sum_top10
    any_matches <- as.character(unique(spec_combo$match)) ## are there any matches?
    if(!"TRUE" %in% any_matches) { ## if no matches then skip
      next
      }
    if("TRUE" %in% any_matches) { ## if are matches then do this
      path_match_within_set[[p]] <- spec_combo %>% filter(match == "TRUE") %>% dplyr::select(gene_name) %>% ## filter for trues and select only gene names
        mutate(path_combo = paste0(paste(path_name, collapse=", "))) ## insert combo type
    }
  }    
  if(length(path_match_within_set) > 0) {
    path_match_in_set[[n]] <- path_match_within_set %>% do.call(rbind, .) %>% as.data.frame()
  }
}
path_match_in_set <- path_match_in_set %>% do.call(rbind, .) %>% as.data.frame() ## combine the list together

save(path_match_in_set, file = "path_match_in_set.Rda")
load("../data/path_match_in_set.Rda")

### getting clustering order from pheatmap
most_genes <- path_match_in_set %>% .$path_combo %>% table() %>% as.data.frame() %>% arrange(desc(Freq)) %>% ## arranging pathway combination based on most genes involved
  'colnames<-' (c("path_combo", "Freq")) %>% dplyr::select(path_combo) %>%
  rownames_to_column(var = "order")

gene_order <- path_match_in_set %>% left_join(., most_genes, by = "path_combo") %>% ## left join the data together
  arrange(order) %>% .$gene_name ## getting the gene name order

## now need to get the gene mutation data and tumor samples 
load("../data/gdc_tumor_coding_complete.Rdata")

## Booleanize mutation count for each tumor sample
missense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$missense) 
splice <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$splice)
frameshdel <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshdel)
frameshins <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshins)
utr3 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr3)
utr5 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr5)
nonstop <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonstop)
nonsense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonsense)

gdc_mut_count_patient <- missense + splice + frameshdel + frameshins + utr3 + utr5 + nonstop + nonsense ## add up all the mutation count
rm(missense, splice, frameshdel, frameshins, utr3, utr5, nonstop, nonsense, gdc_tumor_coding_complete)
gc()
gdc_mut_count_patient[gdc_mut_count_patient > 0] <- 1 ## booleanize data, anything greater than 1 is equal to 1

filtered_mut_data <- gdc_mut_count_patient %>% rownames_to_column(var = "ensembl_gene_id") %>% ## ensembl id to col
  left_join(., gene_data[,c("ensembl_gene_id", "external_gene_name")], by = "ensembl_gene_id") %>% ## left_join data by ensembl_gene_id
  column_to_rownames(var = "external_gene_name") %>% dplyr::select(-ensembl_gene_id) %>% ## replace ensembl_id with gene_name
  .[rownames(.) %in% gene_order, names(.) %in% class_spec] ## filter by genes in top10 path of filtered class and tumor samples associated with it
rm(gdc_mut_count_patient)
gc()

save(filtered_mut_data, gene_order, file = "filtered_mut_data.Rda")

fig4_pheatmap <- pheatmap(filtered_mut_data[gene_order,],
         show_rownames = T,
         show_colnames = F,
         treeheight_row = 0,
         treeheight_col = 0,
         cellheight = 5,
         cellwidth = 1,
         cluster_cols = T,
         cluster_rows = F, ## turn off row clustering
         annotation_legend = F,
         legend_breaks = c(0, 0.25, 0.75, 1),
         legend_labels = c("","Not Mut", "Mut", ""),
         color = c("grey", "firebrick3"),
         clustering_method = "ward.D2",
         fontsize = 10,
         legend = T)

## getting sample orders
pheatmap_order <- fig4_pheatmap$tree_col$order ## getting sample clustering order
sample_order <- names(filtered_mut_data) ## getting the sample_id original order 
sample_order <- sample_order[pheatmap_order] ## reorder cancer by heatmap order

## stopped here ## ignore bottom using as reference

class67_gene_2_path_heatmap <- filtered_mut_data %>% rownames_to_column(var = "gene_name") %>%  ## move rownames to column
  gather(., key = "sample_id", value = "membership", -gene_name) ## wide to long format
class67_gene_2_path_heatmap$membership <- as.character(class67_gene_2_path_heatmap$membership) ## turn boolean values into categories

## set levels based on clustering algorithm above
class67_gene_2_path_heatmap$gene_name <- factor(class67_gene_2_path_heatmap$gene_name, levels = c(rev(rownames(filtered_mut_data))))
class67_gene_2_path_heatmap$sample_id <- factor(class67_gene_2_path_heatmap$sample_id, levels = c(sample_order))

class67_gene_2_path_heatmap <- left_join(class67_gene_2_path_heatmap, path_match_in_set, by = "gene_name") ## add in path_combo for faceting 
save(class67_gene_2_path_heatmap, file = "class67_gene_2_path_heatmap.Rda") 
rm(list = ls(all.names = T))
gc()
```

```{r figure-4, message = FALSE}
load("../data/class67_gene_2_path_heatmap.Rda") ## load data
load("../data/path_match_in_set.Rda")

### getting facet ordering
most_genes <- path_match_in_set %>% .$path_combo %>% table() %>% as.data.frame() %>% arrange(desc(Freq)) %>% ## arranging pathway combination based on most genes involved
  'colnames<-' (c("path_combo", "Freq")) %>% 
  rownames_to_column(var = "order")

gene_order <- path_match_in_set %>% left_join(., most_genes, by = "path_combo") %>% ## left join the data together
  arrange(order) %>% .$gene_name ## getting the gene name order

class67_gene_2_path_heatmap$gene_name <- factor(class67_gene_2_path_heatmap$gene_name, levels = c(rev(gene_order)))

fig4 <- ggplot(data = class67_gene_2_path_heatmap, aes(y=gene_name, x=sample_id, fill = membership)) + 
  geom_tile(size = 0.6) +
  scale_fill_manual(values = c("grey", "firebrick3")) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        legend.position = "none")

load("../data/path_match_in_set.Rda")

gene_order <- path_match_in_set %>% left_join(., most_genes, by = "path_combo") %>% ## left join the data together
  arrange(order) %>% .$gene_name ## getting the gene name order

## right annotation data generation
proteasome_r_anno <- data.frame(gene = gene_order, 
                                psm_gene = 0)

proteasome_r_anno[grep(pattern = "PSM|UBE2|UBB|UBC", x = proteasome_r_anno$gene), "psm_gene"] <- 1

r_anno_color <- c("0" = "white", "1" = "black")
r_anno <- ggplot(proteasome_r_anno) +
  geom_bar(mapping = aes(x = 1, y = gene, fill = as.character(psm_gene)), 
           stat = "identity", 
           width = 1) +
  scale_fill_manual(values = r_anno_color, breaks = "1", name = "", labels = "PSM Gene") +
  theme_void() 
legend <- plot_grid(get_legend(r_anno), ncol = 1)
r_anno <- r_anno + theme(legend.position = "none")

## left annotatoin data generation
l_anno <- path_match_in_set %>% .[match(gene_order, .$gene_name),] ## order by the gene order from the heatmap

## need to get top path names again
load("../data/moderate_enrich_prob.Rda")
enrich_cutoff <- 0.30
filtered_class <- c(6,7)
n_path <- 10
toppath <- moderate_enrich_prob %>% filter(signif <= 2500 & difference >= enrich_cutoff) %>% ## signif is 95% CI with a specified cutoff
  filter(clust %in% filtered_class) %>% ## filter the cluster
  arrange(desc(difference)) %>% ## arrange enrichment score in descending order
  .[, "path"] %>% as.character() %>% unique() %>% .[1:n_path] ## getting the top10 enriched pathways

for(p in toppath){
  message(p)
  l_anno <- l_anno %>% mutate(p = 0) ## create a new col named p
  names(l_anno)[names(l_anno) == 'p'] <- p ## rename col with pathway name
  ## for any path_combo that includes the pathway name change the 0 into a 1
  if(p == "AUF1 (hnRNP D0) binds and destabilizes mRNA") { ## if it has this pathway grep a different way
    l_anno[grep(pattern = "AUF1", x = l_anno$path_combo), p] <- 1
  }
  if(p != "AUF1 (hnRNP D0) binds and destabilizes mRNA") { ## if it is not AUF1 do it normally
  l_anno[grep(pattern = p, x = l_anno$path_combo), p] <- 1 
  }
}

l_anno_ggplot <- list() ## list for storing ggplots
path_legend <- list()

kelly.colours <- c("gray95", "gray13", "gold2", "plum4", "darkorange1",
                   "lightskyblue2", "firebrick", "burlywood3", "gray51", "springgreen4",
                   "lightpink2", "deepskyblue4", "lightsalmon2", "mediumpurple4",
                   "orange", "maroon", "yellow3", "brown4", "yellow4", "sienna4", "chocolate", "gray19")

for(n in 1:length(toppath)){
  message(n)
  p <- toppath[n]
  message(p)
  spec_path <- l_anno[,c("gene_name", p)]
  names(spec_path)[names(spec_path) == p] <- "p"
  anno_color <- c("0" = "white", "1" = kelly.colours[2+n])
  path_ggplot <- ggplot(spec_path) +
  geom_bar(mapping = aes(x = 1, y = gene_name, fill = as.character(p)), 
           stat = "identity", 
           width = 1) +
  scale_fill_manual(values = anno_color, breaks = "1", name = "", labels = p) +
  theme_void() 
  path_legend[[n]] <- plot_grid(get_legend(path_ggplot), ncol = 1) ## extract legend from plot
  l_anno_ggplot[[n]] <- path_ggplot + theme(legend.position = "none") ## remove legend from plot
}

plot <- plot_grid(l_anno_ggplot[[1]], l_anno_ggplot[[2]], l_anno_ggplot[[3]], l_anno_ggplot[[4]], l_anno_ggplot[[5]], l_anno_ggplot[[6]], l_anno_ggplot[[7]], l_anno_ggplot[[8]], l_anno_ggplot[[9]], l_anno_ggplot[[10]], fig4, r_anno, align = "h", ncol = 12, axis = "tb", rel_widths = c(rep(0.1, 10), 15, 0.1))
plot_legend <- plot_grid(path_legend[[1]], path_legend[[2]], path_legend[[3]], path_legend[[4]], path_legend[[5]], path_legend[[6]], path_legend[[7]], path_legend[[8]], path_legend[[9]], path_legend[[10]], ncol = 5)

plot_grid(plot, plot_legend, ncol = 1, rel_heights = c(10, 1.5))

```