### 2020-07-14 text/figure_4_function.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r figure-4-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_2d_3d_coors.tsv") ## load in data
load("../data/gene2_level2_lookup.Rda")
load("../data/pathname2level2.Rda")
load("../data/gdc_reactome_path_analysis.Rda")
load("../data/moderate_enrich_prob.Rda")
load("../data/gene_data.Rda")
# load("../data/gdc_tumor_coding_complete.Rdata")
# 
# missense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$missense) 
# splice <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$splice)
# frameshdel <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshdel)
# frameshins <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshins)
# utr3 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr3)
# utr5 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr5)
# nonstop <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonstop)
# nonsense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonsense)
# 
# gdc_mut_count_patient <- missense + splice + frameshdel + frameshins + utr3 + utr5 + nonstop + nonsense ## add up all the mutation count
# rm(missense, splice, frameshdel, frameshins, utr3, utr5, nonstop, nonsense, gdc_tumor_coding_complete)
# gc()
# gdc_mut_count_patient[gdc_mut_count_patient > 0] <- 1 ## booleanize data, anything greater than 1 is equal to 1
# save(gdc_mut_count_patient, file = "../data/gdc_mut_count_patient_booleanized.Rda")
load("../data/gdc_mut_count_patient_booleanized.Rda")
filtered_class <- 12
n_path <- 10
enrich_cutoff <- 0.30
oncoprint_style <- function(filtered_class, n_path, enrich_cutoff = 0.30) {
  class_spec <- tumor_clusters %>% dplyr::select(V1, knn_clust) %>% filter(knn_clust %in% filtered_class) %>% .[,.$V1] ## sample_id for class specific
  toppath <- moderate_enrich_prob %>% filter(signif <= 2500 & difference >= enrich_cutoff) %>% ## signif is 95% CI with a specified cutoff
  filter(clust %in% filtered_class) %>% ## filter the cluster
  arrange(desc(difference)) %>% ## arrange enrichment score in descending order
  .[, "path"] %>% as.character() %>% unique() %>% .[1:n_path] %>% na.omit() ## getting the top10 enriched pathways
  if(length(toppath) == 0) { ## if no enriched pathways. stop and print error message
    stop("No enriched pathways... select a lower enrich_cutoff")
  }
  if(length(toppath) == 1) { ## if there is only 1 enriched path. stop and print error message
    stop("1 enriched pathway found. Use oncoprint to do your analysis")
  }
  genes_in_path <- gene2level2_lookup %>% .[, match(pathname2level2$PATH_ID, names(.))] %>% ## match path ids
  'colnames<-' (pathname2level2$PATH_NAME) %>% ## replace with path_names
  dplyr::select(c(toppath)) %>% ## filter by top 10 pathways
  rownames_to_column(var = "ensembl_gene_id") %>% left_join(., gene_data[,c("ensembl_gene_id", "external_gene_name")], by = "ensembl_gene_id") %>% column_to_rownames(var = "external_gene_name") %>% dplyr::select(-ensembl_gene_id) %>% ## replace ensembl gene id with gene name
  .[rowSums(.[, -1])>0,]

  make_combinations <- function(x) { ## function to make all the possible combination of genes in pathway
  l <- length(x)
  mylist <- lapply(1:l, function(y) {
    combn(x, y, simplify = FALSE)
  })
  mylist
  }
  path_combo <- make_combinations(names(genes_in_path)) ## possible pathway combinations from 1 to 10

  top10path <- genes_in_path %>% rowSums() %>% as.data.frame() %>% ## getting the sums of genes in pathway
    'colnames<-' ("sum_top10") %>% ## rename col
    rownames_to_column(var = "gene_name") ## make rownames a col

  path_match_in_set <- list() ## making a list for matchings in sets

  for(n in 1:length(path_combo)) {
    message(n)
    path_combo_set <- path_combo[[n]] ## selecting combination set (1 combo, 2 combo, 3 combo, etc)
    path_match_within_set <- list() ## making a list for matching within sets
    for(p in 1:length(path_combo_set)) {
      message(p)
      path_name <- path_combo_set[[p]] ## selecting specific combination pathway names
      spec_combo <- genes_in_path %>% dplyr::select(all_of(path_name)) ## select pathway combo
      if(length(spec_combo) > 1) {
        spec_combo <- spec_combo %>% rowSums() %>% as.data.frame() %>% ## rowSums the pathway to get the gene total membership
        'colnames<-' ("combo") %>% filter(combo == n) ## rename col and filter for combo that is n (n combo)
      }
      if(length(spec_combo) == 1) {
      names(spec_combo) <- "combo"
      }
      spec_combo <- spec_combo %>% rownames_to_column(var = "gene_name") %>% left_join(., top10path, by = "gene_name")
      spec_combo$match <- spec_combo$combo == spec_combo$sum_top10
      any_matches <- as.character(unique(spec_combo$match)) ## are there any matches?
      if(!"TRUE" %in% any_matches) { ## if no matches then skip
        next
        }
      if("TRUE" %in% any_matches) { ## if are matches then do this
        path_match_within_set[[p]] <- spec_combo %>% filter(match == "TRUE") %>% dplyr::select(gene_name) %>% ## filter for trues and select only gene names
        mutate(path_combo = paste0(paste(path_name, collapse=", "))) ## insert combo type
      }
   }    
    if(length(path_match_within_set) > 0) {
      path_match_in_set[[n]] <- path_match_within_set %>% do.call(rbind, .) %>% as.data.frame()
    }
  }
  path_match_in_set <- path_match_in_set %>% do.call(rbind, .) %>% as.data.frame() ## combine the list together
  ### getting clustering order from pheatmap
  most_genes <- path_match_in_set %>% .$path_combo %>% table() %>% as.data.frame() %>% arrange(desc(Freq)) %>% ## arranging pathway combination based on most genes involved
    'colnames<-' (c("path_combo", "Freq")) %>% dplyr::select(path_combo) %>%
    rownames_to_column(var = "order")

  gene_order <- path_match_in_set %>% left_join(., most_genes, by = "path_combo") %>% ## left join the data together
  arrange(order) %>% .$gene_name ## getting the gene name order

  ## gene mutation data and tumor samples - Booleanize mutation count for each tumor sample
  message("working on mutation data")
  filtered_mut_data <- gdc_mut_count_patient %>% rownames_to_column(var = "ensembl_gene_id") %>% ## ensembl id to col
    left_join(., gene_data[,c("ensembl_gene_id", "external_gene_name")], by = "ensembl_gene_id") %>% ## left_join data by ensembl_gene_id
    column_to_rownames(var = "external_gene_name") %>% dplyr::select(-ensembl_gene_id) %>% ## replace ensembl_id with gene_name
    .[rownames(.) %in% gene_order, names(.) %in% class_spec] ## filter by genes in top10 path of filtered class and tumor samples associated with it
  message("Running pheatmap to get coordinates")
  pheatmap_res <- pheatmap(filtered_mut_data[gene_order,],
         show_rownames = T,
         show_colnames = F,
         treeheight_row = 0,
         treeheight_col = 0,
         cellheight = 5,
         cellwidth = 1,
         cluster_cols = T,
         cluster_rows = F, ## turn off row clustering
         annotation_legend = F,
         legend_breaks = c(0, 0.25, 0.75, 1),
         legend_labels = c("","Not Mut", "Mut", ""),
         color = c("grey", "firebrick3"),
         clustering_method = "ward.D2",
         fontsize = 10,
         legend = T)

  pheatmap_order <- pheatmap_res$tree_col$order ## getting sample clustering order
  sample_order <- names(filtered_mut_data) ## getting the sample_id original order 
  sample_order <- sample_order[pheatmap_order] ## reorder cancer by heatmap order
  message("Preparing data for plot")
  gene_2_path_heatmap <- filtered_mut_data %>% rownames_to_column(var = "gene_name") %>%  ## move rownames to column
  gather(., key = "sample_id", value = "membership", -gene_name) ## wide to long format
  gene_2_path_heatmap$membership <- as.character(gene_2_path_heatmap$membership) ## turn boolean values into categories
  ## set levels based on clustering algorithm above
  gene_2_path_heatmap$gene_name <- factor(gene_2_path_heatmap$gene_name, levels = c(rev(gene_order)))
  gene_2_path_heatmap$sample_id <- factor(gene_2_path_heatmap$sample_id, levels = c(sample_order))

  gene_2_path_heatmap <- left_join(gene_2_path_heatmap, path_match_in_set, by = "gene_name") ## add in path_combo for faceting 
  heatmap_ggplot <- ggplot(data = gene_2_path_heatmap, aes(y=gene_name, x=sample_id, fill = membership)) + 
  geom_tile(size = 0.6) +
  scale_fill_manual(values = c("grey", "firebrick3")) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        legend.position = "none") 
  ## left annotatoin data generation
  l_anno <- path_match_in_set %>% .[match(gene_order, .$gene_name),] ## order by the gene order from the heatmap

  for(p in toppath){
    message(p)
    l_anno <- l_anno %>% mutate(p = 0) ## create a new col named p
    names(l_anno)[names(l_anno) == 'p'] <- p ## rename col with pathway name
    ## for any path_combo that includes the pathway name change the 0 into a 1
    if(p == "AUF1 (hnRNP D0) binds and destabilizes mRNA") { ## if it has this pathway grep a different way
     l_anno[grep(pattern = "AUF1", x = l_anno$path_combo), p] <- 1
    }
    if(p == "Signaling by Type 1 Insulin-like Growth Factor 1 Receptor (IGF1R)") { ## if it has this pathway grep a different way
     l_anno[grep(pattern = "Signaling by Type 1 Insulin", x = l_anno$path_combo), p] <- 1
    }
    if(p != "AUF1 (hnRNP D0) binds and destabilizes mRNA" & p != "Signaling by Type 1 Insulin-like Growth Factor 1 Receptor (IGF1R)") { ## if it is not AUF1 do it normally
    l_anno[grep(pattern = p, x = l_anno$path_combo), p] <- 1 
    }
  }

  l_anno_ggplot <- list() ## list for storing ggplots
  path_legend <- list()

  kelly.colours <- c("gray95", "gray13", "gold2", "plum4", "darkorange1",
                     "lightskyblue2", "firebrick", "burlywood3", "gray51", "springgreen4",
                     "lightpink2", "deepskyblue4", "lightsalmon2", "mediumpurple4",
                     "orange", "maroon", "yellow3", "brown4", "yellow4", "sienna4", "chocolate", "gray19")

  for(n in 1:length(toppath)){
    message(n)
    p <- toppath[n]
    message(p)
    spec_path <- l_anno[,c("gene_name", p)]
    names(spec_path)[names(spec_path) == p] <- "p"
    anno_color <- c("0" = "white", "1" = kelly.colours[2+n])
    path_ggplot <- ggplot(spec_path) +
    geom_bar(mapping = aes(x = 1, y = gene_name, fill = as.character(p)), 
            stat = "identity", 
            width = 1) +
    scale_fill_manual(values = anno_color, breaks = "1", name = "", labels = p) +
    theme_void() 
    path_legend[[n]] <- plot_grid(get_legend(path_ggplot), ncol = 1) ## extract legend from plot
    l_anno_ggplot[[n]] <- path_ggplot + theme(legend.position = "none") ## remove legend from plot
  }

  plot <- plot_grid(l_anno_ggplot[[1]], l_anno_ggplot[[2]], l_anno_ggplot[[3]], l_anno_ggplot[[4]], l_anno_ggplot[[5]], l_anno_ggplot[[6]], l_anno_ggplot[[7]], l_anno_ggplot[[8]], l_anno_ggplot[[9]], l_anno_ggplot[[10]], heatmap_ggplot, align = "h", ncol = 11, axis = "tb", rel_widths = c(rep(0.1, 10), 15))
plot_legend <- plot_grid(path_legend[[1]], path_legend[[2]], path_legend[[3]], path_legend[[4]], path_legend[[5]], path_legend[[6]], path_legend[[7]], path_legend[[8]], path_legend[[9]], path_legend[[10]], ncol = 5)

return(plot_grid(plot, plot_legend, ncol = 1, rel_heights = c(10, 1.5)))
}

test1 <- oncoprint_style(filtered_class = c(6,7), n_path = 10)
test12 <- oncoprint_style(filtered_class = 12, n_path = 10)
test11 <- oncoprint_style(filtered_class = 11, n_path = 10)

```