### 2020-07-14 text/figure_s2.Rmd

```{r supplementary-figure-s2-data-generation}
umap_cancertype <- fread("../data/umap_2d_3d_coors.tsv") ## load data
umap_bg <- umap_cancertype[,-"project_code"] ## get background data for facet

ggplot(umap_cancertype, aes(x = x_2d, y = y_2d, colour = Subtype_Selected)) + 
  geom_point(data = umap_bg, aes(x = x_2d, y = y_2d), colour = "grey", size = 0.5) +
  geom_point(size = 0.5) +
  theme_classic() +
  coord_fixed() +
  xlab("UMAP 1") + ylab("UMAP 2") +
  theme(axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    legend.position = "none") +
  facet_wrap(~ project_code, ncol = 5) ## facet by cancer type
```