### 2020-07-20 code/figure1.Rmd
Change all references to umap_2d_3d_coors to umap_3d_coors ; also alter all references to V1 to read sample_id

<div style="text-align: right"> Generation of Somatic Mutation Data </div>
```{r data-set-creation, eval=FALSE}
## Data acquired from GDC
tcga_ssm <- fread("../data/simple_somatic_mutation.open.tsv.gz", header = T, sep = "\t")
## Somatic mutations focused in analysis
selected_mut <- c("3_prime_UTR_variant", "5_prime_UTR_premature_start_codon_gain_variant", "5_prime_UTR_variant", "frameshift_variant", "missense_variant", "splice_acceptor_variant", "splice_donor_variant", "stop_lost", "start_lost", "stop_gained")
## Filtering tcga_ssm for our selected mutations
tcga_ssm <- tcga_ssm[tcga_ssm$consequence_type %in% selected_mut, ]
## Selecting column that we are interested in
tcga_ssm <- data.frame(chr = paste0("chr",tcga_ssm$chromosome),
                   start = tcga_ssm$chromosome_start,
                   end = tcga_ssm$chromosome_end,
                   ref_allele = tcga_ssm$reference_genome_allele,
                   alt_allele = tcga_ssm$mutated_to_allele,
                   type = tcga_ssm$consequence_type,
                   mut_type = tcga_ssm$mutation_type,
                   ensembl_id = tcga_ssm$gene_affected,
                   donor_id = tcga_ssm$icgc_donor_id,
                   sample_id = tcga_ssm$submitted_sample_id,
                   stringsAsFactors = F)
## Removing duplicate rows
tcga_ssm <- tcga_ssm %>% distinct()
## Recategorizing the mutation type
tcga_ssm[which(tcga_ssm$type == "frameshift_variant"), "type"] <- tcga_ssm[which(tcga_ssm$type == "frameshift_variant"), "mut_type"] 
tcga_ssm <- tcga_ssm %>% mutate(type = case_when(type == "insertion of <=200bp" ~ "frameshins_variant", ## relabel values
                                                 type == "deletion of <=200bp" ~ "frameshdel_variant",
                                                 type == "splice_acceptor_variant" ~ "splice_variant",
                                                 type == "splice_donor_variant" ~ "splice_variant",
                                                 type == "start_lost" ~ "nonsense_variant",
                                                 type == "stop_gained" ~ "nonsense_variant",
                                                 type == "5_prime_UTR_premature_start_codon_gain_variant" ~ "UTR_5_prime_variant",
                                                 type == "5_prime_UTR_variant" ~ "UTR_5_prime_variant",
                                                 type == "3_prime_UTR_variant" ~ "UTR_3_prime_variant",
                                                 type == "missense_variant" ~ "missense_variant",
                                                 type == "stop_lost" ~ "nonstop_variant")) %>% 
  dplyr::select(-mut_type) ## rm mut_type col

## Preparing sample metadata ##

complete_sample <- tcga_ssm %>% .[,"sample_id"] %>% unique() %>% .[grep(pattern= "TCGA-", x = .)] # complete TCGA tumor samples

tcga_sample <- fread("../data/sample.tsv.gz", header = T, sep = "\t")
tcga_sample <- tcga_sample %>% dplyr::select(project_code, submitted_sample_id, submitted_specimen_id, icgc_donor_id, submitted_donor_id, study) %>% 
  .[.$submitted_sample_id %in% complete_sample,] %>%
  mutate(type = substring(.$submitted_sample_id, first = 14, last = 15)) %>% ## 01 is primary solid tumor
  filter(type == "01") ## filtering 01 samples
tcga_dup_id <- tcga_sample %>% .[.$icgc_donor_id %in% .[which(duplicated(.$submitted_donor_id)),]$icgc_donor_id,] %>% ## what are the dup donors
  filter(study == "PCAWG") ## selecting dup samples that are PCAWG
tcga_sample <- tcga_sample %>% .[!.$icgc_donor_id %in% tcga_dup_id$icgc_donor_id, ] %>% ## remove the dups from tcga_samples
  rbind(., tcga_dup_id) ## rbind the fixed dups

tcga_donor <- fread("../data/donor.tsv.gz", header = T, sep = "\t")
tcga_donor <- tcga_donor %>% dplyr::select(icgc_donor_id, project_code, submitted_donor_id, donor_sex, donor_vital_status, disease_status_last_followup, donor_age_at_diagnosis, donor_age_at_enrollment, donor_age_at_last_followup, donor_survival_time, donor_interval_of_last_followup) ## selecting columns that we are interested in

sample_sheet <- left_join(tcga_sample, tcga_donor, by = c("submitted_donor_id", "project_code", "icgc_donor_id"))

## Preparing gene metadata ##

# genelist <- unique(as.character(tcga_ssm$ensembl_id)) ## complete genelist
# ensembl_2_entrez_full_list <- bitr(genelist, 
#                                    fromType = "ENSEMBL", 
#                                    toType = c("ENTREZID"), OrgDb = org.Hs.eg.db)
load("../data/ensembl_2_entrez_full_list.Rda")
tcga_ssm <- tcga_ssm[tcga_ssm$ensembl_id %in% ensembl_2_entrez_full_list$ENSEMBL,]

## Now we only have 19067 genes
complete_ensembl <- as.data.frame(unique(as.character(tcga_ssm$ensembl_id)))
names(complete_ensembl) <- "ensembl_id"

selected_mut <- unique(as.character(tcga_ssm$type))
## For every mutation type.. select them.. and create a tsv file 
for(selected_mut in selected_mut) {
  print(selected_mut)
  df <-  tcga_ssm %>% filter(type == selected_mut) %>% ## filter data by selected_mut
    dplyr::select(sample_id, ensembl_id) %>% ## select column sample_id and ensembl_id
    mutate(mutation = 1) %>% ## create a new column with 1's
    distinct() %>% ## remove duplicated rows
    spread(., key = "sample_id", value = "mutation") %>% ## long to wide format
    remove_rownames() %>% ## move ensembl_id column to rownames
    column_to_rownames(var = "ensembl_id") %>% 
    replace(is.na(.), values = 0) %>% ## convert NA's to 0's
    .[, names(.) %in% sample_sheet$submitted_sample_id] ## rm any sample_id not included in complete_sample
  missing_gene <- as.data.frame(matrix(data = 0, nrow = length(complete_ensembl$ensembl_id[!complete_ensembl$ensembl_id %in% rownames(df)]), ncol = length(df)))
  rownames(missing_gene) <- complete_ensembl$ensembl_id[!complete_ensembl$ensembl_id %in% rownames(df)] ## insert missing ensembl_id into row names
  names(missing_gene) <- names(df) ## insert sample_id names into col names
  df <- rbind(df, missing_gene) ## rbind df and missing_gene
  rm(missing_gene) ## rm data
  gc()
  df <- df[match(complete_ensembl$ensembl_id, rownames(df)),] ## match df rownames with genelist
  missing_sample <- as.data.frame(matrix(data = 0, nrow = nrow(df), ncol = length(sample_sheet$submitted_sample_id[!sample_sheet$submitted_sample_id %in% names(df)])))
  rownames(missing_sample) <- rownames(df) ## insert ensembl_id into row names
  names(missing_sample) <- sample_sheet$submitted_sample_id[!sample_sheet$submitted_sample_id %in% names(df)] ## insert missing sample_id names into col names
  df <- cbind(df, missing_sample) ## rbind df and missing_gene
  rm(missing_sample) ## rm data
  gc()
  df <- df[, match(sample_sheet$submitted_sample_id, names(df))]
  assign(paste0(selected_mut), value = df) ## create df in r environment
  rm(df)
  gc()
  }

# ensembl <- useEnsembl(biomart="ensembl", dataset = "hsapiens_gene_ensembl")
 # genelist <- rownames(missense_variant)
 # gene_data <- getBM(
 #   attributes = c(
 #     'ensembl_gene_id',
 #     'external_gene_name',
 #     'gene_biotype',
 #     'chromosome_name', 
 #     'start_position',
 #     'end_position', 
 #     'strand'),
 #   filters = 'ensembl_gene_id',
 #   values = genelist,
 #   mart = ensembl)
 # save(gene_data, file = "gene_data.Rda")
load("../data/gene_data.Rda")
######################

missense_variant <- missense_variant[rownames(missense_variant) %in% gene_data$ensembl_gene_id,]
UTR_3_prime_variant <- UTR_3_prime_variant[rownames(UTR_3_prime_variant) %in% gene_data$ensembl_gene_id,]
UTR_5_prime_variant <- UTR_5_prime_variant[rownames(UTR_5_prime_variant) %in% gene_data$ensembl_gene_id,]
frameshdel_variant <- frameshdel_variant[rownames(frameshdel_variant) %in% gene_data$ensembl_gene_id,]
frameshins_variant <- frameshins_variant[rownames(frameshins_variant) %in% gene_data$ensembl_gene_id,]
splice_variant <- splice_variant[rownames(splice_variant) %in% gene_data$ensembl_gene_id,]
nonstop_variant <- nonstop_variant[rownames(nonstop_variant) %in% gene_data$ensembl_gene_id,]
nonsense_variant <- nonsense_variant[rownames(nonsense_variant) %in% gene_data$ensembl_gene_id,]

## Matching df to the ensembl_id orders in gene_data
match_ensembl <- function(x) {
  x <- x[match(gene_data$ensembl_gene_id, rownames(x)),]
}

missense_variant <- match_ensembl(missense_variant)
UTR_3_prime_variant <- match_ensembl(UTR_3_prime_variant)
UTR_5_prime_variant <- match_ensembl(UTR_5_prime_variant)
frameshdel_variant <- match_ensembl(frameshdel_variant)
frameshins_variant <- match_ensembl(frameshins_variant)
splice_variant <- match_ensembl(splice_variant)
nonstop_variant <- match_ensembl(nonstop_variant)
nonsense_variant <- match_ensembl(nonsense_variant)

genes <- GRanges(
  seqnames = paste("chr", gene_data$chromosome_name, sep=""),
  ranges = IRanges(
    start = gene_data$start_position,
    end = gene_data$end_position),
  strand = gene_data$strand,
  mcols = list(
    ensembl_gene_id = gene_data$ensembl_gene_name,
    hgnc_symbol = gene_data$hgnc_symbol))

gdc_tumor_coding_complete <- SummarizedExperiment(
  assays = list(
    missense   = missense_variant,
    splice     = splice_variant,
    frameshdel = frameshdel_variant,
    frameshins = frameshins_variant,
    utr3       = UTR_3_prime_variant,
    utr5       = UTR_5_prime_variant,
    nonsense   = nonsense_variant,
    nonstop    = nonstop_variant),
  rowRanges = genes,
  colData = sample_sheet)
save(gdc_tumor_coding_complete, file = "../data/gdc_tumor_coding_complete.Rdata")
rm(list = ls(all.names = T))
gc()
```
<div style="text-align: right"> Reactome Pathway Creation </div>
```{r reactome-pathway-creation, eval=FALSE}
load("../data/gdc_tumor_coding_complete.Rdata")
load("../data/gene_data.Rda")
## Preparing level_0 to level_1 pathway conversion table. level_0 is the master pathway. level_1 is the parent pathway
main2first <- fread(file = "../data/reactome_main2first_level.txt", header = T, sep = "\t")

main2first <- main2first %>% .[,2:3] %>%
  'colnames<-' (c("first_level", "main_level")) ## rename colnames

## Filtering out master pathways from first2second_level
first2second <- fread(file = "../data/reactome_first2second_level.txt", header = F, sep = "\t")
first2second <- first2second %>% 'colnames<-' (c("first_level", "second_level")) %>% ## rename colnames
  .[grep(pattern = "-HSA-", .$first_level)] %>% ## keep human pathways
  .[!.$first_level %in% unique(main2first$main_level),] %>% ## remove main level pathways from first level
  .[!.$first_level %in% unique(.$second_level),] ## remove second level pathways from first level

ensembl2reactome <- fread(file = "../data/Ensembl2Reactome_All_Levels.txt", header = F, sep = "\t") ## ensembl to pathway list acquired from reactome database
ensembl2reactome <- ensembl2reactome %>% .[,c(1:2,4)] %>% ## select columns ensembl, pathway_id, pathway_name columns
  'colnames<-' (c("ENSEMBL", "PATH_ID", "PATH_NAME")) %>% ## rename colnames
  .[grep(pattern = "-HSA-", .$PATH_ID),] %>% ## keep human pathways
  .[.$PATH_ID %in% first2second$second_level,] ## include only second level pathways

## Removing pathways that are not molecular functions such as pathways that related to disease, defect, bacteria, disorder, myosin and etc
rm_pathways <- c("R-HSA-1222499", "R-HSA-1226099", "R-HSA-1300642", "R-HSA-1300645", "R-HSA-162906",  "R-HSA-1643713", "R-HSA-168254",
                 "R-HSA-212436", "R-HSA-2160456", "R-HSA-2219528", "R-HSA-2474795", "R-HSA-2534343", "R-HSA-2644603", "R-HSA-2978092",
                 "R-HSA-3296482", "R-HSA-3304351", "R-HSA-3560782", "R-HSA-3781860", "R-HSA-388396",  "R-HSA-3906995", "R-HSA-4791275",
                 "R-HSA-5339562", "R-HSA-5387390", "R-HSA-5545483", "R-HSA-5576886", "R-HSA-5576890", "R-HSA-5576892", "R-HSA-5576893",
                 "R-HSA-5576894", "R-HSA-5578768", "R-HSA-5578775", "R-HSA-5579029", "R-HSA-5602358", "R-HSA-5609975", "R-HSA-5619084",
                 "R-HSA-5619102", "R-HSA-5632927", "R-HSA-5632928", "R-HSA-5632968", "R-HSA-5632987", "R-HSA-5663084", "R-HSA-5687613",
                 "R-HSA-6802957", "R-HSA-8862803", "R-HSA-8876384", "R-HSA-9005891", "R-HSA-9605310", "R-HSA-9616333", "R-HSA-9616334",
                 "R-HSA-9629232", "R-HSA-9630750", "R-HSA-9632693", "R-HSA-9645722", "R-HSA-9646303", "R-HSA-9646304")

## Ensembl to reactome pathway lookup table
gene2level2_lookup <- ensembl2reactome %>% dplyr::select(-PATH_NAME) %>% ## remove PATH_NAME col
  .[.$ENSEMBL %in% gene_data$ensembl_gene_id,] %>% ## only include genes that are in the gene_data 
  mutate(member = 1) %>% ## add a column for ensembl to pathway membership
  distinct() %>% ## keep unique rows
  spread(data = ., key = "PATH_ID", value = "member") %>% ## long to wide format
  column_to_rownames(var = "ENSEMBL") %>% ## ensembl_id as rownames
  .[,!names(.) %in% rm_pathways] %>% ## filter rm_pathways from lookup table
  replace(is.na(.), values = 0) ## replace NA's with 0's
#save(gene2level2_lookup, file = "gene2_level2_lookup.Rda")
## Pathway names to level 2 lookup table
pathname2level2 <- ensembl2reactome %>% dplyr::select(-ENSEMBL) %>% ## remove ENSEMBL col
  .[.$PATH_ID %in% names(gene2level2_lookup),] %>% ## only include pathways in gene2level2_lookup
  distinct(.) ## keep unique rows
#save(pathname2level2, file = "pathname2level2.Rda")
rm(rm_pathways, ensembl2reactome, main2first, first2second)
gc()
```
<div style="text-align: right"> Filter tissue-specific low expressed genes </div>
```{r filter-tissue-specific-low-expressed-genes, eval=FALSE}
gtf_file <- "../data/gencode.v30.annotation.gtf"
gencode_gtf <- rtracklayer::import(gtf_file)
gene_info <- as.data.frame(mcols(gencode_gtf)[, c("type", "gene_id", "transcript_id", "gene_type", "transcript_type", "gene_name")])#; rm(gencode_gtf)
## We only want transcript only 
gene_info <- gene_info[gene_info$type == "transcript",]
## We only want protein_coding only
gene_info <- gene_info[gene_info$gene_type == "protein_coding",]
gene_info <- gene_info[gene_info$transcript_type == "protein_coding",]
## Filtering for transcript_id, gene_id, gene_name columns
g2t <- gene_info %>% dplyr::filter(type == "transcript") %>% dplyr::select(transcript_id, gene_id, gene_name)
## Condense df to only having unique data. 
good.g2t <- as_tibble(unique(g2t))

## df with transcript_id and entrez id                      
gencode_entrez <- readr::read_tsv("../data/gencode.v30.metadata.EntrezGene.gz", col_names = FALSE)
head(gencode_entrez[1:10,])
colnames(gencode_entrez) <- c("transcript_id", "entrez_id")
## Joining data that are joined together by transcript_id
entrez_g2t <- left_join(good.g2t, gencode_entrez)
head(entrez_g2t[1:10,])
## Remove transcript_id column
entrez_g2t$transcript_id <- NULL
## Remove . and the numbers following from the ensembl id.
entrez_g2t$gene_id <- gsub("\\..*", "", entrez_g2t$gene_id)
## Removing data that does not have entrez id. 83091 out of 83647 had NA values
table(is.na(entrez_g2t$entrez_id))
entrez_g2t <- entrez_g2t[!is.na(entrez_g2t$entrez_id),]
entrez_g2t <- unique(entrez_g2t[, c("gene_id", "gene_name", "entrez_id")])
## Are there any duplicates in our df? Yes. Only one dupped ensembl id.
table(duplicated(entrez_g2t$gene_id))
## These are pseudogenes... We will remove them all.
dups_gene_id <- entrez_g2t[which(duplicated(entrez_g2t$gene_id)),]
## We're removing the rows that are associated with this pseudogene
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285607")),]
## It's gone
which(entrez_g2t$gene_id == "ENSG00000285607")
## Are there any duplicate gene names? Yes! 14 of them
table(duplicated(entrez_g2t$gene_name))
dups_gene_name <- entrez_g2t[which(duplicated(entrez_g2t$gene_name)),]
entrez_g2t[entrez_g2t$gene_name == "TBCE",]
## ENSG00000285053  is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285053")),]
entrez_g2t[entrez_g2t$gene_name == "PDE11A",]
## ENSG00000284741 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000284741")),]
entrez_g2t[entrez_g2t$gene_name == "ATXN7",]
## ENSG00000285258 is not in the database anywhere
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285258")),]
entrez_g2t[entrez_g2t$gene_name == "MATR3",]
## ENSG00000280987 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000280987")),]
entrez_g2t[entrez_g2t$gene_name == "SOD2",]
## ENSG00000285441 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285441")),]
entrez_g2t[entrez_g2t$gene_name == "ABCF2",]
## ENSG00000285292 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285292")),]
entrez_g2t[entrez_g2t$gene_name == "PINX1",]
## ENSG00000258724 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000258724")),]
entrez_g2t[entrez_g2t$gene_name == "HSPA14",]
## ENSG00000284024 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000284024")),]
entrez_g2t[entrez_g2t$gene_name == "IGF2",]
## ENSG00000284779 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000284779")),]
entrez_g2t[entrez_g2t$gene_name == "ATF7",]
## ENSG00000267281 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000267281")),]
entrez_g2t[entrez_g2t$gene_name == "DIABLO",]
## ENSG00000284934 is a homolog
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000284934")),]
entrez_g2t[entrez_g2t$gene_name == "GGT1",]
## ENSG00000286070 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000286070")),]
entrez_g2t[entrez_g2t$gene_name == "SCO2",]
## ENSG00000130489 no longer exist
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000130489")),]
entrez_g2t[entrez_g2t$gene_name == "TMSB15B",]
## ENSG00000269226 is a readthrough
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000269226")),]

## Are there any dupped entrez id? 
dups_entrez <- entrez_g2t[which(duplicated(entrez_g2t$entrez_id)),]

entrez_g2t[entrez_g2t$entrez_id == "55486",]
## ENSG00000283765 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000283765")),]
entrez_g2t[entrez_g2t$entrez_id == "55300",]
## ENSG00000281028 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000281028")),]
entrez_g2t[entrez_g2t$entrez_id == "653082",]
## ENSG00000286094 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000286094")),]
entrez_g2t[entrez_g2t$entrez_id == "8622",]
## ENSG00000284762 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000284762")),]
entrez_g2t[entrez_g2t$entrez_id == "221468",]
## ENSG00000286105 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000286105")),]
entrez_g2t[entrez_g2t$entrez_id == "1124",]
## ENSG00000285162 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285162")),]
entrez_g2t[entrez_g2t$entrez_id == "102723849",]
## ENSG00000283765 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000273520")),]
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000262461")),]
entrez_g2t[entrez_g2t$entrez_id == "883",]
## ENSG00000286112 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000286112")),]
entrez_g2t[entrez_g2t$entrez_id == "22891",]
## ENSG00000261130 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285551")),]
entrez_g2t[entrez_g2t$entrez_id == "54816",]
## ENSG00000285253 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285253")),]
entrez_g2t[entrez_g2t$entrez_id == "23370",]
## ENSG00000268861 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000268861")),]
entrez_g2t[entrez_g2t$entrez_id == "284391",]
## ENSG00000286098 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000286098")),]
entrez_g2t[entrez_g2t$entrez_id == "440519",]
## ENSG00000283201 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000283201")),]
entrez_g2t[entrez_g2t$entrez_id == "149951",]
## ENSG00000285382 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000285382")),]
entrez_g2t[entrez_g2t$entrez_id == "388820",]
## ENSG00000268861 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000278961")),]
entrez_g2t[entrez_g2t$entrez_id == "8471",]
## ENSG00000286098 not found on ncbi
entrez_g2t <- entrez_g2t[-c(which(entrez_g2t$gene_id == "ENSG00000260548")),]

## manual gene removal complete

load("../data/TCGA.RNA.Rda")

## Put entrez ids as rownames and shortening the TCGA ids since the last three portions of the TCGA barcode is coded to explain the test performed and etc. 
tumor_samples <- gdc_tumor_coding_complete %>% colData(.) %>% 
  as.data.frame(.) %>%
  dplyr::select(project_code, submitted_specimen_id) %>%
  mutate(submitted_specimen_id = substring(text = .$submitted_specimen_id, first = 1, last = 15))

TCGA.RNA <- TCGA.RNA %>% dplyr::select(-gene_symbol) %>%
  remove_rownames(.) %>% column_to_rownames(var = "gene_id") %>% 
  .[rownames(.) %in% entrez_g2t$entrez_id,] %>%
  'colnames<-' (substring(text = names(.), first = 1, last = 15)) %>% ## only need the first 15 character of barcode
  .[, names(.) %in% tumor_samples$submitted_specimen_id] ## only include tumor samples in our analysis

g2e_list <- entrez_g2t[entrez_g2t$entrez_id %in% rownames(TCGA.RNA),]

express_no_express_lookup <- list()
for(i in unique(tumor_samples$project_code)) {
  df <- which(tumor_samples$project_code == i) ## which rows are the spec cancer types
  df <- tumor_samples[df,] ## select only those rows
  gdc <- TCGA.RNA[, names(TCGA.RNA) %in% df$submitted_specimen_id] ## filter the norm_gdc_gene_exp dataset with those tumor samples
  mean_gdc <- as.data.frame(apply(gdc, MARGIN = 1, function(x) {
    mean(x)
  })) ## take the mean for each gene across the tumor samples
  names(mean_gdc) <- i
  express_no_express_lookup[[i]] <- mean_gdc ## store data in list
  rm(df, gdc, mean_gdc)
  gc
}
express_no_express_lookup <- do.call(cbind, express_no_express_lookup) ## cbind list
express_no_express_lookup[is.na(express_no_express_lookup)] <- 0

## 
gdc_ensembl_id <- gdc_tumor_coding_complete %>% rowRanges(.) %>% as.data.frame(.) %>% ## getting the ensembl_ids in our gdc dataset
  rownames(.)
g2e_list <- g2e_list[g2e_list$gene_id %in% gdc_ensembl_id,] ## subsetting the g2e_list to only have ensembl_ids in our gdc dataset

express_no_express_lookup <- express_no_express_lookup[rownames(express_no_express_lookup) %in% g2e_list$entrez_id, ]
express_no_express_lookup[express_no_express_lookup <= 10] <- 0 ## Set threshold to >10 for gene expression
express_no_express_lookup[express_no_express_lookup >10] <- 1

## convert entrez id to ensembl_id
express_no_express_lookup <- express_no_express_lookup %>% .[match(g2e_list$entrez_id, rownames(.)),] %>% ## match the entrez_id express_no_express_lookup with g2e_list
  'rownames<-' (g2e_list$gene_id) ## replace entrez_id with

## Booleanize mutation count for each tumor sample
missense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$missense) 
splice <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$splice)
frameshdel <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshdel)
frameshins <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshins)
utr3 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr3)
utr5 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr5)
nonstop <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonstop)
nonsense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonsense)

gdc_mut_count_patient <- missense + splice + frameshdel + frameshins + utr3 + utr5 + nonstop + nonsense ## add up all the mutation count
rm(missense, splice, frameshdel, frameshins, utr3, utr5, nonstop, nonsense)
gc()
gdc_mut_count_patient <- gdc_mut_count_patient[rownames(gdc_mut_count_patient) %in% rownames(express_no_express_lookup),] ## filter data with genes in the express_no_express_lookup
gdc_mut_count_patient[gdc_mut_count_patient > 0] <- 1 ## booleanize data, anything greater than 1 is equal to 1

gdc_gene_exp_adj <- list()
for(i in unique(tumor_samples$project_code)) {
  df <- which(tumor_samples$project_code == i) ## which rows are my spec cancer types
  df <- tumor_samples[df,] ## select only those rows
  gdc <- gdc_mut_count_patient[, names(gdc_mut_count_patient) %in% rownames(df)] ## filter the norm_gdc_gene_exp dataset with those tumor samples
  exp_data <- as.data.frame(express_no_express_lookup[, i]) ## select correct lookup column for the cancer type
  rownames(exp_data) <- rownames(express_no_express_lookup) ## Insert rownames to exp_data
  gdc <- gdc[match(rownames(exp_data), rownames(gdc)),] ## Match ensembl_id of exp_data with gdc data
  gdc_adj <- as.data.frame(apply(gdc, MARGIN = 2, function(x) {
    exp_data * x
  })) ## Function to multiply lookup to the data for the specific cancer type
  names(gdc_adj) <- names(gdc) ## Rename gdc_adj with tumor samples names
  gdc_gene_exp_adj[[i]] <- gdc_adj ## Store data to list
  rm(df, gdc, exp_data, gdc_adj)
  gc()
}
## Combining list into a dataframe.
gdc_gene_exp_adj <- do.call(cbind, gdc_gene_exp_adj) ## cbind list
names(gdc_gene_exp_adj) <- gsub("^.*\\.", "", names(gdc_gene_exp_adj)) ## tumor sample names have cancer type name in it. Removing it using gsub
rm(gdc_mut_count_patient, express_no_express_lookup, TCGA.RNA, gdc_tumor_coding_complete, gencode_gtf, dups_entrez, dups_gene_id, dups_gene_name, g2t, good.g2t, gene_info, g2e_list, entrez_g2t, gencode_entrez)
gc()
save(gdc_gene_exp_adj, file = "gdc_gene_exp_adj.Rda")
```
<div style="text-align: right"> Genes to pathway transformation </div>
```{r genes-to-pathways-transformation, eval=FALSE}
load("../data/gdc_gene_exp_adj.Rda")
load("../data/gene2_level2_lookup.Rda")
sample_list <- as.character(names(gdc_gene_exp_adj)) ## Create sample list
complete_path <- as.character(names(gene2level2_lookup)) ## Get names of pathways

df_gene <- as.data.frame(t(gdc_gene_exp_adj)) ## Transpose data so samples are rows and genes are columns
df_gene <- df_gene > 0 ## if 1 == TRUE if 0 == FALSE
df_gene <- as.data.frame(df_gene) ## turn matrix back to df

df_gene <- df_gene %>% rownames_to_column(var = "sample") %>% 
  gather(key = "gene", value = "mutation", -sample) %>% ## Wide to long format
  .[.$mutation,] %>% ## Keep TRUE's (which are 1's); Remove FALSE's (0's)
  dplyr::select(-mutation) ## Remove column named mutation

path_lookup <- as.data.frame(t(gene2level2_lookup)) ## Transpose data so paths are rows and genes are columns
path_lookup <- path_lookup > 0 ## if 1 == TRUE if 0 == FALSE
path_lookup <- as.data.frame(path_lookup) ## turn matrix back to df

path_lookup <- path_lookup %>% rownames_to_column(var = "path") %>%
  gather(key = "gene", value = "path_member", -path) %>% ## Wide to long format
  .[.$path_member,] %>% ## Keep TRUE's (which are 1's); Remove FALSE's (0's)
  dplyr::select(-path_member)

df_lookup <- df_gene %>% left_join(., path_lookup, by = "gene") %>% ## Match df_gene and lookup dataframes
  na.omit(.) ## Omit NA's that were generated due to no matches

## getting the genes involved in our analysis
genes_in_pathanalysis <- unique(df_lookup$gene)
write.table(genes_in_pathanalysis, file = "genes_in_pathanalysis.tsv", quote = FALSE, sep = "\t", row.names =  FALSE) ## save data

df_lookup <- as.data.frame(table(df_lookup$sample, df_lookup$path)) ## Generate 1's for the matches

df_lookup <- df_lookup %>% spread(key = "Var1", value = "Freq") %>% ## Long to wide format
  remove_rownames() %>%
  column_to_rownames(var = "Var2") ## path_id to rownames

## Insert missing tumor samples due to no mutations
df_missing <- sample_list[!(sample_list %in% names(df_lookup))] 
mx <- as.data.frame(matrix(data = 0, nrow = nrow(df_lookup), ncol = length(df_missing))) ## Create mx to store data
names(mx) <- df_missing ## Missing tumor sample names in col names
rownames(mx) <- rownames(df_lookup) ## Path names in rownames
df_lookup <- cbind(df_lookup, mx) ## cbind data

## Insert missing pathway due to no mutations
path_missing <- complete_path[!c(complete_path %in% rownames(df_lookup))]
mx <- as.data.frame(matrix(data = 0, nrow = length(path_missing), ncol = ncol(df_lookup))) ## Create mx to store data
names(mx) <- names(df_lookup) ## Insert tumor sample names
rownames(mx) <- path_missing ## Insert missing path names
df_lookup <- rbind(df_lookup, mx) ## cbind data

pathway_order <- rownames(df_lookup) ## need to perserve the order of pathway for the next operation
gdc_reactome_path_analysis <- df_lookup %>% mutate_if(is.numeric, ~1 * (. > 0)) %>% as.data.frame() %>% ## Mutate anything >0 to 1's (Booleanize data)
  'rownames<-' (pathway_order) %>%
  rownames_to_column(var = "PATH_ID") %>%
  left_join(., pathname2level2, by = "PATH_ID") %>%
  dplyr::select(-PATH_ID) %>% 
  remove_rownames() %>% column_to_rownames(var = "PATH_NAME")
rm(df_missing, path_missing, mx, df_gene, sample_list, complete_path, pathway_order, df_lookup)
gc()
save(gdc_reactome_path_analysis, file = "gdc_reactome_path_analysis.Rda")
```
<div style="text-align: right"> Figure 1 data </div>
```{r tumor-clusters, eval = FALSE}
####################
## Figure 1A data ##
####################
load("../data/gdc_reactome_path_analysis.Rda")
load("../data/gdc_tumor_coding_complete.Rdata")

path_daisy <- daisy(t(gdc_reactome_path_analysis), type = list(symm  = c(1:col(gdc_reactome_path_analysis))))
path_daisy <- as.dist(as.matrix(path_daisy))

path_daisy_row <- daisy(gdc_reactome_path_analysis, type = list(symm  = c(1:nrow(gdc_reactome_path_analysis))))
path_daisy_row <- as.dist(as.matrix(path_daisy_row))

pathway_dendro <- pheatmap(gdc_reactome_path_analysis,
         show_rownames = F,
         show_colnames = F,
         cellheight = 1,
         cellwidth = 0.5,
         clustering_distance_cols = path_daisy,
         clustering_distance_rows = path_daisy_row,
         cluster_cols = T,
         cluster_rows = T,
         cutree_cols = 19,
         annotation_legend = F,
         legend_breaks = c(0, 0.25, 0.75, 1),
         legend_labels = c("","Not Mut", "Mut", ""),
         color = c("grey", "firebrick3"),
         clustering_method = "ward.D2",
         legend = T)
dev.off()
pathway_dendro <- pathway_dendro$tree_row$order ## getting pathway row ordering
df <- rownames(gdc_reactome_path_analysis)
pathway_dendro <- df[pathway_dendro]

tumor_samples <- gdc_tumor_coding_complete %>% colData(.) %>%  ## get sample_id data from gdc
  as.data.frame(.) %>%
  dplyr::select(project_code, submitted_specimen_id) %>%  
  mutate(project_code = gsub(pattern = "\\-.*", replacement = "", x = .$project_code))

cancer_type <- tumor_samples %>% .$project_code %>% unique %>% .[order(.)] ## get unique cancer_type

cancer_dendro <- list() ## getting cancer type specific tumor sample clustering info

for(n in cancer_type) {
  cancer_info <- tumor_samples %>% rownames_to_column(var = "sample_id") %>% ## move sample_id to column
    filter(project_code == n) %>% ## filter for specific cancer type
    .[, "sample_id"] ## retain only sample_id
  cancer_path <- gdc_reactome_path_analysis %>% .[, names(.) %in% cancer_info] ## only including filtered sample_id
  cancer_heatmap <- pheatmap(cancer_path[pathway_dendro,], ## using pathway row order 
                          show_rownames = F,
                          show_colnames = F,
                          cellheight = 1,
                          cellwidth = 0.5,
                          cluster_cols = T,
                          cluster_rows = F,
                          annotation_legend = F,
                          color = c("grey", "firebrick3"),
                          clustering_method = "single", ## ordering is single linkage
                          legend = T)
  cancer_order <- cancer_heatmap$tree_col$order ## cancer order
  cancer_path <- names(cancer_path) ## getting the sample_id original order 
  cancer_dendro[[n]] <- cancer_path[cancer_order] ## reorder cancer by heatmap order
  rm(cancer_info, cancer_path)
  gc()
}
dev.off()

for(n in cancer_type[1]) { ## getting the order of clustering by cancer type by alphabetical order
  print(n)
  cancer_order <- as.data.frame(do.call(cbind, cancer_dendro[n])) ## read in first list
  names(cancer_order) <- "sample_id"
  cancer_order$type <- n
  for(p in cancer_type[-1]) {
    print(p)
    df1 <- as.data.frame(do.call(cbind, cancer_dendro[p])) ## read in rest of list in loop
    names(df1) <- "sample_id"
    df1$type <- p
    cancer_order <- rbind(cancer_order, df1) ## combine the two df and repeat the loop until end of cancer_type
    rm(df1)
    gc()
  }
}

cancer_dendro <- cancer_order %>% .[,"sample_id"] %>% as.character(.) ## sample_id ordering

cancer_disrupt_heatmap <- gdc_reactome_path_analysis %>% rownames_to_column(var = "path_name") %>%  ## move rownames to column
  gather(., key = "sample_id", value = "membership", -path_name) ## wide to long format
cancer_disrupt_heatmap$membership <- as.character(cancer_disrupt_heatmap$membership) ## turn boolean values into categories

## set levels based on clustering algorithm above
cancer_disrupt_heatmap$path_name <- factor(cancer_disrupt_heatmap$path_name, levels = c(rev(pathway_dendro)))
cancer_disrupt_heatmap$sample_id <- factor(cancer_disrupt_heatmap$sample_id, levels = c(cancer_dendro))

cancer_disrupt_heatmap <- left_join(cancer_disrupt_heatmap, cancer_order, by = "sample_id") ## add in cancer type information for faceting
save(cancer_disrupt_heatmap, file = "cancer_disrupt_heatmap.Rda")

####################
## Figure 1B data ##
####################
gdc_path <- gdc_reactome_path_analysis %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "sample") %>% ## move path_names to column
  .[match(rownames(tumor_samples), .$sample), ]

gdc_path_mca <- left_join(gdc_path, tumor_samples[, c(1, 3)],
  by = c("sample" = "submitted_sample_id")) %>%
  column_to_rownames(var = "sample")

gdc_path_sample_names <- rownames(gdc_path_mca)
gdc_path_tumor_types <- gdc_path_mca$project_code.x

gdc_path_mca$project_code.x <- NULL

## create a categorical variable from the boolean variable
pathway_list <- make_clean_names(colnames(gdc_path_mca))
pathway_count <- length(pathway_list)
gdc_path_mca <- as.data.frame(lapply(gdc_path_mca, function(x) {
  factor(x, levels = c(0, 1), labels = c("no", "yes"))
}))
colnames(gdc_path_mca) <- pathway_list
gdc_path_mca$project_code <- gdc_path_tumor_types
rownames(gdc_path_mca) <- gdc_path_sample_names

## perform mca
mca_res <- MCA(X = gdc_path_mca[, pathway_list], graph = FALSE)

expl_var <- mca_res$eig[, 2]
expl_var_perm <- matrix(NA, ncol = length(mca_res$eig[, 2]), nrow = 100)

set.seed(4242)
for (k in seq(100)) {
  gdc_perm <- as.data.frame(lapply(gdc_path_mca[, pathway_list], sample))
  MCA_perm <- MCA(X = gdc_perm, graph = FALSE)
  expl_var_perm[k, ] <- MCA_perm$eig[, 2]
}

randomize <- tibble(
  eigenvalue = 1:length(expl_var),
  expl_var = expl_var,
  perm = colMeans(expl_var_perm),
  pvalue = rowSums(t(expl_var_perm) >= expl_var) / 100
)

optimal_mca_dimensions <- randomize %>%
  filter(pvalue < 0.05) %>%
  filter(eigenvalue == max(eigenvalue)) %>%
  pull(eigenvalue)

mca_res <- MCA(
  X = gdc_path_mca[, pathway_list],
  ncp = optimal_mca_dimensions, graph = FALSE
)

save(randomize, mca_res, file = "figure1B_data.rda")

####################
## Figure 1C data ##
####################
rng <- RNGseq(max_n_pt * n_iter, 42)
r <- rng[[3649]]

rngtools::setRNG(r)
gdc.umap <- umap(mca_res$ind$coord, 
                 n_components = 3,
                 n_neighbors = round(sqrt(nrow(mca_res$ind$coord))), # 87
                 n_epochs = 1000,
                 n_threads = 1)
gdc_umap_coors <- as.data.frame(gdc.umap) %>%
  rename(plot_x = V1, plot_y = V2, plot_z = V3)
rownames(gdc_umap_coors) <- rownames(gdc_path_mca)

####################
## Figure 1D data ##
####################
n_iter <- 40
max_n_pt <- 200
score_mca <- matrix(as.numeric(NA), nrow = max_n_pt - 2, ncol = n_iter)

n_tumors <- nrow(mca_res$ind$coord)

cl <- makeCluster(8)
registerDoParallel(cl)

for (i in 3:max_n_pt) {
  print(paste0("MIN POINTS = ", i))
  score_iter <- foreach(1:n_iter,
    .combine = c,
    .export = c("mca_res"),
    .packages = c("dbscan", "uwot"),
    r = rng[(i - 1) * n_iter + 1:n_iter]
  ) %dopar% {
    rngtools::setRNG(r)
    umap_init_mca <- umap(mca_res$ind$coord, 
                          min_dist = 1e-6, 
                          n_components = 3, 
                          n_neighbors = round(sqrt(nrow(mca_res$ind$coord))), 
                          n_epochs = 1000, 
                          n_threads = 1)
    res <- hdbscan(umap_init_mca, minPts = i)
    score_iter_temp <- sum(res$membership_prob < 0.05) / n_tumors
    return(score_iter_temp)
  }
  score_mca[i - 2, ] <- score_iter
}

score <- rowMeans(score_mca)
score_df <- tibble(minPts = 3:max_n_pt, score = score)

sample_df <- score_mca %>%
  as.data.frame() %>%
  as_tibble() %>%
  add_column(minPts = 3:max_n_pt, .before = 1) %>%
  pivot_longer(-minPts) %>%
  group_by(minPts)

optimal_minPts <- score_df %>%
  filter(score == min(score)) %>%
  pull(minPts)

save(score_df, sample_df, optimal_minPts, file = "optimal_minpts_score.rda")

####################
## Figure 1E data ##
####################
rngtools::setRNG(r)

cluster_umap <- umap(mca_res$ind$coord,
  min_dist = 1e-6,
  n_components = 3,
  n_neighbors = round(sqrt(nrow(mca_res$ind$coord))),
  n_epochs = 1000,
  n_threads = 1
)
colnames(cluster_umap) <- c("cluster_x", "cluster_y", "cluster_z")
hdb_cluster_assignment <- hdbscan(cluster_umap, minPts = optimal_minPts)
umap_3d_coors <- gdc_umap_coors %>%
  bind_cols(as.data.frame(cluster_umap)) %>%
  mutate(
    clust = as.factor(hdb_cluster_assignment$cluster),
    member_prob = hdb_cluster_assignment$membership_prob,
    sample_id = rownames(gdc_umap_coors)
  )
## clust 0 is unclassified cluster

####################
## Figure 1F data ##
####################
## reclassifying clust 0 using our classified cluster as training test and unclassified as test set
umap_3d_coors$clust_knn <- umap_3d_coors$clust
umap_3d_coors$clust_knn[umap_3d_coors$clust == 0] <- class::knn(
  train = umap_3d_coors[which(umap_3d_coors$clust != 0), c("cluster_x", "cluster_y", "cluster_z")],
  test = umap_3d_coors[which(umap_3d_coors$clust == 0), c("cluster_x", "cluster_y", "cluster_z")],
  cl = umap_3d_coors$clust[umap_3d_coors$clust != 0], k = 87
)

####################
## Figure 1G data ##
####################
missense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$missense) ## reading in mutation data from SE
splice <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$splice)
frameshdel <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshdel)
frameshins <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshins)
utr3 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr3)
utr5 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr5)
nonstop <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonstop)
nonsense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonsense)

mutation_count <- missense + splice + frameshdel + frameshins + utr3 + utr5 + nonstop + nonsense ## add up all the mutation count
rm(missense, splice, frameshdel, frameshins, utr3, utr5, nonstop, nonsense)
gc()
mutation_count <-mutation_count %>% .[rownames(.) %in% rownames(gene2level2_lookup),] %>% ## filter data with genes in our pathway lookup table
  colSums() %>% as.data.frame() %>% ## sum the mutation count
  'colnames<-' (c("mut_count")) %>% ## rename colname
  rownames_to_column(var = "sample_id") ## move rownames to column

umap_3d_coors <- left_join(umap_3d_coors, mutation_count, by = "sample_id")

####################
## Figure 1H data ##
####################
tumor_samples <- gdc_tumor_coding_complete %>% colData(.) %>% 
  as.data.frame(.) %>%
  dplyr::select(project_code) %>% 
  mutate(project_code = gsub(pattern = "\\-.*", replacement = "", x = .$project_code)) %>%
  rownames_to_column(var = "sample_id")

umap_3d_coors <- left_join(umap_3d_coors, tumor_samples, by = "sample_id")

####################
## Figure 1I data ##
####################
## getting subtype information on our cancer type
subtypes <- PanCancerAtlas_subtypes() 
tumor_subtypes <- substring(text = tumor_samples$submitted_specimen_id, first = 1, last = 12) ## sample_id for subsetting subtypes
subtypes <- subtypes %>% dplyr::select(pan.samplesID, Subtype_Selected) %>% ## select the relevant columns
  mutate(Subtype_Selected = gsub(pattern = ".*\\.", replacement = "", x = .$Subtype_Selected)) %>% ## remove the cancer acronym at the beginning
  mutate(pan.samplesID = substring(text = .$pan.samplesID, first = 1, last = 12)) %>% ## interested in 
  distinct() %>% ## only unique columns
  .[.$pan.samplesID %in% tumor_subtypes, ] %>% ## include sample_ids that are in our analysis
  filter(Subtype_Selected != "Normal") %>%
  .[!.$pan.samplesID %in% .[which(duplicated(.$pan.samplesID)),]$pan.samplesID, ] ## remove dups

#save(subtypes, file = "subtypes.Rda")
load("../data/subtypes.Rda")

umap_3d_coors <- umap_3d_coors %>%
  mutate(pan.samplesID = substring(text = .$sample_id, first = 1, last = 12)) %>%
  left_join(., subtypes, by = "pan.samplesID") %>% ## left_join subtype data to umap
  dplyr::select(-pan.samplesID)

##########################
## Figure 1K and L data ##
##########################
## getting tumor_metastasis, staging
tcga_clinical_tumor_staging <- list()
for(n in cancer_type) {
  print(n)
  cancername <- paste0("TCGA-",n)
  query <- GDCquery(project = cancername,
                  data.category = "Clinical",
                  data.type = "Clinical Supplement",
                  data.format = "BCR Biotab",
                  file.type = "patient")
  GDCdownload(query)
  clinical <- GDCprepare(query)[[1]]
  if("ajcc_pathologic_tumor_stage" %in% colnames(clinical) & "ajcc_tumor_pathologic_pt" %in% colnames(clinical) == T) {
    cat("Yep, ajcc info in there!\n")
    clinical <- clinical[-c(1:2), c("bcr_patient_barcode","ajcc_pathologic_tumor_stage", "ajcc_tumor_pathologic_pt", "ajcc_metastasis_pathologic_pm")]
  }
  if(!"ajcc_pathologic_tumor_stage" %in% colnames(clinical) & "clinical_stage" %in% colnames(clinical) & "ajcc_tumor_pathologic_pt" %in% colnames(clinical)== T) {
    cat("Nope, ajcc stage not there, but pathologic scores is!\n")
    clinical <- clinical[-c(1:2), c("bcr_patient_barcode", "clinical_stage", "ajcc_tumor_pathologic_pt", "ajcc_metastasis_pathologic_pm")]
    names(clinical) <- c("bcr_patient_barcode","ajcc_pathologic_tumor_stage", "ajcc_tumor_pathologic_pt", "ajcc_metastasis_pathologic_pm")
  }
  if(!"ajcc_pathologic_tumor_stage" %in% colnames(clinical) & "clinical_stage" %in% colnames(clinical) & !"ajcc_tumor_pathologic_pt" %in% colnames(clinical) & "pathologic_T" %in% colnames(clinical) == T) {
    cat("Nope, ajcc info not there!\n")
    clinical <- clinical[-c(1:2), c("bcr_patient_barcode", "clinical_stage", "pathologic_T", "pathologic_M")]
    names(clinical) <- c("bcr_patient_barcode","ajcc_pathologic_tumor_stage", "ajcc_tumor_pathologic_pt", "ajcc_metastasis_pathologic_pm")
    }
  if(!"ajcc_pathologic_tumor_stage" %in% colnames(clinical) & !"clinical_stage" %in% colnames(clinical) == T) next
  if("ajcc_pathologic_tumor_stage" %in% colnames(clinical) & !"ajcc_tumor_pathologic_pt" %in% colnames(clinical) == T) next
  tcga_clinical_tumor_staging[[n]] <- clinical
}

tcga_clinical_tumor_staging <- do.call(rbind, tcga_clinical_tumor_staging) ## rbind the list together
tcga_clinical_tumor_staging <- tcga_clinical_tumor_staging[tcga_clinical_tumor_staging$bcr_patient_barcode %in% tumor_subtypes,] ## We have 6633 out of 7607 tumor data
missing_sample <- as.data.frame(matrix(data = "NA", nrow = length(tumor_subtypes[!tumor_subtypes %in% tcga_clinical_tumor_staging$bcr_patient_barcode]), ncol = length(tcga_clinical_tumor_staging) -1 ))

tcga_clinical_tumor_staging <- missing_sample %>% 'colnames<-' (names(tcga_clinical_tumor_staging[-1])) %>% ## colnames based on tcga_clinical_tumor_staging
  mutate(bcr_patient_barcode = tumor_subtypes[!tumor_subtypes %in% tcga_clinical_tumor_staging$bcr_patient_barcode]) %>%
  .[,c(4,1,2,3)] %>% ## reorder columns for rbinding
  rbind(tcga_clinical_tumor_staging, .) ## rbind

save(tcga_clinical_tumor_staging, file = "tcga_clinical_tumor_staging.Rda")

umap_3d_coors <- umap_3d_coors %>%
  mutate(bcr_patient_barcode = substring(text = .$sample_id, first = 1, last = 12)) %>%
  left_join(., tcga_clinical_tumor_staging, by = "bcr_patient_barcode") %>% ## left_join clinical data to umap
  dplyr::select(-bcr_patient_barcode)

M0 <- umap_3d_coors$ajcc_metastasis_pathologic_pm %in% c("cM0 (i+)", "M0")
M1 <- umap_3d_coors$ajcc_metastasis_pathologic_pm %in% c("M1", "M1a", "M1b", "M1c")
NA.value <- umap_3d_coors$ajcc_metastasis_pathologic_pm %in% c("[Discrepancy]", "[Not Applicable]", "[Not Available]", "MX")

umap_3d_coors$ajcc_metastasis_pathologic_pm[M0] <- "M0"
umap_3d_coors$ajcc_metastasis_pathologic_pm[M1] <- "M1"
umap_3d_coors$ajcc_metastasis_pathologic_pm[NA.value] <- "NA"

stage1 <- umap_3d_coors$ajcc_pathologic_tumor_stage %in% c("Stage I", "Stage IA", "Stage IA1", "Stage IA2", "Stage IB", "Stage IB1", "Stage IB2", "Stage IC")
stage2 <- umap_3d_coors$ajcc_pathologic_tumor_stage %in% c("Stage II", "Stage IIA", "Stage IIA1", "Stage IIA2", "Stage IIB", "Stage IIC")
stage3 <- umap_3d_coors$ajcc_pathologic_tumor_stage %in% c("Stage III", "Stage IIIA", "Stage IIIB", "Stage IIIC", "Stage IIIC1", "Stage IIIC2")
stage4 <- umap_3d_coors$ajcc_pathologic_tumor_stage %in% c("Stage IV", "Stage IVA", "Stage IVB", "Stage IVC")
unknown <- umap_3d_coors$ajcc_pathologic_tumor_stage %in% c("Stage X", "[Discrepancy]", "[Not Applicable]", "I/II NOS", "NA", "[Not Available]")

umap_3d_coors$ajcc_pathologic_tumor_stage[stage1] <- "Stage I"
umap_3d_coors$ajcc_pathologic_tumor_stage[stage2] <- "Stage II"
umap_3d_coors$ajcc_pathologic_tumor_stage[stage3] <- "Stage III"
umap_3d_coors$ajcc_pathologic_tumor_stage[stage4] <- "Stage IV"
umap_3d_coors$ajcc_pathologic_tumor_stage[unknown] <- "Not Available"

write.table(x = umap_3d_coors, file = "umap_3d_coors.tsv", quote = FALSE, sep = "\t", row.names = FALSE) ## save the data
```

```{r fig-1a-1l, warning=FALSE, fig.height=11, fig.width=8.5, message = FALSE}
load("../data/cancer_disrupt_heatmap.Rda")
umap_cancertype <- fread("../data/umap_3d_coors.tsv")

kelly.colours <- c("gray95", "gray13", "gold2", "plum4", "darkorange1",
                   "lightskyblue2", "firebrick", "burlywood3", "gray51", "springgreen4",
                   "lightpink2", "deepskyblue4", "lightsalmon2", "mediumpurple4",
                   "orange", "maroon", "yellow3", "brown4", "yellow4", "sienna4", "chocolate", "gray19")

fig1A <- ggplot(data = cancer_disrupt_heatmap, aes(y=sample_id, x=path_name, fill = membership)) + 
  geom_tile(size = 0.6) +
  scale_fill_manual(values = c("grey", "firebrick3")) +
  facet_grid(type ~ ., scales = "free", space = "free", switch = "y") +
scale_y_discrete(position = "left") +
  labs(tag = "A") +
  ylab("") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank(),
        strip.text.y = element_text(size = 12, hjust = 0),
        strip.text.y.left = element_text(angle = 0),
        strip.background = element_blank(),
        legend.position = "none")

## fixing the text cutoff in heatmap
 fig1Ag <- ggplotGrob(fig1A)

 for(i in which(grepl("strip-l", fig1Ag$layout$name))){
   fig1Ag$grobs[[i]]$layout$clip <- "off"
 }

 fig1A <- as.ggplot(fig1Ag)
 rm(fig1Ag)

optimal_mca_dimensions <- randomize %>%
  filter(pvalue < 0.05) %>%
  filter(eigenvalue == max(eigenvalue)) %>%
  pull(eigenvalue)

permut_p <- randomize %>%
  filter(eigenvalue < 50) %>%
  ggplot(aes(x = eigenvalue, y = expl_var / 100)) +
  geom_line(aes(colour = "#998ec3")) +
  geom_line(aes(y = perm / 100, colour = "#f1a340")) +
  scale_color_identity(guide = guide_legend(title = ""), 
                       labels = c("Eigenvalues", "Permuted Eigenvalues")) +
  scale_y_continuous(labels = scales::percent) +
  xlab("MCA Eigenvalues") +
  ylab("explained variance") +
  theme_classic()

pvalue_p <- randomize %>%
  filter(eigenvalue < 50) %>%
  ggplot(aes(x = eigenvalue, y = pvalue)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept = optimal_dimensions, color = "red") +
  ggtitle(label = "", subtitle = paste0("Significant dimensions = ", optimal_dimensions)) +
  xlab("MCA Eigenvalues") +
  ylab("P-value") +
  theme_classic()

fig1B <- permut_p /
  pvalue_p + plot_annotation(title = "Significance of MCA Eigenvalues")

## umap monochrome 2d
fig1C <- ggplot(umap_cancertype, aes(x=x_2d, y=y_2d)) +
  geom_point(colour = "grey", size = 0.25) +
  xlab("umap_1") +
  ylab("umap_2") +
  theme_classic() +
  labs(tag = "C") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank())
## umap hdb_scan
fig1E <- ggplot(subset(umap_cancertype, clust !=0), aes(x = x_2d, y = y_2d, color = as.factor(clust))) +
  geom_point(data = umap_cancertype, aes(x = x_2d, y = y_2d), colour = "grey", size = 0.25) +
  geom_point(size = 0.25) + scale_color_manual(values = kelly.colours[-c(1:2)]) +
  guides(color = F) +
  theme_classic() +
  xlab("umap_1") +
  ylab("umap_2") +
  theme_classic() +
  labs(tag = "E") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank())

## umap knn_clust 
fig1F <- ggplot(subset(umap_cancertype), aes(x = x_2d, y = y_2d, color = as.factor(knn_clust))) +
  geom_point(size = 0.25) + scale_color_manual(values = kelly.colours[-c(1:2)]) +
  guides(color = F) +
  theme_classic() +
  xlab("umap_1") +
  ylab("umap_2") +
  theme_classic() +
  labs(tag = "F") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank())

umap_cancertype$knn_clust <- factor(umap_cancertype$knn_clust, levels = c(1:12))

## mutation burden plot
mut_order <- list() ## store data in list 
for(n in 1:max(as.numeric(as.character(umap_cancertype$knn_clust)))) { ## loop through all the clusters
  message(n)
  mut_order[[n]] <- umap_cancertype %>% filter(knn_clust == n) %>%  ## filter for specific class
    dplyr::select(sample_id, x_2d, y_2d, knn_clust, mut_count) %>% ## select relevant columns
    arrange(mut_count) ## arrange mutation in increasing order
}
mut_order <- mut_order %>% do.call(rbind, .) %>% as.data.frame() ## rbind list

mut_order$sample_id <- factor(mut_order$sample_id, levels = mut_order$sample_id)
mut_order$knn_clust <- factor(mut_order$knn_clust, levels = c(1:12))

## Need the breaks positions for the x-axis
previous_class <- 0 ## starting position 1
position_of_previous_class <- 0 ## starting position 2
xaxis_breaks <- list() ## list for for loop
for(n in 1) { ## loop through 1 to 12 clusters
  message(n)
  class_spec <- round(mut_order %>% filter(knn_clust == n) %>% nrow(.) / 2, digits = 0) ## calculate halfway mark of class 
  xaxis_breaks[[n]] <- class_spec + previous_class + position_of_previous_class ## adding up all variables
  previous_class <- class_spec ## storing class halfway mark
  for(n in 2:max(as.numeric(as.character(mut_order$knn_clust)))) {
    message(n)
    position_of_previous_class <- xaxis_breaks[[n-1]] ## getting the previous class position
    class_spec <- round(mut_order %>% filter(knn_clust == n) %>% nrow(.) / 2, digits = 0) ## calculate halfway mark of new class
    xaxis_breaks[[n]] <- class_spec + previous_class + position_of_previous_class ## adding up all variables
    previous_class <- class_spec ## storing class halfway mark
  }
}
xaxis_breaks <- as.data.frame(do.call(rbind, xaxis_breaks)) ## rbind list
xaxis_breaks <- mut_order[rownames(mut_order) %in% xaxis_breaks$sample_id,] ## find the tumor samples in these positions

total_n <- as.data.frame(table(mut_order$knn_clust)) ## getting total n in cluster

mut_order$log10mut <- log10(mut_order$mut_count + 1) ## mutation count log transformation + pseudo count

mut_median <- mut_order %>% 
  group_by(knn_clust) %>% 
  dplyr::summarize(mut_med = median(log10mut)) ## calculate median for each class

fig1G <- ggplot(mut_order, aes(x = sample_id, y = log10mut, color = knn_clust)) + 
  geom_point(size = 0.25) + 
  ylab("Log10(Mutation Count)") + 
  xlab("") +
  labs(tag = "G") +
  theme_classic() +
  facet_wrap(~ factor(knn_clust, levels = c(1:12)), ncol = 12, scales = "free_x") +
  geom_hline(data= mut_median, aes(yintercept=mut_med)) +
  scale_y_continuous(sec.axis = sec_axis(~.)) + ## dup y-axis
  scale_x_discrete(breaks = c(as.character(mut_order$sample_id[300]), as.character(xaxis_breaks$sample_id)), labels = c("n =" ,total_n$Freq)) +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks.x = element_blank(),
        axis.line.x = element_line(colour = "black", size = 0.5),
        axis.line.x.top = element_line(colour = "black", size = 0.5),
        #strip.background.x = element_blank(),
        strip.placement = "outside",
        strip.text.x = element_text(size = 12),
        legend.position = "none") +
  scale_color_manual(values = kelly.colours[-c(1:2)])

## code to keep only the bottom border line of the facet
fig1G <- ggplotGrob(fig1G)
lg <- linesGrob(x=unit(c(0,1),"npc"), y=unit(c(0,0),"npc"), 
                   gp=gpar(col="black", lwd=4))

for (k in grep("strip-t",fig1G$layout$name)) {
  fig1G$grobs[[k]]$grobs[[1]]$children[[1]] <- lg
}

fig1G <- as.ggplot(fig1G)

## umap by cancer type
fig1H <- ggplot(umap_cancertype[umap_cancertype$project_code %in% c("COAD", "BRCA", "HNSC", "PAAD"),], aes(x = x_2d, y = y_2d, colour = project_code)) +
  geom_point(data = umap_cancertype, aes(x = x_2d, y = y_2d), colour = "grey", size = 0.25) +
  geom_point(size = 0.25) + 
  theme_classic() +
  ylab("umap_2") +
  xlab("umap_1") +
  labs(tag = "H", color = "") +
  scale_color_manual(labels = c("BRCA", "COAD", "HNSC", "PAAD"), 
                     values = c('#fb9a99', '#b15928', '#6a3d9a', 'springgreen4')) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.position = "bottom")

## umap of brca subtype
fig1I <- ggplot(umap_cancertype %>% filter(project_code == "BRCA" & !is.na(Subtype_Selected)), aes(x = x_2d, y = y_2d, colour = Subtype_Selected)) +
  geom_point(data = umap_cancertype, aes(x = x_2d, y = y_2d), colour = "grey", size = 0.25) +
  geom_point(size = 0.25) +
  theme_classic() +
  ylab("umap_2") +
  xlab("umap_1") +
  labs(tag = "I", color = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.position = "bottom")

## umap brca somatic mutations
load("../data/gdc_gene_exp_adj.Rda")
load("../data/gene_data.Rda")
umap_cancertype <- fread("../data/umap_3d_coors.tsv")
brca <- gene_data %>% .[.$external_gene_name %in% c("BRCA1", "BRCA2"),] %>% dplyr::select(ensembl_gene_id, external_gene_name) ## Getting BRCA gene information
brca <- gdc_gene_exp_adj %>% .[rownames(.) %in% brca$ensembl_gene_id, ] %>% ## we only want brca gene
  rownames_to_column(var = "ensembl_gene_id") %>%
  left_join(., brca, by = "ensembl_gene_id") %>% 
  remove_rownames()  %>% column_to_rownames(var = "external_gene_name") %>% ## replace ensembl_id with gene_name
  dplyr::select(-ensembl_gene_id) %>% ## get ride of ensembl_id  col
  t() %>% as.data.frame() %>% ## correct orientation for umap_cancertype df
  rownames_to_column(var = "sample_id") %>% ## getting data ready for leftjoin 
  mutate(BRCA1 = case_when(BRCA1 == "1" ~ "BRCA1",
                           BRCA1 == "0" ~ "NA"),
         BRCA2 = case_when(BRCA2 == "1" ~ "BRCA2",
                           BRCA2 == "0" ~ "NA")) %>%
  mutate(mut_type = gsub('NA,|,NA', '', paste(.$BRCA1, .$BRCA2, sep = ","))) %>% ## combine mutation cause some patients can have both
  dplyr::select(-BRCA1, -BRCA2) %>% ## remove col we don't need
  left_join(., umap_cancertype[,c("sample_id", "x_2d", "y_2d", "project_code")], by = "sample_id") ## leftjoin by sample id

fig1J <- ggplot(brca %>% filter(mut_type != "NA"), aes(x = x_2d, y = y_2d, colour = mut_type)) +
  geom_point(data = brca, aes(x = x_2d, y = y_2d), colour = "grey", size = 0.25) +
  geom_point(size = 0.25) + 
  theme_classic() +
  ylab("umap_2") +
  xlab("umap_1") +
  labs(color = "", tag = "J") +
  scale_color_manual(labels = c("BRCA1", "BRCA1,BRCA2", "BRCA2"), 
                     values = c('#33a02c', '#e31a1c', '#ff7f00')) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.position = "bottom")

## umap stage
fig1K <- ggplot(subset(umap_cancertype, ajcc_pathologic_tumor_stage != "Not Available"), aes(x = x_2d, y = y_2d, colour = ajcc_pathologic_tumor_stage)) +
  geom_point(data = umap_cancertype, aes(x = x_2d, y = y_2d), colour = "grey", size = 0.25) +
  geom_point(size = 0.25) + 
  theme_classic() +
  ylab("umap_2") +
  xlab("umap_1") +
  labs(tag = "K", color = "") +
  scale_color_manual(labels = c("Stage I", "Stage II", "Stage III", "Stage IV"), 
                     values = c('darkgoldenrod3','dodgerblue3','yellow3','tomato2')) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.position = "bottom")

## umap metastasis
fig1L <- ggplot(umap_cancertype, aes(x = x_2d, y = y_2d)) +
  geom_point(data = umap_cancertype, aes(x = x_2d, y = y_2d), colour = "grey", size = 0.25) +
  geom_point(data = subset(umap_cancertype, ajcc_metastasis_pathologic_pm == "M0"), aes(x = x_2d, y = y_2d, colour = ajcc_metastasis_pathologic_pm), size = 0.25) + 
  geom_point(data = subset(umap_cancertype, ajcc_metastasis_pathologic_pm == "M1"), aes(x = x_2d, y = y_2d, colour = ajcc_metastasis_pathologic_pm), size = 0.25) + 
  ylab("umap_2") +
  xlab("umap_1") +
  labs(tag = "L", color = "") +
  scale_color_manual(labels = c("M0", "M1"), 
                     values = c('#a6cee3','#1f78b4')) +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.position = "bottom")

layout <- '
AAAABBCCDD
AAAAEEEFFF
AAAAGGGGGG
AAAAGGGGGG
AAAAHHIIJJ
AAAAKKLL##
'

fig1A + plot_spacer() + fig1C + plot_spacer() + fig1E + fig1F + fig1G + fig1H + fig1I + fig1J + fig1K + fig1L + plot_layout(design = layout)

```

<p style="FONT-SIZE:10px; COLOR:#000000; LINE-HEIGHT:12px;">**Figure 1:
Classification of tumors by pathway disruptions.** a) Each of 377 selected
Reactome pathways (columns) is classified as disrupted if one or more genes is
mutated in the tumor sample (rows). Tumors are grouped by tissue of origin using
standardized abrreviations from the TCGA project. b) MCA analysis was performed
restricted to the 36 informative components. c) The resulting MCA was projected
down to 3 dimensions for the purpose of density based clustering (only the first
2 dimensions are shown). Each dot represents a tumor d) HDBSCAN was used to
perform hierarchical density based clustering tuned to minimize the fraction of
unassigned tumors. e) HDBSCAN clustered tumors were visualized on the UMAP
projection; tumors are colored by cluster membership (see legend). This results
in some tumors remaining unclassified (grey points). f) kNN was used to assign
membership to the remaining tumors. g) Violin plots of sample mutation counts
for tumors in each class. h) Projection of cancer type identity onto UMAP.
Abbrevs: BRCA = Breast cancer, COAD = Colorectal adenocarcinoma, HNSC = Head and
Neck Squamous Cell, PAAD = Pancreatic adenocarcinoma. i) Projection of breast
cancer subtypes onto the UMAP. j) projection of BRCA1/2 somatic mutation onto the
UMAP. k) projection of tumor stage onto the UMAP,
regardless of cancer type. l) Projection of metastatic status onto the UMAP.
Abbrevs: M0 = non-metastatic tumors, M1 = metastatic tumors.</p>