### 2020-07-27 text/survival-methods.Rmd
pull out survival methods into seperate .Rmd

We modeled baseline longevity in the U.S. with these different factors.
<mark>Key assumptions? (Room here for further descriptions and equations)</mark>
Using this base survival model, we constructed a likelihood function with
parameters describing the overall lethality of each cancer type as a modifier
$k_{tissue}$ of the rate parameter $r_{0}$, and cluster specific modifiers of
$k$ and optimized model fit using stan[@stanref]. The $k$ in our model can be
taken as, roughly, the death rate relative to the baseline rates as depicted in
**supplementary figure S5**. Thus, a value of $k=2$ means that a given cancer is
twice as deadly.

With an accurate model for baseline longevity in place, we sought to model
cancer as an event based modifier of long-term survival. Thus any factors or
events such as cancer would modify the baseline risk function as a multiplier.
The resulting coefficients can thus be interpreted as increasing baseline risk
by <mark>x-fold.</mark> Baseline risk was modeled as cumulative survival for
female vs. male caucasians, blacks and asians (hispanics are not clearly marked
in the tumor data; see **supplementary figure S5** for comparisons of raw CDC
data and baseline risk models). Thus we are able to accurately model baseline
risk of death taking these factors into account. The data themselves encompass
all other causes of death for these populations including cancer, thus freeing
our model from concerns about other confounding factors ( _e.g._ Diabetes). For
demonstration purposes we also illustrate some examples of the expected survival
for a population that receives a diagnosis with doubled risk of death at ages
40, 50, 60, 70, and 80 (**supplementary figure S6**).

As an aside, we did not explicitly account for stage at diagnosis in our model
for the following reasons. In the tumor sample data, stage at diagnosis is
confounded with cancer type because some cancers are screened agressively (
_e.g._ colorectal and prostate cancers) while others are diagnosed typically
after they become problematic for the patient's lifestyle ( _e.g._ ovarian and
pancreatic cancers). Secondly, such a model specification would suffer from
added noise because the staging data are not well standardized across cancer
types, have different criteria, and because it is unclear what the relationship
between stage and advancement of disease is (for example some sub-stage 4 tumors
are metastatic). To compound this latter issue, our tumor set is vastly
under-powered given the uneven representation of stage across cancer types.