### 2020-07-27 text/survival-methods.Rmd

In order to understand how the diagnosis of cancer changes the expectation for
longevity of patients, we first began by modeling the longevity of the overall
population of the US based on CDC life tables. Although childhood cancers exist
and are tragic, most cancer patients have diagnosis later in life. We modeled
baseline longevity using a risk rate function:

$$R(t) = R_{20}\exp(a(A-20) + D(t) k a (\frac{t}{365}))$$

Time is split into two components. $A$ going from 0 up to the age of the patient
at diagnosis and then remaining fixed, whereas $t$ starts at 0 upon diagnosis
and represents the days since diagnosis, a field recorded in the dataset. $D(t)$
is an indicator for whether diagnosis has occurred or not (before diagnosis D(t)
= 0, after it equals 1). The model essentially assumes that the risk of dying
per year at age 20 years is $R_{20}$ and that it increases exponentially with a
constant rate $a$ up to diagnosis, and then with a new rate $ka$
thereafter. Though the model does extend back to earlier ages, relatively few
patients were under age 20.

Given this risk per unit time, the probability to die at time $T+dT$ is the
probability to survive to time $T$ which is $(1-P(T))$ and then die in the
remaining interval which is $R(T)dT$

$$dP = (1-P(T)) R(T) dT$$

leading to the differential equation for the cumulative probability of death
$P(T)$

$$\frac{dP}{dT} = R(T)(1-P(T))$$

The solution of this ODE is:

$$P(T) = 1 - \exp(-\int_0^T R(t) dt)$$

And the density of deaths per unit time is the derivative with respect to T:

$$p(T) = R(T) \exp(-\int_0^T R(t)dt)$$

The $R(T)$ function has two important parameters, the $a$ value which represents
the background risk rates, and the $k$ value, which is a function of both cancer
tissue type, and the cancer genomic class.

$$k = k_{tis} k_{cl}$$

The product $k_{tis}k_{cl}$ is obviously symmetric between the two factors. To
disambiguate the meaning of the two, the prior distribution of $k_{cl}$ has peak
probability at 1 and a relatively narrow width, due to the fact that it is a
tissue specific multiplicative modifying factor, this allows the overall
magnitude of k to be primarily determined by the $k_{tis}$ value whose prior is
significantly less constrained so that there is a wider range of dangers across
tissue types.

The $k$ in our model can be taken as, roughly, the aging rate relative to the
baseline rates as depicted in **supplementary figure S5**. Thus, a value of
$k=2$ means that a given cancer causes you to experience the same risk in time
$dt$ as a non-cancer patient would experience at the same age, in time $2dt$.

Using this model, we split the data by Male/Female sex, and by White, vs Black
race as these were the main well-identified categories available to us in the
dataset. Sex is known to be a risk factor for death, with males dying at
slightly higher rates for all ages of interest in our dataset. Black vs White
race is an important category due to differences in access to care and other
socioeconomic and education related factors.

Using the model described above we fit the data using Stan via the rstan package
in R. Posterior inferences are shown in **FIXME WHICH FIGURE?**.

**supplementary figure S5** gives a comparison on log-scale between the post-fit
survival curves and the CDC data it is based on. For demonstration purposes we
also illustrate some examples of the expected survival for a population that
receives a diagnosis with doubled risk of death at ages 40, 50, 60, 70, and 80
(**supplementary figure S6**).

As an aside, we did not explicitly account for stage at diagnosis in our model
for the following reasons. In the tumor sample data, stage at diagnosis is
confounded with cancer type because some cancers are screened aggressively (
_e.g._ colorectal and prostate cancers) while others are diagnosed typically
after they become problematic for the patient's lifestyle ( _e.g._ ovarian and
pancreatic cancers). Secondly, such a model specification would suffer from
added noise because the staging data are not well standardized across cancer
types, have different criteria, and because it is unclear what the relationship
between stage and advancement of disease is (for example some sub-stage 4 tumors
are metastatic). To compound this latter issue, our tumor set is vastly
under-powered given the uneven representation of stage across cancer types.