### 2020-08-08 code/figure2.Rmd
Add Figure 4 forest heatmap with figure legend. Update text to reflect figure 4. Adjust manuscript.rmd to work with docx

```{r oncoprint-top100-plus-panel-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_3d_coors.tsv") ## load in data
load("../data/gdc_reactome_path_analysis.Rda")

######################
## oncoprint top100 ##
######################
onco_color <- c("0" = "grey", "1" = "firebrick3")

top100path <- gdc_reactome_path_analysis %>% rowSums() %>% as.data.frame() %>% rownames_to_column(var = "path_name") %>% arrange(desc(.)) %>% .[1:100,] %>% .[, "path_name"] ## getting the top100 mutated pathways in class

top100path <- gdc_reactome_path_analysis[top100path, ] ## selecting the top 100 paths

top100path[top100path == "1"] <- "MUT" ## categorize boolean values
top100path[top100path == "0"] <- " "
col <- c("MUT" = "firebrick3")
heatmap_legend <- list(title = "", at = c("MUT"),
                       labels = c("Mutation"))
alter_fun = list(
    background = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), 
            gp = gpar(fill = "#CCCCCC", col = NA))
    },
    # bug red
    MUT = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), 
            gp = gpar(fill = col["MUT"], col = NA))
    })
  column_title <- "Top 100 Pathway PanCancer"
  onco_res <- oncoPrint(top100path,
    alter_fun = alter_fun, col = col,
    column_title = column_title, heatmap_legend_param = heatmap_legend,
    remove_empty_columns = F, remove_empty_rows = F, right_annotation = NULL,
    top_annotation = NULL
    )

clust_col_order <- names(top100path[,onco_res@column_order])
path_row_order <- rownames(top100path[onco_res@row_order,])

onco_ggplot <- gdc_reactome_path_analysis %>% .[rownames(.) %in% path_row_order,] %>% ## only include the top 10 paths
  rownames_to_column(var = "path_name") %>% ## path_name to col
  gather(., key = "sample_id", value = "mutation_status", -path_name) %>% ## wide to long format
  mutate(path_name = factor(.$path_name, levels = c(rev(path_row_order))), ## factor to force oncoprint order
         sample_id = factor(.$sample_id, levels = c(clust_col_order)))
save(onco_ggplot, file = "onco_ggplot.Rda")

###################
## barplot panel ##
###################

load("../data/kelly.colours.rda")
load("../data/moderate_enrich_prob.Rda")

enrich_pathways <- moderate_enrich_prob %>%
  group_by(clust) %>%
  filter(difference > 0.3 & signif <= 2500) %>%
  mutate(pathway = make_clean_names(path)) %>%
  rename(clust_knn = "clust") %>%
  select(clust_knn, pathway) %>%
  mutate(clust_knn = as.numeric(clust_knn)) %>%
  mutate(specific = TRUE)

pathway_counts <- tumor_clusters %>%
  left_join(t(gdc_reactome_path_analysis) %>% 
              as.data.frame() %>% 
              clean_names() %>% 
              rownames_to_column("sample_id")) %>% 
  group_by(clust_knn) %>%
  summarise(total = n(), across(intrinsic_pathway_for_apoptosis:signaling_by_the_b_cell_receptor_bcr, sum)) %>%
  mutate(across(intrinsic_pathway_for_apoptosis:signaling_by_the_b_cell_receptor_bcr, ~ .x / total)) %>%
  select(-total) %>%
  pivot_longer(-clust_knn, names_to = "pathway", values_to = "proportion") %>%
  group_by(clust_knn) %>%
  filter(pathway %in% enrich_pathways$pathway) %>%
  left_join(enrich_pathways %>% select(clust_knn, pathway, specific), by = c("clust_knn", "pathway"))

pathway_counts[is.na(pathway_counts$specific), ]$specific <- FALSE

pathway_mat <- pathway_counts %>%
  select(-specific) %>%
  pivot_wider(names_from = pathway, values_from = proportion) %>%
  ungroup() %>%
  select(-clust_knn) %>%
  as.matrix()

pathway_dist <- dist(t(pathway_mat))
pathway_clust <- as.hclust(rev(as.dendrogram(seriate(pathway_dist, method = "OLO_ward")[[1]])))
pathway_counts$pathway <- factor(pathway_counts$pathway, levels = colnames(pathway_mat)[pathway_clust$order])
bp.cluster.colors <- c(kelly.colours[3:12], "gray15")
names(bp.cluster.colors) <- 1:11

pathway_counts <- pathway_counts %>%
  mutate(color = case_when(specific ~ clust_knn, TRUE ~ 11)) %>%
  mutate(color = factor(color))
save(pathway_counts, bp.cluster.colors, file = "barplot_pathways_fig2b.rda")

#####################
## oncoprint panel ##
#####################
clust_col_order <- list()
path_row_order <- list()

for(n in 1:max(tumor_clusters$clust_knn)) {
  class_spec_sample_id <- tumor_clusters %>% .[,c("sample_id", "clust_knn")] %>% filter(clust_knn == n) %>% .[,.$sample_id] ## get class spec sample ids
  top10path <- gdc_reactome_path_analysis %>% .[, names(.) %in% class_spec_sample_id] %>% rowSums() %>% as.data.frame() %>% rownames_to_column(var = "path_name") %>% arrange(desc(.)) %>% .[1:10,] %>% .[, "path_name"] ## getting the top10 mutated pathways in class
  top10path <- gdc_reactome_path_analysis[top10path, class_spec_sample_id]
  top10path[top10path == "1"] <- "MUT" ## categorize boolean values
  top10path[top10path == "0"] <- " "
  col <- c("MUT" = "firebrick3")
heatmap_legend <- list(title = "", at = c("MUT"),
                       labels = c("Mutation"))
alter_fun = list(
    background = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), 
            gp = gpar(fill = "#CCCCCC", col = NA))
    },
    # bug red
    MUT = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), 
            gp = gpar(fill = col["MUT"], col = NA))
    })
  column_title <- "test"
  onco_res <- oncoPrint(top10path,
    alter_fun = alter_fun, col = col,
    column_title = column_title, heatmap_legend_param = heatmap_legend,
    remove_empty_columns = F, remove_empty_rows = F, right_annotation = NULL,
    top_annotation = NULL
    )
  clust_col_order[[n]] <- names(top10path[,onco_res@column_order])
  path_row_order[[n]] <- rownames(top10path[onco_res@row_order,])
}

save(clust_col_order, path_row_order, file = "onco_clust.Rda")
```

```{r figure-2-plot, fig.align= 'center', fig.height= 7, fig.width = 7}
#out.height='90%', out.width='90%'
## top100 pathways pancancer oncoprint
load("../data/onco_ggplot.Rda") ## load data
onco_color <- c("0" = "grey", "1" = "firebrick3") ## color scheme for oncoprint

fig2A <- ggplot(data = onco_ggplot, aes(x=sample_id, y=path_name)) + 
  geom_tile(aes(fill = as.character(mutation_status))) +
  scale_fill_manual(values = onco_color) +
  theme_classic() +
  xlab("Pan-Cancer (n=7,607 tumor samples)") +
  ylab("377 pathways") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        #plot.tag = element_text(size = 8, hjust = 0, vjust = 0),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.position = "none")

load("../data/effect_heatmap.Rda") ## load data
effect_colors = c("1" = "#2d004b", "2" = "#542788", "3" = "#8073ac", "4" = "#b2abd2", "5" = "#d8daeb", "6" = "white", "7" = "#fee0b6", "8" = "#fdb863", "9" = "#e08214","10" = "#b35806","11" ="#7f3b08") ## color scheme

guide_axis_label_trans <- function(label_trans = identity, ...) {
  axis_guide <- guide_axis(...)
  axis_guide$label_trans <- rlang::as_function(label_trans)
  class(axis_guide) <- c("guide_axis_trans", class(axis_guide))
  axis_guide
}

guide_train.guide_axis_trans <- function(x, ...) {
  trained <- NextMethod()
  trained$key$.label <- x$label_trans(trained$key$.label)
  trained
}

fig2B <- ggplot(data = effect_heatmap, aes(y = cluster, x = path_name)) +
  geom_tile(size = 1, aes(fill = effect_size)) +
  scale_fill_fermenter(
    type = "div",
    palette = "PuOr",
    n.breaks = 12,
    limits = c(-1, 1),
    name = "Effect Size",
    labels = c(-1, "", 0.5, "", "", 0, "", "", 0.5, "", 1)
  ) +
  facet_wrap(~cluster, ncol = 1, strip.position = "right", scales = "free_y") +
  guides(fill = guide_colorsteps(title.position = "top", title.hjust = 0.5)) +
  xlab("377 pathways, ward.D2 clustered") +
  ylab("Clusters") +
  scale_x_discrete(position = "top") +
  guides(y.sec = guide_axis_label_trans(~ paste(.x))) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    axis.ticks = element_blank(),
    strip.text.y = element_text(size = 10, hjust = .5),
    strip.text.y.right = element_text(angle = 0),
    strip.background = element_blank(),
    panel.background = element_blank(),
    panel.spacing = unit(0, "lines"),
    legend.position = "bottom",
    legend.key.height = unit(0.3, "lines"),
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 5),
    legend.margin=margin(0,0,0,0),
    legend.box.margin=margin(-10,-10,0,-10)
  )

load("../data/barplot_pathways_fig2b.rda")
fig2C <- pathway_counts %>% ggplot(aes(x = pathway, y = proportion, fill = color)) +
  geom_col() +
  facet_wrap(~clust_knn, ncol = 1, strip.position = "right") +
  ylab("Proportion of mutated samples") +
  xlab("Significant pathways") +
  theme_classic() +
  scale_fill_manual(values = bp.cluster.colors) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    #strip.text.y = element_text(size = 10, hjust = .5),
    #plot.tag = element_text(size = 12, hjust = 0, vjust = 0),
    strip.text.y.right = element_text(angle = 0),
    strip.background = element_blank(),
    legend.position = "none"
  )

layout <- '
AAABBBB
AAABBBB
AAABBBB
AAACCCC
AAACCCC
AAACCCC
AAACCCC
'

fig2A + fig2B + fig2C + plot_layout(design = layout) + plot_annotation(tag_levels = "A")
```

<p style="FONT-SIZE:10px; COLOR:#000000; LINE-HEIGHT:12px;">**Figure 2: Pan-cancer enrichment of pathway disruptions.** a) Waterfall plot; columns represent tumors. Each row represents a pathway. The pathways are ranked such that the pathway with the most disruptions across all 7,607 tumors is the first row _etc._ and tumors are ordered to prioritize the top pathway disruptions successively. The top 100 most mutated pathways are shown. Red signifies pathway disrupted pathways, grey signifies not disrupted. b) Heatmap shows the relative enrichment of each pathway (columns) within each numbered cluster (rows). Effect size is displayed as colors (orange, enrichment; purple, depletion). c) Proportion of mutated samples in the each significant pathway (columns; union set of pathways with effect size >= 0.30 in each cluster) within each numbered cluster (rows).</p>