### 2020-08-13 code/figure1v2.Rmd
Split figure 1 into 4 figures, update figure4.rmd with simon's fix, add page breaks

```{r figure-1-data-generation, eval = FALSE}
###################
## Figure 1 data ##
###################
load("../data/gdc_reactome_path_analysis.Rda")
load("../data/gdc_tumor_coding_complete.Rdata")
load("../data/gene2_level2_lookup.Rda")

path_daisy <- daisy(t(gdc_reactome_path_analysis), type = list(symm  = c(1:col(gdc_reactome_path_analysis))))
path_daisy <- as.dist(as.matrix(path_daisy))

path_daisy_row <- daisy(gdc_reactome_path_analysis, type = list(symm  = c(1:nrow(gdc_reactome_path_analysis))))
path_daisy_row <- as.dist(as.matrix(path_daisy_row))

pathway_dendro <- pheatmap(gdc_reactome_path_analysis,
         show_rownames = F,
         show_colnames = F,
         cellheight = 1,
         cellwidth = 0.5,
         clustering_distance_cols = path_daisy,
         clustering_distance_rows = path_daisy_row,
         cluster_cols = T,
         cluster_rows = T,
         cutree_cols = 19,
         annotation_legend = F,
         legend_breaks = c(0, 0.25, 0.75, 1),
         legend_labels = c("","Not Mut", "Mut", ""),
         color = c("grey", "firebrick3"),
         clustering_method = "ward.D2",
         legend = T)
dev.off()
pathway_dendro <- pathway_dendro$tree_row$order ## getting pathway row ordering
df <- rownames(gdc_reactome_path_analysis)
pathway_dendro <- df[pathway_dendro]

tumor_samples <- gdc_tumor_coding_complete %>% colData(.) %>%  ## get sample_id data from gdc
  as.data.frame(.) %>%
  dplyr::select(project_code, submitted_sample_id, submitted_specimen_id) %>%  
  mutate(project_code = gsub(pattern = "\\-.*", replacement = "", x = .$project_code))

cancer_type <- tumor_samples %>% .$project_code %>% unique %>% .[order(.)] ## get unique cancer_type

cancer_dendro <- list() ## getting cancer type specific tumor sample clustering info

for(n in cancer_type) {
  cancer_info <- tumor_samples %>% rownames_to_column(var = "sample_id") %>% ## move sample_id to column
    filter(project_code == n) %>% ## filter for specific cancer type
    .[, "sample_id"] ## retain only sample_id
  cancer_path <- gdc_reactome_path_analysis %>% .[, names(.) %in% cancer_info] ## only including filtered sample_id
  cancer_heatmap <- pheatmap(cancer_path[pathway_dendro,], ## using pathway row order 
                          show_rownames = F,
                          show_colnames = F,
                          cellheight = 1,
                          cellwidth = 0.5,
                          cluster_cols = T,
                          cluster_rows = F,
                          annotation_legend = F,
                          color = c("grey", "firebrick3"),
                          clustering_method = "single", ## ordering is single linkage
                          legend = T)
  cancer_order <- cancer_heatmap$tree_col$order ## cancer order
  cancer_path <- names(cancer_path) ## getting the sample_id original order 
  cancer_dendro[[n]] <- cancer_path[cancer_order] ## reorder cancer by heatmap order
  rm(cancer_info, cancer_path)
  gc()
}
dev.off()

for(n in cancer_type[1]) { ## getting the order of clustering by cancer type by alphabetical order
  print(n)
  cancer_order <- as.data.frame(do.call(cbind, cancer_dendro[n])) ## read in first list
  names(cancer_order) <- "sample_id"
  cancer_order$type <- n
  for(p in cancer_type[-1]) {
    print(p)
    df1 <- as.data.frame(do.call(cbind, cancer_dendro[p])) ## read in rest of list in loop
    names(df1) <- "sample_id"
    df1$type <- p
    cancer_order <- rbind(cancer_order, df1) ## combine the two df and repeat the loop until end of cancer_type
    rm(df1)
    gc()
  }
}

cancer_dendro <- cancer_order %>% .[,"sample_id"] %>% as.character(.) ## sample_id ordering

cancer_disrupt_heatmap <- gdc_reactome_path_analysis %>% rownames_to_column(var = "path_name") %>%  ## move rownames to column
  gather(., key = "sample_id", value = "membership", -path_name) ## wide to long format
cancer_disrupt_heatmap$membership <- as.character(cancer_disrupt_heatmap$membership) ## turn boolean values into categories

## set levels based on clustering algorithm above
cancer_disrupt_heatmap$path_name <- factor(cancer_disrupt_heatmap$path_name, levels = c(rev(pathway_dendro)))
cancer_disrupt_heatmap$sample_id <- factor(cancer_disrupt_heatmap$sample_id, levels = c(cancer_dendro))

cancer_disrupt_heatmap <- left_join(cancer_disrupt_heatmap, cancer_order, by = "sample_id") ## add in cancer type information for faceting
save(cancer_disrupt_heatmap, file = "cancer_disrupt_heatmap.Rda") 
```

```{r, fig.width = 9, fig.height= 4}
load("../data/cancer_disrupt_heatmap.Rda")
umap_cancertype <- fread("../data/umap_3d_coors.tsv")

kelly.colours <- c("gray95", "gray13", "gold2", "plum4", "darkorange1",
                   "lightskyblue2", "firebrick", "burlywood3", "gray51", "springgreen4",
                   "lightpink2", "deepskyblue4", "lightsalmon2", "mediumpurple4",
                   "orange", "maroon", "yellow3", "brown4", "yellow4", "sienna4", "chocolate", "gray19")

fig1A <- ggplot(data = cancer_disrupt_heatmap, aes(x=sample_id, y=path_name, fill = membership)) + 
  geom_tile(size = 0.6) +
  scale_fill_manual(values = c("grey", "firebrick3")) +
  facet_grid(.~type, scales = "free", space = "free", switch = "x") +
scale_y_discrete(position = "left") +
  xlab("") +
  ylab("377 Reactome Pathways") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank(),
        strip.text.y = element_text(size = 12, hjust = 0),
        strip.text.y.left = element_text(angle = 0),
        strip.text.x = element_text(angle = 90),
        strip.background = element_blank(),
        legend.position = "none")

## fixing the text cutoff in heatmap
 fig1Ag <- ggplotGrob(fig1A)

 for(i in which(grepl("strip-b", fig1Ag$layout$name))){
   fig1Ag$grobs[[i]]$layout$clip <- "off"
 }

 fig1A <- as.ggplot(fig1Ag)
 rm(fig1Ag)

opar <- par(mfrow = c(3,3), mai = c(0.1, 0.1, 0.1, 0.1))

fig1A
```