### 2020-08-13 code/figure2v2.Rmd

```{r figure-2-data-generation, eval = FALSE}
####################
## Figure 1B data ##
####################
gdc_path <- gdc_reactome_path_analysis %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "sample") %>% ## move path_names to column
  .[match(tumor_samples$submitted_sample_id, .$sample), ]

gdc_path_mca <- left_join(gdc_path, tumor_samples[, c("project_code", "submitted_sample_id")],
  by = c("sample" = "submitted_sample_id")) %>%
  column_to_rownames(var = "sample")

gdc_path_sample_names <- rownames(gdc_path_mca) ## Getting sample id order
gdc_path_tumor_types <- gdc_path_mca$project_code ## Getting cancer type order

gdc_path_mca$project_code <- NULL ## Remove project_code column from df

## create a categorical variable from the boolean variable
pathway_list <- make_clean_names(colnames(gdc_path_mca))
pathway_count <- length(pathway_list)
gdc_path_mca <- as.data.frame(lapply(gdc_path_mca, function(x) {
  factor(x, levels = c(0, 1), labels = c("no", "yes"))
}))
colnames(gdc_path_mca) <- pathway_list
gdc_path_mca$project_code <- gdc_path_tumor_types
rownames(gdc_path_mca) <- gdc_path_sample_names

## perform mca
mca_res <- MCA(X = gdc_path_mca[, pathway_list], graph = FALSE)

expl_var <- mca_res$eig[, 2]
expl_var_perm <- matrix(NA, ncol = length(mca_res$eig[, 2]), nrow = 100)

set.seed(4242)
for (k in seq(100)) {
  gdc_perm <- as.data.frame(lapply(gdc_path_mca[, pathway_list], sample))
  MCA_perm <- MCA(X = gdc_perm, graph = FALSE)
  expl_var_perm[k, ] <- MCA_perm$eig[, 2]
}

randomize <- tibble(
  eigenvalue = 1:length(expl_var),
  expl_var = expl_var,
  perm = colMeans(expl_var_perm),
  pvalue = rowSums(t(expl_var_perm) >= expl_var) / 100
)

optimal_mca_dimensions <- randomize %>%
  filter(pvalue < 0.05) %>%
  filter(eigenvalue == max(eigenvalue)) %>%
  pull(eigenvalue)

mca_res <- MCA(
  X = gdc_path_mca[, pathway_list],
  ncp = optimal_mca_dimensions, graph = FALSE
)

save(randomize, mca_res, file = "figure1B_data.rda")

####################
## Figure 1C data ##
####################
n_iter <- 40
max_n_pt <- 200
rng <- RNGseq(max_n_pt * n_iter, 42)
r <- rng[[3649]]

rngtools::setRNG(r)
gdc.umap <- umap(mca_res$ind$coord, 
                 n_components = 3,
                 n_neighbors = round(sqrt(nrow(mca_res$ind$coord))), # 87
                 n_epochs = 1000,
                 n_threads = 1)
gdc_umap_coors <- as.data.frame(gdc.umap) %>%
  rename(plot_x = V1, plot_y = V2, plot_z = V3)
rownames(gdc_umap_coors) <- rownames(gdc_path_mca)

####################
## Figure 1D data ##
####################
n_iter <- 40
max_n_pt <- 200
score_mca <- matrix(as.numeric(NA), nrow = max_n_pt - 2, ncol = n_iter)

n_tumors <- nrow(mca_res$ind$coord)

cl <- makeCluster(8)
registerDoParallel(cl)

for (i in 3:max_n_pt) {
  print(paste0("MIN POINTS = ", i))
  score_iter <- foreach(1:n_iter,
    .combine = c,
    .export = c("mca_res"),
    .packages = c("dbscan", "uwot"),
    r = rng[(i - 1) * n_iter + 1:n_iter]
  ) %dopar% {
    rngtools::setRNG(r)
    umap_init_mca <- umap(mca_res$ind$coord, 
                          min_dist = 1e-6, 
                          n_components = 3, 
                          n_neighbors = round(sqrt(nrow(mca_res$ind$coord))), 
                          n_epochs = 1000, 
                          n_threads = 1)
    res <- hdbscan(umap_init_mca, minPts = i)
    score_iter_temp <- sum(res$membership_prob < 0.05) / n_tumors
    return(score_iter_temp)
  }
  score_mca[i - 2, ] <- score_iter
}

score <- rowMeans(score_mca)
score_df <- tibble(minPts = 3:max_n_pt, score = score)

sample_df <- score_mca %>%
  as.data.frame() %>%
  as_tibble() %>%
  add_column(minPts = 3:max_n_pt, .before = 1) %>%
  pivot_longer(-minPts) %>%
  group_by(minPts)

optimal_minPts <- score_df %>%
  filter(score == min(score)) %>%
  pull(minPts)

save(score_df, sample_df, optimal_minPts, file = "optimal_minpts_score.rda")

####################
## Figure 1E data ##
####################
rngtools::setRNG(r)

cluster_umap <- umap(mca_res$ind$coord,
  min_dist = 1e-6,
  n_components = 3,
  n_neighbors = round(sqrt(nrow(mca_res$ind$coord))),
  n_epochs = 1000,
  n_threads = 1
)
colnames(cluster_umap) <- c("cluster_x", "cluster_y", "cluster_z")
hdb_cluster_assignment <- hdbscan(cluster_umap, minPts = optimal_minPts)
umap_3d_coors <- gdc_umap_coors %>%
  bind_cols(as.data.frame(cluster_umap)) %>%
  mutate(
    n_pathways_mut = rowSums(gdc_path[, -1]),
    clust = as.factor(hdb_cluster_assignment$cluster),
    member_prob = hdb_cluster_assignment$membership_prob,
    sample_id = rownames(gdc_umap_coors)
  )
## clust 0 is unclassified cluster

####################
## Figure 1F data ##
####################
## reclassifying clust 0 using our classified cluster as training test and unclassified as test set
umap_3d_coors$clust_knn <- umap_3d_coors$clust
umap_3d_coors$clust_knn[umap_3d_coors$clust == 0] <- class::knn(
  train = umap_3d_coors[which(umap_3d_coors$clust != 0), c("cluster_x", "cluster_y", "cluster_z")],
  test = umap_3d_coors[which(umap_3d_coors$clust == 0), c("cluster_x", "cluster_y", "cluster_z")],
  cl = umap_3d_coors$clust[umap_3d_coors$clust != 0], k = 87
)
```

```{r figure-2-plot, fig.width = 9}
load("../data/figure1B_data.rda")
optimal_mca_dimensions <- randomize %>%
  filter(pvalue < 0.05) %>%
  filter(eigenvalue == max(eigenvalue)) %>%
  pull(eigenvalue)
umap_cancertype <- data.table::fread("../data/umap_3d_coors.tsv")
kelly.colours <- c("gray95", "gray13", "gold2", "plum4", "darkorange1",
                   "lightskyblue2", "firebrick", "burlywood3", "gray51", "springgreen4",
                   "lightpink2", "deepskyblue4", "lightsalmon2", "mediumpurple4",
                   "orange", "maroon", "yellow3", "brown4", "yellow4", "sienna4", "chocolate", "gray19")
ditto_color <- c(
  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
  "#0072B2", "#D55E00", "#CC79A7", "#666666",
  "#AD7700", "#1C91D4", "#007756", "#D5C711",
  "#005685", "#A04700", "#B14380", "#4D4D4D",
  "#FFBE2D", "#80C7EF", "#00F6B3", "#F4EB71",
  "#06A5FF", "#FF8320", "#D99BBD", "#8C8C8C",
  "#FFCB57", "#9AD2F2", "#2CFFC6", "#F6EF8E",
  "#38B7FF", "#FF9B4D", "#E0AFCA", "#A3A3A3",
  "#8A5F00", "#1674A9", "#005F45", "#AA9F0D",
  "#00446B", "#803800", "#8D3666", "#3D3D3D")
cluster.colours <- c("grey", kelly.colours[c(3:12)])
names(cluster.colours) <- c(0:10)
load("../data/optimal_minpts_score.rda")

opar <- par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(4,6,2,0) + 0.1)
with(randomize %>% filter(eigenvalue < 50), plot(x = eigenvalue, y = expl_var / 100, 
                                                 type = "l", yaxt = "n", bty = "l",
                                                 col = "#998ec3", 
                                                 xlab = "", ylab = ""))
with(randomize %>% filter(eigenvalue < 50), lines(x = eigenvalue, y = perm / 100,
                                                  col = "#f1a340"))
yticks_val <- scales::breaks_pretty(n=5)(randomize %>% filter(eigenvalue < 50) %>% pull(expl_var) / 100)
axis(2, at=yticks_val, lab=scales::percent(yticks_val), las = 1)
title(ylab = "explained variance", line = 3.5)
title(xlab = "MCA Eigenvalues", line = 2)
legend("topright", legend=c("Eigenvalues", "Permuted Eigenvalues"),
       col=c("#998ec3", "#f1a340"), lty=1, cex=0.8)
### D
par(mar = c(2,2,0,0), bg = NA)
with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "white", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
### E
par(mar = c(4,6,2,0) + 0.1)
with(sample_df, plot(x = minPts, y = value, 
                     col = "grey", pch = 20, yaxt = "n",
                     xlab = "", ylab = "", bty = "l",
                     las = 1, cex = 0.3))
with(score_df, lines(x = minPts, y = score,
                     col = "black", las = 1))
abline(v = optimal_minPts, col = "red")
title(ylab = "score", line = 3.5)
title(xlab = "Minimum Cluster Size", line = 2)
axis(2, at=c(0.25, 0.35, 0.45, 0.55), las = 1)

### C
with(randomize %>% filter(eigenvalue < 50), plot(x = eigenvalue, y = pvalue, 
                                                 type = "l", bty = "l",
                                                 col = "black", las = 1,
                                                 xlab = "", ylab = ""))
with(randomize %>% filter(eigenvalue < 50), points(x = eigenvalue, y = pvalue, pch = 20,
                                                   col = "black"))
abline(v = optimal_mca_dimensions, col = "red")
title(main = paste0("Significant dimensions = ", optimal_mca_dimensions), adj = 0, font = 3, cex.main = .8)
title(ylab = "P-value", line = 3.5)
title(xlab = "MCA Eigenvalues", line = 2)
### F
par(mar = c(2,2,0,0), bg = NA)
with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust)], pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
### G
with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)

plot.new()
```