### 2020-08-14 text/discussion.Rmd

### Classification of tumors independent of tissue-of-origin

One of the biggest hurdles in cancer research is the sparsity of data; 
~20,000 protein-coding genes is comparable with the number of tumor samples,
even with multiple mutations per sample. And that is not accounting for the fact
that ~30,000 non-coding genes and hundreds of thousands of non-coding regulatory
elements exist with oncogenic or tumor suppressor activity [@deesMuSiC2012;
@tamboreroOncodriveCLUST2013;
@lawrenceDiscoverySaturationAnalysis2014;
@kumarStatisticallyIdentifyingTumor2015; @jiangWITER2019;
@zhaoModelingPositive2019; @noncoding-mutations-Zhang]. Hence, we sought to
simplify the problem by employing a "knowledge-base driven analysis"
[@pathways-Atul-Butte], investigating cancer as a disease of basic cellular and
biochemical pathways, which we accomplished by translating gene-level mutations
into pathway level disruptions. Our approach differs from previously described
methods [@PATHNET] in that we chose to focus on the pathway as the unit of
disruption instead of the gene, employing the conceit that individual mutations
may be sufficient to disrupt or misdirect pathway activity (as is often the case
with driver mutations). Rather than evaluating the burden of mutations on each
pathway within or across tumors, which can be biased by the number of genes or
the coverage of pathway-linked genes in the genome, we scored each pathway in
binary fashion, allowing one mutation or multiple mutations to count equally for
each sample.

To our knowledge, this approach has not previously been attempted in the
literature in spite of its relative simplicity. We limited our analysis to
mutations with likely deleterious effects, including missense, nonsense,
frame-shift, stop-gain, untranslated region and splice site mutations in genes that are actively
expressed in each cancer type, thus avoiding bias from transcription coupled
repair. Our method of filtration differs from the "rank-and-cut" method of
[@PCAWG-MAIN] but represents a reasonable attempt to account for the same
biases. We also filtered genes on the basis of pathway membership in biochemical
pathways, excluding curated gene sets related to diseases, syndromes, or classes
of proteins with shared catalytic activity or conserved domains ( _e.g._ transcription factors) which are potentially problematic
[@pathways-Atul-Butte]. We chose this approach to limit redundancy, which is
already high in the remaining pathways for some driver genes, and to exclude
biologically unrelated collections of genes. We do not account for non-coding
mutations, amplification events or structural variants, though we hope to
include them in subsequent studies (see below). Our pathway-centric analysis
also allows us to circumvent the problem that some cancer-promoting mutations
likely occur at low enough frequencies as to be indistinguishable from
background. Previous attempts to capture these associations using
Bayesian networks have yielded important new insights into this problem and new
genes [@hornNetSigNetworkbasedDiscovery2018] that would not have been possible
with traditional analysis, supporting the case that pathway-centric approaches
have the potential to increase statistical power.

The hypothesis that cancer is the result of dysfunction from
a limited number of basic cellular processes common to all living human cells
was first introduced and later expanded on in a pair of classic essays by
Hanahan and Weinberg [@weinberg2000; @hallmarks_II]. In this view, all tumors
regardless of tissue of origin acquire six essential phenotypes; sustaining
proliferative signaling, evading growth suppressors, replicative immortality,
evading apoptosis, supportive angiogenesis, tissue invasion and metastasis
[@weinberg2000]. Advancements in the field over the ensuing decade led to the
inclusion of new tumor promoting processes, namely deregulation of cellular
energy cycle, tumor promoting inflammation, evasion of the immune system, genome
instability, and on top of these acquired traits a new appreciation for the
concept of the tumor microenvironment and its contribution to the evolution and
promotion of metastatic disease [@hallmarks_II].

An alternative hypothesis is that every tumor belongs to one of a large number
of syndromes which are unique to each tissue-of-origin, that share some
mechanisms and treatment strategies. Recent publication of TCGA consortium
papers present a view largely, and surprisingly, consistent with this hypothesis
[@hoadleyMultiplatformAnalysis2014; @hoadleyCellofOriginPatternsDominate2018].
Perhaps owing to the intractable complexity of genomics, proteomics, and
patient metadata in all its forms, the inescapable conclusion thus far is that
tissue-of-origin remains the most important driver of tumor characteristics at
every scale and by every measure. Our observations contrast with this view, and
instead support an interpretation of publicly available data in which all tumors
manifest one of a limited number of metastable phenotypes resulting from
disruptions in basic pathways represented by our classes (**figure 1**).

We attempted to account for our clusters in terms of more trivial explanations.
For example, it could be that the classes we identified are consistent with a
signature of overall disease progression, such that successive, adjacent classes
on the UMAP projection exhibit increasing specificity. We identified no such
trend in the number of mutations (see **figure 4**), the relative staging
(**figure 2D**) or metastasis (**figure 2E**); and each class instead was
associated with unique combinations of pathways (as in **figures 5B, C**).

We find that most tumors segregate into one of ten biologically distinct
clusters, regardless of the cancer type (**figure 2F, 3A** and **supplementary
figure S1**). Some cancer types ( _e.g._ Pancreatic:PAAD in the RAS
mutant cluster 8, and Kidney Chromophobe:KICH in cluster 1, see **supplementary
figure S1**) are unevenly distributed, though we could not identify any cancers
that were exclusive to a single tumor class. Only cancer types with the fewest
samples were found to be absent from one or more classes at all. Somewhat
surprisingly to us, this finding extends to histological subtypes of breast,
head and neck cancers, leukemias, _etc._ (see **Figure 3B** and **Supplemental
Figure S2**). This result implies that many, if not most, histological subtypes
could reflect differences in cell-of-origin, rather than fundamental differences
in proximal cause. The four major subtypes of breast cancer correspond to histological
and molecular expression profiles that define them and how they respond to experimental
stimuli [@breast-subtypes2]. It has been hypothesized that differences in the
molecular regulators of development in the precursor cell types present in the
breast epithelium drive the various histological phenotypes [@breast-subtypes;
@breast-subtypes3]. Consistent with this view we found that breast tumor samples
of the Luminal A subtype were heavily biased toward membership in clusters 1 and
7, and basal tumor samples were biased toward cluster 6 (**figure 3B**).
However, both subtypes also contained samples in every other class (without
exception), and Luminal B and Her2 positive samples are distributed across
clusters. We interpret these data to mean that inherited cell-of-origin
molecular profiles may predispose certain precursor cell types within the breast
epithelium to forming tumors of one class or another but are not determinative;
this view is compatible with the previously stated hypothesis but opens the way
for a more granular view of individual tumors. It would be surprising if we did
not observe bias for some cancers and subtypes amongst our classes, since some
treatment regimens have greater efficacy for patients
of a given cancer or histological subtype [@breast-subtypes2]. Nonetheless, the
basis for some tumors being treatment-refractory in spite of receiving the
standard of clinical care for diagnostic markers remains elusive. Doubtless some
of this is due to bad luck, as tumors can metastasize and remain dormant years
before they are detected at distal sites, or resistant clones may have already
arisen at undetectable levels [@hallmarks_II], but our analysis suggests the
possibility of identifying more informative molecular, histological or cellular
subtypes that could form a basis for future separation of patients into
different precision treatment regimens.

### Tissue specific manifestation of pathway-centric disruptions

Our results illustrate how unique combinations of mutations in pan-cancer driver genes with tissue-specific pathway disruptions result in common categories when viewed at the level of the pathway knowledge-base. For example, we find that for "snRNP assembly", one of the unique driver pathways of cluster 5, the top gene in breast cancer is represented by TPR (mutated in 26% of samples). By contrast the top gene for head and neck and lower grade gliomas is RANBP (20% and 13% of samples, respectively). For some other clusters, _e.g._ 7, 8 & 9, the top genes are known drivers such as TP53, PIK3CA and KRAS. None of these genes is 100% mutated in any cluster however, and therefore it is the combinations with secondary mutations, many of which are tissue-specific, that round out the complement of pathways in all clusters. Some top genes are found in most of the clusters (supplemental figure S5), in spite of the fact that they contribute to many cluster-specific pathways. This can only be explained as a result of other less common driver genes complementing the pathway disruption of the unique combination of driver genes that are disrupted in each tumor, and we speculate that many of these are relatively tissue-specific in terms of their sensitivity to mutation. Consistent with our observations, @moonlight showed how tissue- and cancer-specific driver genes and druggable targets of pathway moduli can be discovered by machine learning on integrated multi-omics datasets. Marrying these approaches in future work will greatly enhance our understanding of tumor biology.

### Incompleteness of the pathway disruption data

Finally, we must remark on the limitations of our work exemplified in two of the
classes, 1 and 10, for which we did not find many distinctive associations with
pathways (**figure 5B, C**). Class 1 had a relatively low proportion of mutated
pathways across the board, although it is broadly enriched in many of the same
tumor-promoting pathways common to the other groups. This likely does not
reflect a true difference in stages of tumor evolution, as our data show clearly
that this class is as likely to contain stage IV metastatic tumors as it is to
contain those of stage I (**figure 3D, E**). This class likely represents a
group of tumors with aberrations in methylation, copy number alterations,
structural variants, or combinations thereof which we excluded from our
analysis. Consistent with this, kidney chromophobe and thyroid cancers have high
proportions of structural variations _vs._ other variant types [@PCAWG-MAIN] and are heavily
skewed to cluster 1 membership (**Supplementary figure S1**). Likewise, Class 10
represents a group of hyper-mutated tumors that harbor so many mutations that
virtually no pathway is unaffected. It seems likely that a significant fraction
of the “mutant” samples for each pathway are burdened with excess passenger mutations. This could be
addressed with application of ever-increasingly sophisticated filtering likely passenger mutations [ _e.g._ @spliceAI;
@deep-learning-clinical-variants]. In the future, we hope to incorporate these
other data into a comprehensive pathway-centric analysis as we have done here
for point mutations and indels.

### Estimates of survival reveal pathway-dependent differences

By modeling CDC longevity data as a baseline risk function we showed that each
of our pathway disruption classes exhibit cancer-type specific effects on
expected survival (**figure 7**). Contrary to what one might expect at the
outset, we found that models assuming different survival based on class and
cancer type fit better than models in which they are independent
factors. However, considering that within each cancer type there are different
clinical standards of care, and even within classes of drugs the preferred
treatment can very between cancers, it makes sense that we observe tissue-specific class effects. Contrast the situation with ovarian _vs._ breast cancer, which
are both hormonally driven cancers, for example. Ovarian cancer has but one main
treatment axis, platinum, whereas breast cancer patients have a variety of
treatment regimens based on molecular subtype and other factors. It is easy to
see how standard of care for one cancer and class could produce different
outcomes than the same class of a different cancer type. Unfortunately, given
the diversity drug classes and treatments (including non-chemical treatments
such as surgery and palliative care), we lack sufficient power 
to explore these variables in the TCGA data. It is our hope that 
future studies will help to distinguish between treatment-specific
effects on survival given different pathway disruption classes.

### Implications for the evolution of cancer

Our findings imply that there are separate processes in the etiology of cancer
that can be broadly thought of as general cancer promoting, class-specific
mutations and metastasis. General cancer promoting processes must include
factors that relate to genome stability and immortality, as "enabling
characteristics" of the cancer phenotype [@hallmarks_II]. Such pathways are
disrupted in all or most of the defined classes (**supplementary figure S1**)
and are frequently the result of aberrations involving common driver genes such as
BRCA1/2, MMR genes, mitotic checkpoints, Cohesion complexes, _etc._
(**supplementary figure S3** and **supplemental file 1**). Class-specific
evolution must involve the acquisition of disruptions to pathways that may
individually be harmful ( _e.g._ highly proliferative cells are more likely to
senesce) but together produce more specialized cancer phenotype and increased
fitness (as in **figure 6** and **tables 1** and **ST2**). Importantly, our
observations do not imply the order in which these mutations should accumulate.
This could be addressed in a future study by integrating our analysis with
evolutionary analysis of clonality, drawing inferences from variant allele
frequencies as in @PCAWG-EVO. However, since many of the genes in the
non-cluster-specific pathways involve the known driver genes, it is reasonable to
hypothesize that these mutations promote or enable the subsequent acquisition of
class-specific defects via random mutation and natural selection, thus producing
the classes we observed. In support of this, the pan-cancer analysis of whole
genomes consortium (PCAWG) found that oncogenic driver
mutations are highly enriched in early arising clones, whereas later arising
clones have much greater diversity in driver mutations [@PCAWG-EVO]. Moreover,
driver genes that are known to be responsible for discrete mutation signatures
such as APOBEC, BRCA1 and BRCA2 produce mutational hotspots reflecting the
varying selective pressures in different tissues [@PCAWG-MAIN].

One of the drawbacks of bulk tumor whole genome sequencing data is the problem
of tumor heterogeneity. Consortium samples are likely to contain contamination
from support tissue, stroma, inflammatory cells, immune cells of the innate and
adaptive immune systems, all potentially harboring cancer supporting mutations
[@bhowmick_review]. We think it will be instructive to explore these ideas in
the context of single cell experiments.

### On metastasis as a convergence of phenotypes

We report that pathway enrichment in metastatic tumors across all classes
(pooled) yielded generally lower effect sizes and larger p-values than the
cluster specific analysis, suggesting that signal is diluted when the samples
are analyzed together, and supporting the view that, within the parameters of our analysis, metastasis has
class-specific requirements. Since the number of metastatic samples is
relatively low this part of our analysis is likely underpowered and subject to
expanded analysis with larger cohorts. Nonetheless, the class-specific
metastatic pathway enrichment we observe (in **table 1**) could be a result of
the pathways directly promoting metastatic phenotypes in these classes, or
because they acquire a stronger or more complete oncogenic phenotype, as we
don’t know the order in which they were acquired. The fact that most metastatic
enrichment is cluster-specific and has a tendency to overlap with
cluster-specific pathways from non-metastatic tumors of neighboring clusters
favors the latter interpretation, and suggests a mechanism of complementarity,
in which newly acquired mutations result in similar classes converging on
one or more deadly phenotypes that capture most of the critical features of end-stage
cancer. One explanation for this could be that tumors must accumulate mutations
in a certain order or face elimination due to any of the selective bottlenecks
they face (senescence, the immune system, _etc._) [@hallmarks_II]. Some clusters
were characterized by cluster-specific pathways that seem superficially
unrelated; for example cluster 2 has mutations in the proteasome subunits,
cluster 3 has mutations impacting protein translation, whereas cluster 5 has
defects in RNA metabolism and clusters 7, 8 and 9 have defects in the more
traditional extra-cellular signaling pathways (as exemplified in **table S1**).
It may be that the primary effects of mutations in protein and RNA metabolism in
clusters 3 and 5 point to currently underappreciated mechanisms impacting hallmark
pathways. Clusters 3 and 5 are also marked by a decidedly different set of
pathways in metastasis. Clusters 3 and 5 notably have (in addition to signaling
pathways) mutations in glucagon and glycogen metabolism, respectively, in their
metastases (**table 1**), consistent with energy metabolism as one of the new
hallmarks [@hallmarks_II], but also distinguishing these classes from the others
in the order or nature of hallmark acquisition. Thus, even at relatively low
power, our analysis of metastasis further supports the pathway-centric approach
as a fruitful way to uncover real differences in underlying biology. As a major
caveat to this, we acknowledge that many therapies _do_ target highly-specific
driver genes, markers, and signaling pathways ( _e.g._ TP53, EGFR or HER2), but
understanding the broader context of the genetic background and pathway
vulnerability of tumors containing such markers may aid in creating smarter
combination therapies. We submit that when we discover the requirements of each
cluster with respect to pathway disruptions and metastasis we may be able to
target them therapeutically and prevent further adaptation.