### 2020-08-16 text/introduction.Rmd
edit full document

Recent advances in technology have made the study of cancer possible at
unprecedented levels of detail and scale. Array and next-generation sequencing
technologies have made it possible to assay germline and somatic mutations,
expression profiles, DNA methylation, and copy number variations. To take
advantage of these tools several large consortia, including the cancer genome
atlas (TCGA) and pan-cancer analysis of whole genomes (PCAWG), sequenced large
numbers of tumors and collected data from multiple assays that could be analyzed
together, with the goal of integration and increased understanding of the
mechanisms of cancer.

Considerable progress has been made analyzing these data. Statistical analyses
identify hundreds of global and tissue-specific cancer driver genes
[@deesMuSiC2012; @tamboreroOncodriveCLUST2013;
@lawrenceDiscoverySaturationAnalysis2014;
@kumarStatisticallyIdentifyingTumor2015; @tokheimEvaluatingEvaluationCancer2016;
@jiangWITER2019; @zhaoModelingPositive2019] using approaches aimed at detecting
when genes are mutated at a greater rate than expected due to chance. It has
been estimated that on the order of three mutations in key oncogenes and/or
tumor suppressors would be sufficient to transform a normal cell to a cancerous
state [@vogelsteinPathCancerThree2015].

Other studies aim at providing a comprehensive view of mutations, gene
expression and genomic signatures, with the express goal of understanding the
common themes of all cancer independent of tissue of origin. Understanding
cancer as a disease of the cell has long been a goal of the field and is perhaps
best characterized in essays by @weinberg2000; @hallmarks_II (see also
Discussion). The first of these studies in the genomic era identified 11 subtypes
from 12 cancer types, using integrative analysis with co-equal weighting of gene
expression, methylation, copy number and proteomics data
[@hoadleyMultiplatformAnalysis2014]. The principal finding was that
tissue-of-origin is the predominant driving factor, though a significant
proportion of tumors (~10%) could be reclassified independent of
tissue-of-origin. In a second study involving 33 cancer types and a much greater
number of tumors [@shenIntegrativeClusteringMultiple2009], the authors identify
28 clusters that could be further subdivided into organ specific groups,
including pan-gastrointestinal, pan-gynecological, pan-squamous,
pan-gynecological/ squamous and pan-kidney
[@hoadleyCellofOriginPatternsDominate2018].

More recently, attempts have been made to consider pan-cancer mutations and
other data as disruptions of normal pathway activity. @PCAWG-PATH
carried out near-exhaustive analysis of pathway enrichment using whole genome
data. Horn _et al._ [@hornNetSigNetworkbasedDiscovery2018] introduce approaches
to address the sparseness of mutations, allowing the potential importance of a
gene to be influenced by the aggregate of mutations in its nearest network
neighbors. This method is powerful especially when applied to a small number (on
the order of hundreds) of tumors. One of the challenges in the field remains
that beyond a couple hundred common driver genes and mutations, there are likely
thousands of moderate effect genes that occur at such low frequency--given
tissue diversity--and low sample numbers that they are impossible to detect
using positive selection theory. Some groups have attempted to address this
challenge by applying machine learning to cancer data sets to discover groups of
functionally related genes as they interact with larger pathways and networks
[@kimPancancerAnalysisSomatic2018; @mourikis-ml-pathways; @moonlight]. These
studies have proven very effective at highlighting some potentially fundamental
disease phenotypes at the pathway level that cross cancers with different
origins at the cellular and tissue level.

In this study, we attempt to address these questions by focusing on discrete
molecular pathways as the unit of disruption in cancer. Our approach strips all
cell-type-specific information from the mutation data and equates gene-level
mutations to cell-biological pathway disruptions. We use this heuristic to
evaluate all cancers and show that, surprisingly, tumors fall into categories of
common pathway disruptions with distinctive origins arising from tissue-specific
gene alterations and having unique prognosis in each cancer type.