### 2020-08-20 text/introduction.Rmd

Recent advances in technology have greatly expanded our view into the mechanisms
of cancer at the systems biology level. Array and next-generation sequencing
technologies have made it possible to assay germline and somatic mutations,
expression profiles, DNA methylation, and copy number variations from the same
tissue samples. To take advantage of these tools several large consortia,
including the cancer genome atlas (TCGA) and pan-cancer analysis of whole
genomes (PCAWG), sequenced large numbers of tumors and collected data from
multiple assays that could be analyzed together, with the goal of integration
and increased understanding of the mechanisms of cancer.

Considerable progress has been made analyzing these data. Statistical analyses
identify hundreds of global and tissue-specific cancer driver genes
[@deesMuSiC2012; @tamboreroOncodriveCLUST2013;
@lawrenceDiscoverySaturationAnalysis2014;
@kumarStatisticallyIdentifyingTumor2015; @tokheimEvaluatingEvaluationCancer2016;
@jiangWITER2019; @zhaoModelingPositive2019] using approaches aimed at detecting
when genes are mutated at a greater rate than expected due to chance. It has
been estimated that fewer than five mutations in key oncogenes and/or tumor
suppressors would be sufficient to transform a normal cell to a cancerous state
[@vogelsteinPathCancerThree2015; @Iranzo-Network]. Somewhat paradoxically,
mutations in cancer driver genes are commonplace in somatic tissues and
correlate with age and environmental exposures [@mutations-healthy-tissues].

Other studies provide a comprehensive view of mutations, gene expression and
genomic signatures, with the express goal of understanding the common themes of
all cancer independent of tissue of origin. Understanding cancer as a disease of
the cell has long been a goal of the field as characterized in essays by
@weinberg2000; @hallmarks_II. The first of these studies in the genomic era
identified 11 subtypes from 12 cancer types, using integrative analysis with
co-equal weighting of gene expression, methylation, copy number and proteomics
data [@hoadleyMultiplatformAnalysis2014]. The principal finding was that
tissue-of-origin is the predominant driving factor, though a significant
proportion of tumors (~10%) could be reclassified independent of
tissue-of-origin. In a second study involving 33 cancer types and a much greater
number of tumors [@shenIntegrativeClusteringMultiple2009], the authors identify
28 clusters that could be further subdivided into organ specific groups,
including pan-gastrointestinal, pan-gynecological, pan-squamous,
pan-gynecological/squamous and pan-kidney [@hoadleyCellofOriginPatternsDominate2018].

More recently, attempts have been made to consider pan-cancer mutations and
other data as disruptions of normal pathway activity. @PCAWG-PATH carried out
near-exhaustive analysis of pathway enrichment using whole genome data. Network
approaches address the sparseness of mutations, allowing genes to be influenced
by the aggregate of mutations in the nearest network neighbors
[@hornNetSigNetworkbasedDiscovery2018; @Iranzo-Network]. These methods are
powerful especially when applied to a small number (on the order of hundreds) of
tumors. One of the challenges in the field remains that beyond a couple hundred
common driver genes and mutations, there may be thousands of moderate effect
genes that occur at such low frequency--given tissue diversity--and low sample
numbers that they are impossible to detect using positive selection theory. Some
groups have attempted to address this challenge by applying machine learning to
cancer data sets to discover groups of functionally related genes as they
interact with larger pathways and networks [@kimPancancerAnalysisSomatic2018;
@mourikis-ml-pathways; @moonlight]. These studies have proven very effective at
highlighting some potentially fundamental disease phenotypes at the pathway
level that cross cancers with different origins at the cellular and tissue
level.

In this study, we attempt to address these questions by focusing on discrete
molecular pathways as the unit of disruption in cancer. Our approach strips all
cell-type-specific information from the mutation data and equates gene-level
mutations to cell-biological pathway disruptions. We use this heuristic to
evaluate all cancers and show that, surprisingly, tumors that exhibit
tissue-specific gene mutation patterns nonetheless fall into common categories
of pathway disruption having unique prognoses in each cancer type.