### 2020-08-21 code/figure9.Rmd
add figure9.Rmd

```{r, message = FALSE, fig.width = 9, fig.height= 9}
kelly.colours <- c("gray95", "gray13", "gold2", "plum4", "darkorange1",
                   "lightskyblue2", "firebrick", "burlywood3", "gray51", "springgreen4",
                   "lightpink2", "deepskyblue4", "lightsalmon2", "mediumpurple4",
                   "orange", "maroon", "yellow3", "brown4", "yellow4", "sienna4", "chocolate", "gray19")

cluster.colours <- c("grey", kelly.colours[c(3:12)])
names(cluster.colours) <- c(0:10)

kclassplot = function(dfsamps,tisnames){
	kclnames = grep("k\\[",names(dfsamps), value=TRUE)
	dfclas = dfsamps[,kclnames]
	dfclas = stack(dfclas)
	names(dfclas) = c("value","coefname")
	dfclas$class = as.factor(as.numeric(gsub("k\\[(.*),(.*)\\]","\\2",dfclas$coefname)))
	dfclas$tiss = tisnames[as.numeric(gsub("k\\[(.*),(.*)\\]","\\1",dfclas$coefname))]
	ggplot(dfclas,aes_(quote(value),quote(class), fill = quote(class))) + geom_density_ridges() +
		facet_wrap(. ~ tiss,nrow=3) + theme_minimal() + scale_fill_brewer(palette = "Paired") +
		xlim(0,3) + theme(legend.position = "none") + ggtitle("Class Specific Coefficients")
}

pclass = kclassplot(dfsamps,levels(cancerdat$type))

pclass+mytheme+scale_fill_manual(values = cluster.colours)
```