### 2020-08-21 code/figure1v2.Rmd

```{r, fig.width = 9, fig.height= 4}
load("../data/cancer_disrupt_heatmap.Rda") ## Load  in data processed in figure1.Rmd

fig1 <- ggplot(data = cancer_disrupt_heatmap, aes(x=sample_id, y=path_name, fill = membership)) + 
  geom_tile(size = 0.6) +
  scale_fill_manual(values = c("grey", "firebrick3"),
                    labels = c("Not Mutated", "Mutated")) +
  facet_grid(.~type, scales = "free", space = "free", switch = "x") + ## facet by cancer type; switch: move facet labels top to bottom
  xlab("") + ylab("377 Reactome Pathways") + labs(fill = "") +
  guides(fill = guide_legend(override.aes = list(size = 2))) + ## make fill legend shape smaller
  theme(axis.text.x = element_blank(), ## turn off x axis text
        axis.text.y = element_blank(), ## turn off y axis text
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank(), ## remove tick marks
        strip.text.x = element_text(angle = 90), ## rotate x-axis
        strip.background = element_blank(),
        legend.position = "bottom", ## legend parameters
        legend.key.height = unit(1, "lines"),
        legend.text = element_text(size = 8),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,-10,0,-10))

## fixing the text cutoff in heatmap
 fig1 <- ggplotGrob(fig1)

 for(i in which(grepl("strip-b", fig1$layout$name))){ ## grep all text strips at the bottom of plot 
   fig1$grobs[[i]]$layout$clip <- "off" ## turn off clipping of text
 }

fig1 <- as.ggplot(fig1) ## turn back fig into 
fig1
```