### 2020-08-22 text/discussion.Rmd
reduce discussion, remove references to the results

### Classification of tumors independent of tissue-of-origin

One of the biggest hurdles in cancer research is the sparsity of data; ~20,000
protein-coding genes is comparable with the number of tumor samples, even with
multiple mutations per sample. And that is not accounting for the fact that
~30,000 non-coding genes and hundreds of thousands of non-coding regulatory
elements exist with oncogenic or tumor suppressor activity [@deesMuSiC2012;
@tamboreroOncodriveCLUST2013; @lawrenceDiscoverySaturationAnalysis2014;
@kumarStatisticallyIdentifyingTumor2015; @jiangWITER2019;
@zhaoModelingPositive2019; @noncoding-mutations-Zhang]. Hence, we sought to
simplify the problem by employing a "knowledge-base driven analysis"
[@pathways-Atul-Butte], investigating cancer as a disease of basic cellular and
biochemical pathways, which we accomplished by translating gene-level mutations
into pathway level disruptions. Our approach differs from previously described
methods [@PATHNET] in that we chose to focus on the pathway (defined in the
methods) as the unit of disruption instead of the gene, employing the conceit
that individual mutations may be sufficient to disrupt or misdirect pathway
activity (as is often the case with driver mutations). Rather than evaluating
the burden of mutations on each pathway within or across tumors, which can be
biased by the number of genes or the coverage of pathway-linked genes in the
genome, we scored each pathway in binary fashion, allowing one mutation or
multiple mutations to count equally for each sample.

To our knowledge, this approach has not previously been attempted in spite of
its relative simplicity. We limited our analysis to mutations with likely
deleterious effects, including missense, nonsense, frame-shift, stop-loss,
untranslated region and splice site mutations in genes that are actively
expressed in each cancer type, thus avoiding bias from transcription coupled
repair. Our method of filtration differs from the "rank-and-cut" method of
[@PCAWG-MAIN] but represents a reasonable attempt to account for the same
biases. We also filtered genes on the basis of pathway membership in biochemical
pathways, excluding curated gene sets related to diseases, syndromes, or classes
of proteins with shared catalytic activity or conserved domains which are
potentially problematic [@pathways-Atul-Butte]. We chose this approach to limit
redundancy and exclude biologically unrelated collections of genes. Our
pathway-centric analysis allows us to circumvent the problem that some
cancer-promoting mutations likely occur at such low frequencies as to be
indistinguishable from background. Previous attempts to capture these
associations using Bayesian networks have yielded important new insights into
this problem and new genes [@hornNetSigNetworkbasedDiscovery2018] that would not
have been possible with traditional analysis, supporting the case that
pathway-centric approaches have the potential to increase statistical power.

The hypothesis that cancer is the result of dysfunction from a limited number of
basic cellular processes common to all living human cells was introduced and
later expanded on in a pair of essays by Hanahan and Weinberg [@weinberg2000;
@hallmarks_II]. In this view, all tumors regardless of tissue of origin acquire
six essential phenotypes; sustaining proliferative signaling, evading growth
suppressors, replicative immortality, evading apoptosis, supportive
angiogenesis, tissue invasion and metastasis [@weinberg2000]. Advancements in
the field over the ensuing decade led to the inclusion of new tumor promoting
processes, namely deregulation of cellular energy cycle, tumor promoting
inflammation, evasion of the immune system, genome instability, and on top of
these acquired traits a new appreciation for the concept of the tumor
microenvironment and its contribution to the evolution and promotion of
metastatic disease [@hallmarks_II].

An alternative hypothesis is that every tumor belongs to one of a large number
of syndromes which are unique to each tissue-of-origin, that share some
mechanisms and treatment strategies. Recent publication of TCGA consortium
papers present a view largely, and surprisingly, consistent with this hypothesis
[@hoadleyMultiplatformAnalysis2014; @hoadleyCellofOriginPatternsDominate2018].
Perhaps owing to the intractable complexity of genomics, proteomics, and
patient metadata in all its forms, the inescapable conclusion thus far is that
tissue-of-origin remains the most important driver of tumor characteristics at
every scale and by every measure. Our observations contrast with this view, and
instead support an interpretation of publicly available data in which all tumors
manifest one of a limited number of metastable phenotypes resulting from
disruptions in basic pathways as represented by our clusters.

We attempted to account for our clusters in terms of more trivial explanations.
For example, it could be that the clusters we identified are consistent with a
signature of overall disease progression, such that successive, adjacent
clusters on the UMAP projection exhibit increasing specificity. We identified no
such trend in the number of mutations, the relative staging or metastasis; and
each cluster instead was associated with unique combinations of pathways.

We find that most tumors segregate into one of ten biologically distinct
clusters, regardless of the cancer type. Some cancer types are unevenly
distributed, though we could not identify any cancers that were exclusive to a
single tumor cluster. Only cancer types with the fewest samples were found to be
absent from one or more clusters at all. Somewhat surprisingly to us, this
finding extends to histological subtypes of breast, head and neck cancers,
leukemias, _etc._ This result implies that many, if not most, histological
subtypes could reflect differences in cell-of-origin, rather than fundamental
differences in proximal cause. The four major subtypes of breast cancer
correspond to histological and molecular expression profiles that define them
and how they respond to experimental stimuli [@breast-subtypes2]. It has been
hypothesized that differences in the molecular regulators of development in the
precursor cell types present in the breast epithelium drive the various
histological phenotypes [@breast-subtypes; @breast-subtypes3]. Consistent with
this view we found that breast tumor samples of the Luminal A subtype were
heavily biased toward membership in clusters 1 and 7, and basal tumor samples
were biased toward cluster 6. However, both subtypes also contained samples in
every other cluster (without exception), and Luminal B and Her2 positive samples
are distributed across clusters. We interpret these data to mean that inherited
cell-of-origin molecular profiles may predispose certain precursor cell types
within the breast epithelium to forming tumors of one cluster or another but are
not determinative; this view is compatible with the previously stated hypothesis
but opens the way for a more granular view of individual tumors. It would be
surprising if we did not observe bias for some cancers and subtypes amongst our
classes, since some treatment regimens have greater efficacy for patients of a
given cancer or histological subtype [@breast-subtypes2]. Nonetheless, the basis
for some tumors being treatment-refractory in spite of receiving the standard of
clinical care for diagnostic markers remains elusive. Doubtless some of this is
due to bad luck, as tumors can metastasize and remain dormant years before they
are detected at distal sites, or resistant clones may have already arisen at
undetectable levels [@hallmarks_II], but our analysis suggests the possibility
of identifying more informative molecular, histological or cellular subtypes
that could form a basis for future separation of patients into different
precision treatment regimens.

### Tissue specific manifestation of pathway-centric disruptions

Our results illustrate how unique combinations of mutations in pan-cancer driver
genes with tissue-specific pathway disruptions result in common categories when
viewed at the level of the pathway knowledge-base. Top cancer driver genes (e.g.
PIK3C and TP53) are found in most of the clusters, in spite of the fact that
they contribute to many cluster-specific pathways. This can only be explained as
a result of other less common driver genes complementing the pathway disruption
of the unique combination of driver genes that are disrupted in each tumor, and
we speculate that many of these are relatively tissue-specific in terms of their
sensitivity to mutation. Consistent with our observations, @moonlight showed how
tissue- and cancer-specific driver genes and druggable targets of pathway moduli
can be discovered by machine learning on integrated multi-omics datasets.
Marrying these approaches in future work will greatly enhance our understanding
of tumor biology.

### Incompleteness of the pathway disruption data

Finally, we must remark on the limitations of our work exemplified in two of the
clusters, 1 and 10, for which we did not find many distinctive associations with
pathways. Cluster 1 had a relatively low proportion of mutated pathways across
the board, although it is broadly enriched in many of the same tumor-promoting
pathways common to the other groups. This likely does not reflect a true
difference in stages of tumor evolution, as our data show clearly that this
cluster is as likely to contain stage IV metastatic tumors as it is to contain
those of stage I. This cluster likely represents a group of tumors with
aberrations in methylation, key deletions, or other structural variants.
Consistent with this, kidney chromophobe and thyroid cancers have high
proportions of structural variations _vs._ other variant types [@PCAWG-MAIN] and
are heavily skewed to cluster 1 membership. Likewise, cluster 10 represents a
group of hyper-mutated tumors that harbor so many mutations that virtually no
pathway is unaffected. It seems likely that a significant fraction of the
“mutant” samples for each pathway are burdened with excess passenger mutations.
This could be addressed with application of ever-increasingly sophisticated
filtering of likely passenger mutations [ _e.g._ @spliceAI;
@deep-learning-clinical-variants]. In the future, we hope to incorporate these
other data into a comprehensive pathway-centric analysis as we have done here
for point mutations and indels.

### Estimates of survival reveal pathway-dependent differences

By modeling CDC longevity data as a baseline risk function we showed that each
of our pathway disruption clusters exhibit cancer-type specific effects on
expected survival. Contrary to our initial expectation, we found that models
assuming different survival based on cluster $\times$ cancer type fit better than
models in which they are independent factors. However, considering that within
each cancer type there are different clinical standards of care, and even within
classes of drugs the preferred treatment can very between cancers, it makes
sense that we observe tissue-specific cluster effects. Contrast the situation
with ovarian _vs._ breast cancer, which are both hormonally driven cancers, for
example. Ovarian cancer has but one main treatment axis, platinum, whereas
breast cancer patients have a variety of treatment regimens based on molecular
subtype and other factors. Standard of care for one cancer and cluster could
produce different outcomes than the same cluster of a different cancer type.
Unfortunately, given the diversity of drug classes and treatments (including
non-chemical treatments such as surgery and palliative care), we lack sufficient
power to explore these variables in the TCGA data. It is our hope that future
studies will help to distinguish between treatment-specific effects on survival
given different pathway disruption clusters.

### Implications for the evolution of cancer

Our findings imply that there are separate processes in the etiology of cancer
that can be broadly thought of as general cancer promoting, cluster-specific
mutations and metastasis. General cancer promoting processes must include
factors that relate to genome stability and immortality, as "enabling
characteristics" of the cancer phenotype [@hallmarks_II]. Such pathways are
disrupted in all or most of the defined clusters
and are frequently the result of aberrations involving common driver genes such
as BRCA1/2, MMR genes, mitotic checkpoints, Cohesion complexes, _etc._
Cluster-specific
evolution must involve the acquisition of disruptions to pathways that may
individually be harmful ( _e.g._ highly proliferative cells are more likely to
senesce) but together produce more specialized cancer phenotype and increased
fitness. Importantly, our
observations do not imply the order in which these mutations should accumulate.
This could be addressed in a future study by integrating our analysis with
evolutionary analysis of clonality, drawing inferences from variant allele
frequencies as in @PCAWG-EVO. However, since many of the genes in the
non-cluster-specific pathways involve the known driver genes, it is reasonable
to hypothesize that these mutations promote or enable the subsequent acquisition
of cluster-specific defects via random mutation and natural selection, thus
producing the clusters we observed. In support of this, the pan-cancer analysis
of whole genomes consortium (PCAWG) found that oncogenic driver mutations are
highly enriched in early arising clones, whereas later arising clones have much
greater diversity in driver mutations [@PCAWG-EVO]. Moreover, driver genes that
are known to be responsible for discrete mutation signatures such as APOBEC,
BRCA1 and BRCA2 produce mutational hotspots reflecting the varying selective
pressures in different tissues [@PCAWG-MAIN].

One of the drawbacks of bulk tumor whole genome sequencing data is the problem
of tumor heterogeneity. Consortium samples are likely to contain contamination
from support tissue, stroma, inflammatory cells, immune cells of the innate and
adaptive immune systems, all potentially harboring cancer supporting mutations
[@bhowmick_review]. We think it will be instructive to explore these ideas in
the context of single cell experiments.

### On metastasis as a convergence of phenotypes

We report that pathway enrichment in metastatic tumors across all clusters
(pooled) yielded generally lower effect sizes and larger p-values than the
cluster specific analysis, suggesting that signal is diluted when the samples
are analyzed together, and supporting the view that metastasis has
cluster-specific requirements. Since the number of metastatic samples is
relatively low, this part of our analysis is likely underpowered and subject to
expanded analysis with larger cohorts. The fact that most metastatic enrichment
is cluster-specific and has a tendency to overlap with cluster-specific pathways
from non-metastatic tumors of neighboring clusters suggests a mechanism of
complementarity, in which newly acquired mutations result in similar clusters
converging on one or more deadly phenotypes that capture most of the critical
features of end-stage cancer. Thus, even at relatively low power, our analysis
of metastasis further supports the pathway-centric approach as a fruitful way to
uncover biologically meaningful differences between tumors against the noisy
backdrop of tissue specific profiles. As a major caveat to this, we acknowledge
that many therapies _do_ target highly-specific driver genes, markers, and
signaling pathways ( _e.g._ TP53, EGFR or HER2), but understanding the broader
context of the genetic background and pathway vulnerability of tumors containing
such markers may aid in creating smarter combination therapies. We submit that
when we discover the requirements of each cluster with respect to pathway
disruptions and metastasis we may be able to target them therapeutically and
prevent further adaptation.