### 2020-08-24 code/figure3.Rmd

```{r sankey-plot}
umap_cancertype <- fread("../data/umap_3d_coors.tsv") ## load data
load("../data/kelly.colours.rda")
load("../data/ditto.colours.rda")

sankey.data <- umap_cancertype %>% dplyr::select(sample_id, project_code, clust_knn) %>% ## select relevant columns
  gather(., key = "type", value = "value", -sample_id) ## wide to long format

## Cancer names in alphabetical order
cancerorder <- unique(umap_cancertype$project_code)
cancerorder <- cancerorder[order(cancerorder)]
## Cluster in numerical order
clustorder <- unique(umap_cancertype$clust_knn)
clustorder <- as.character(clustorder[order(clustorder)])

## Factor to force ordering
sankey.data$type <- factor(sankey.data$type, levels = c( "clust_knn", "project_code"))
sankey.data$value <- factor(sankey.data$value, levels= c(cancerorder, clustorder))

## Color scheme for sankey plot
sankey.colours <- c(ditto_colours[2:24],kelly.colours[3:12])
names(sankey.colours) <- c(cancerorder, clustorder)

## Sankey plot
sankey.plot <- ggplot(sankey.data,
       aes(x = type, stratum = value, alluvium = sample_id,
           fill = value, label = value)) +
  scale_fill_manual(values = sankey.colours) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback") +
  geom_stratum(alpha = 0.5) +
  geom_text(stat = "stratum", size = 3) +
  theme_classic() +
  xlab(label = "") +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        plot.margin = unit(c(-0.4, -3, -1, -2.5), "cm"),
        panel.background = element_rect(fill = "transparent"))
```

```{r, fig.width= 6, fig.height= 6}
# start new page
plot.new()

# setup layout
gl <- grid.layout(nrow=8, ncol=3)

## Setup viewports - How figures will be laid out
vp.1 <- viewport(layout.pos.col=1:2, layout.pos.row=1:8)
vp.2 <- viewport(layout.pos.col=3, layout.pos.row=1:2) 
vp.3 <- viewport(layout.pos.col=3, layout.pos.row=3:4) 
vp.4 <- viewport(layout.pos.col=3, layout.pos.row=5:6) 
vp.5 <- viewport(layout.pos.col=3, layout.pos.row=7:8)

# init layout
pushViewport(viewport(layout=gl))

## vp.1 - Figure A - sankey plot
pushViewport(vp.1) ## Enter vp.1 viewpoint
print(sankey.plot, newpage = FALSE) ## Plot ggplot
mtext("A", side=3, line=2.6, cex=1.2, adj=-0.17)
popViewport() ## Exit vp.1 viewpoint

## vp.2 - Figure B - BRCA subtype
pushViewport(vp.2)
par(new=TRUE, fig=gridFIG())
par(mar = c(0,0,0,0), mai=c(0.1,0.1,0.1,0.1), cex.lab = 0.8, cex = 0.8) ## setting plot parameters

brca_subtype <- umap_cancertype %>% filter(project_code == "BRCA") %>% ## Filter for selected cancer type
  dplyr::select(plot_x, plot_y, plot_z, Subtype_Selected) ## Select relevant column 
brca_subtype[is.na(brca_subtype)] <- "NA" ## Convert NA to character "NA"

## Color scheme for subtype
B_name <- unique(brca_subtype$Subtype_Selected) ## What are the subtype names
B_name <- B_name[order(B_name,na.last=T)] ## Order the names ABC and have NA last
B_color <- ditto_colours[c(2:(length(B_name) + 1))] ## Selecting enough colors from ditto
names(B_color) <- B_name ## Assigning subtypes to color

with(brca_subtype,
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = B_color, pch = 21, cex = 0.8, lwd = 0.2, 
               theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
## Axis labels
legend("top", inset = c(0), legend=names(B_color[1:2]), col=B_color[1:2], pch = 20, bty = "n", horiz = T, cex = 0.8)
legend("top", inset = c(0.07), legend=names(B_color[3:4]), col=B_color[3:4], pch = 20, bty = "n", horiz = T, cex = 0.8)
legend("top", inset = c(0.14), legend=names(B_color[5:6]), col=B_color[5:6], pch = 20, bty = "n", horiz = T, cex = 0.8)
title(xlab="UMAP 2", line=-0.25)
mtext("B", side=3, line=-1.2, cex=1.2, adj=0)

popViewport() ## exit

## vp.3 - Figure C - BRCA somatic mutations
pushViewport(vp.3)
par(new=TRUE, fig=gridFIG())
par(mar = c(0,0,0,0), mai=c(0.1,0.1,0.1,0.1), cex.lab = 0.8, cex = 0.8) ## setting plot parameters

load("../data/gdc_gene_exp_adj.Rda") ## load data
load("../data/gene_data.Rda")

brca <- gene_data %>% .[.$external_gene_name %in% c("BRCA1", "BRCA2"),] %>% dplyr::select(ensembl_gene_id, external_gene_name) ## Getting BRCA gene information
brca <- gdc_gene_exp_adj %>% .[rownames(.) %in% brca$ensembl_gene_id, ] %>% ## Include BRCA gene
  rownames_to_column(var = "ensembl_gene_id") %>% ## Move ensembl id to column
  left_join(., brca, by = "ensembl_gene_id") %>% ## left join with brca df 
  remove_rownames()  %>% column_to_rownames(var = "external_gene_name") %>% ## replace row with gene symbol
  dplyr::select(-ensembl_gene_id) %>% ## get rid of ensembl_id col
  t() %>% as.data.frame() %>% ## correct orientation for umap_cancertype df
  rownames_to_column(var = "sample_id") %>% ## getting data ready for leftjoin 
  mutate(BRCA1 = case_when(BRCA1 == "1" ~ "BRCA1", ## Convert booleans to categorical variables
                           BRCA1 == "0" ~ "NA"),
         BRCA2 = case_when(BRCA2 == "1" ~ "BRCA2",
                           BRCA2 == "0" ~ "NA")) %>%
  mutate(mut_type = gsub('NA,|,NA', '', paste(.$BRCA1, .$BRCA2, sep = ","))) %>% ## combine mutation cause some patients can have both
  dplyr::select(-BRCA1, -BRCA2) %>% ## remove col we don't need
  left_join(., umap_cancertype[,c("sample_id", "plot_x", "plot_y", "plot_z", "project_code")], by = "sample_id") ## leftjoin by sample id

C_colors <- c('#33a02c', '#e31a1c', '#ff7f00') ## Color scheme 
names(C_colors) <- c("BRCA1", "BRCA2", "BRCA1,BRCA2") ## Assigning mutation type to color
with(brca %>% filter(mut_type != "NA"), ## Filter samples that do not have a mutation in BRCA1 or BRCA2
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = C_colors, pch = 21, cex = 0.8, lwd = 0.2,
               theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
## Axis labels
legend("top", inset = c(0), legend=names(C_colors[c(1:2)]), col=C_colors[c(1:2)], pch = 20, bty = "n", horiz = T, cex = 0.8)
legend("top", inset = c(0.07), legend=names(C_colors[c(3)]), col=C_colors[3], pch = 20, bty = "n", horiz = T, cex = 0.8)
title(xlab="UMAP 2", line=-0.25)
mtext("C", side=3, line=-1.2, cex=1.2, adj=0)

popViewport()

## vp.4 - Figure D - Stage
pushViewport(vp.4)
par(new=TRUE, fig=gridFIG())
par(mar = c(0,0,0,0), mai=c(0.1,0.1,0.1,0.1), cex.lab = 0.8, cex = 0.8) ## setting plot parameters

D_colors <- c('darkgoldenrod3','dodgerblue3','yellow3','tomato2') ## Color scheme
names(D_colors) <- c("Stage I", "Stage II", "Stage III", "Stage IV") ## Assign stage to color
with(umap_cancertype %>% filter(ajcc_pathologic_tumor_stage != "Not Available"), ## Filter samples that do not have stage info
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = D_colors, pch = 21, cex = 0.8, lwd = 0.2,
               theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
## Axis labels
legend("top", inset = c(0), legend=names(D_colors[1:2]), col=D_colors[1:2], pch = 20, bty = "n", horiz = T, cex = 0.8)
legend("top", inset = c(0.07), legend=names(D_colors[3:4]), col=D_colors[3:4], pch = 20, bty = "n", horiz = T, cex = 0.8)
title(xlab="UMAP 2", line=-0.25)
mtext("D", side=3, line=-1.2, cex=1.2, adj=0)

popViewport()

## vp.5 - Figure E - Metastasis
pushViewport(vp.5)
par(new=TRUE, fig=gridFIG())
par(mar = c(0,0,0,0), mai=c(0.1,0.1,0.1,0.1), cex.lab = 0.8, cex = 0.8) ## setting plot parameters

E_colors <- c('#a6cee3','#1f78b4') ## Color scheme
names(E_colors) <- c("M0", "M1") ## Assign metastasis to color
with(umap_cancertype %>% filter(ajcc_metastasis_pathologic_pm %in% c("M0", "M1")), ## Only M0 and M1 tumor samples
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = E_colors, pch = 21, cex = 0.8, lwd = 0.2,
               theta = 0, phi = 65, scale = F,
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
## Axis labels
legend("top", inset = c(0), legend=names(E_colors), col=E_colors, pch = 20, bty = "n", horiz = T, cex = 0.8)
title(xlab="UMAP 2", line=-0.25)
mtext("E", side=3, line=-1.2, cex=1.2, adj=0)

popViewport()
```