### 2020-08-24 code/figure2.Rmd

```{r figure-2-plot, fig.width = 9}
umap_cancertype <- data.table::fread("../data/umap_3d_coors.tsv") ## load data
load("../data/kelly.colours.rda")
cluster.colours <- c("grey", kelly.colours[c(3:12)]) ## Color scheme for cluster
names(cluster.colours) <- c(0:10) ## assigning cluster to the colors

theta <- c(0, 90, 270)
phi <- c(65, 15, 15)

opar <- par(mfrow = c(1,3), mai = c(0.1, 0.1, 0.1, 0.1)) ## plot layout 2x3
par(mar = c(2,2,0,0), bg = NA) ## setting plot margins for the following plots

for(n in 1) {
  with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                  bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2, # clust_knn is knn result
                                  theta = theta[n], phi = phi[n], scale = T,
                                  xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                  colvar = NULL))
  title(xlab="UMAP 2", line=-5.2)
  for(p in 2:length(phi)) {
    with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2, 
                                theta = theta[p], phi = phi[p], scale = T,
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
  }
}

```