### 2020-08-25 text/abstract.Rmd

Cancer is a complex disease involving dysfunctional cellular metabolism, other
basic biochemical processes, and the microenvironment. However, in spite of some
generally agreed upon unifying principles [@weinberg2000; @hallmarks_II],
molecular signatures remain largely indistinguishable from tissue-of-origin,
presenting a major barrier for precision health and individualized medicine. To
address this challenge, we reduced mutation data to disruptions in a select set
of pathways relevant to basic cell biology, from DNA replication to cellular
communication. Using dimensionality reduction techniques, we assigned tumor
samples into ten clusters distinct from their tissues-of-origin and largely free
of bias from mutational burden or clinical stage. We showed that these tumor
classes vary in prognosis by modeling relative risk of death in different
classes and cancer types. We identify cluster-specific mutations in different
tissues, demonstrating that tissue-specific signatures contribute to common
cellular phenotypes. Moreover, driver gene mutations involved in replication
fidelity and genome stability, such as _BRCA1_ and _Rb1_, are equally
distributed among our clusters, contrary to the expectation that such genes are
avatars of molecular subtype. We also investigated pathways that distinguish
between metastatic and non-metastatic tumors, and found that most meaningful
differences are cluster-specific. Some metastasis enriched pathways from one
cluster are cluster-specific pathways from non-metastatic tumors of another
cluster, suggesting phenotypic convergence, and potentially highlighting key
pathways to focus on for prevention of cluster-specific relapse. Taken as a
whole, our observations suggest that common driver genes combine with
tissue-specific disruptions in tumor-promoting pathways to produce a limited
number of distinct molecular phenotypes. Eventually these signatures converge on
one or more end-stage disease phenotypes. Thus, we present a coherent view of
global tumor biology, and explain how common cellular dysfunction might arise
from tissue-specific mutations.