### 2020-08-25 code/figureSM2.Rmd

```{r SM2}
## mortality data from CDC:
namesmort = c("agerng","pdead","nsurv","ndie","pyrs","pyrsabove","expectedlife")
menmort = read_csv("../data/survival/lifetables/Table02.csv",skip=2)
## drop the final row, it's just a text info statement
menmort = menmort[1:NROW(menmort)-1,] 
wommort = read_csv("../data/survival/lifetables/Table03.csv",skip=2)
wommort = wommort[1:NROW(wommort)-1,]

names(menmort) = namesmort
names(wommort) = namesmort
yrs=seq(0,100,by=1);
menmort$yrnum = yrs
wommort$yrnum = yrs

load("../code/survival/survmodel.stansave")

sampsum = summary(samps)

asamps = as.array(samps)

r20f = mean(asamps[,,"r20[1]"])
r20m = mean(asamps[,,"r20[2]"])
agrf = mean(asamps[,,"agerate[1]"])
agrm = mean(asamps[,,"agerate[2]"])

#sprintf("r20f mean = %f",r20f)
#sprintf("r20m mean = %f",r20m)
#sprintf("agrf mean = %f", agrf)
#sprintf("agrm mean = %f", agrm)

ourlccdf = function(t,age,agerate,k,r20){
    r20 = r20 * 1e-5
    agerate = agerate*.05
    return(-r20*((365*exp((agerate*k*t)/365+age*agerate-20*agerate))/(agerate*k)
        -(365*exp(age*agerate-20*agerate))/(agerate*k)));
}

ccdfsimpmkr = function(r20,agert){
    function(age){
        exp(ourlccdf(age*365,0,agert,1,r20))
    }
}

mccdf = ccdfsimpmkr(r20m,agrm)
fccdf = ccdfsimpmkr(r20f,agrf)

mytheme = theme(
#    axis.text.x = element_blank(),
#    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
#    axis.ticks = element_blank(),
#    axis.line = element_blank(),
    strip.text.y.right = element_text(angle = 0),
    strip.background = element_blank(),
    legend.key.height = unit(0.4, "lines"),
    legend.title = element_blank(),
    #legend.position = "none",
    title = element_text(size=10)
  )

mortpl <- ggplot(menmort) + 
    geom_line(aes(yrnum,nsurv/100000, color = "CDC men"),data=menmort) +
    geom_line(aes(yrnum,nsurv/100000, color ="CDC women"),data=wommort) +
    stat_function(fun=function(x) mccdf(x), aes(color = "Model fit men")) +
    stat_function(fun=function(x) fccdf(x), aes(color = "Model fit women")) + 
    labs(title="1-cdf(age) for Men and Women (CDC Data vs model fit)",x="Age (yrs)",y="1-cdf", color = "") +
    theme_classic() +
    theme(axis.title.x = element_text(size = 10),
          axis.title.y = element_text(size = 10),
          title = element_text(size = 10),
          legend.position = c(0.85, 0.9),
          legend.text = element_text(size = 8),
          legend.key.size = unit(0.4,"cm")) +
    scale_color_manual(values = c("CDC men" = "#0072B2", "CDC women" = "#D55E00", 
                                  "Model fit men" = "#009E73", "Model fit women" = "#E69F00"))

print(mortpl)
```