### 2020-08-26 text/abstract.Rmd
incorporate Megan Hitchens' comments, fix spaces after parens

Cancer is a complex disease involving disrupted cellular metabolism, basic
biochemical processes, and the microenvironment. However, despite some generally
agreed upon unifying principles [@weinberg2000; @hallmarks_II], molecular
signatures remain largely indistinguishable from tissue-of-origin, presenting a
major barrier for precision health and individualized medicine. To address this
challenge, we reduce mutation data to disruptions in a select set of pathways
relevant to basic cell biology, from DNA replication to cellular communication.
Using dimensionality reduction techniques, we assign tumor samples into ten
clusters distinct from tissue-of-origin and largely free of bias from mutational
burden or clinical stage. We show that the clusters vary in prognosis by
modeling relative risk of death by cancer type and cluster. We identify
cluster-specific mutations in different tissues, demonstrating that
tissue-specific signatures contribute to common cellular phenotypes. Moreover,
germline risk genes involved in replication fidelity and genome stability are
equally distributed among clusters, contrary to the expectation that such genes
are avatars of molecular subtype. We investigate metastatic and non-metastatic
pathways, and show that most differences are cluster-specific. Some metastatic
pathways from one cluster are cluster-specific pathways from non-metastatic
tumors of another cluster, suggesting phenotypic convergence. Taken as a whole,
our observations suggest that common driver genes combine with tissue-specific
disruptions in tumor-promoting pathways to produce a limited number of distinct
molecular phenotypes. Thus, we present a coherent view of global tumor biology,
and explain how common cellular dysfunction might arise from tissue-specific
mutations.