### 2020-08-28 text/discussion.Rmd
cut all text to conform to Cell's 45k character limit

### Classification of tumors independent of tissue-of-origin

One of the biggest hurdles in cancer research is the sparsity of data; ~20,000
protein-coding genes is comparable with the number of tumor samples, even with
multiple mutations per sample. We sought to
simplify the problem by employing a "knowledge-base driven analysis"
[@pathways-Atul-Butte], investigating cancer as a disease of basic cellular and
biochemical pathways, which we accomplished by translating gene-level mutations
into pathway level disruptions. Our approach differs from previously described
methods [@PATHNET] in that we chose to focus on the pathway (defined in the
methods) as the unit of disruption instead of the gene, where individual
mutations may be sufficient to alter pathway activity.

To our knowledge, this approach has not previously been attempted despite
its relative simplicity. We limited our analysis to mutations with likely
deleterious effects, including missense, nonsense, frame-shift, stop-loss,
untranslated region and splice site mutations in genes that are actively
expressed in each cancer type, thus avoiding bias from transcription coupled
repair. Our method of filtration differs from the "rank-and-cut" method of
[@PCAWG-MAIN] but represents a reasonable attempt to account for the same
biases. We also restricted our analysis to biochemical
pathways, excluding curated gene sets related to diseases, syndromes, or classes
of proteins with shared catalytic activity or conserved domains which are
potentially problematic [@pathways-Atul-Butte]. We chose this approach to limit
redundancy and exclude biologically unrelated collections of genes.

The hypothesis that cancer is the result of dysfunction from a limited number of
basic cellular processes common to eukaryotic cells was introduced and
later expanded on in a pair of essays by Hanahan and Weinberg [@weinberg2000;
@hallmarks_II]. An alternative hypothesis is that every tumor belongs to one of a large number
of syndromes which are unique to each tissue-of-origin, that share some
mechanisms and treatment strategies. Recent publication of TCGA consortium
papers present a view largely, and surprisingly, consistent with this latter hypothesis
[@hoadleyMultiplatformAnalysis2014; @hoadleyCellofOriginPatternsDominate2018].
Perhaps owing to the intractable complexity of genomics, proteomics, and
patient metadata in all its forms, the inescapable conclusion thus far is that
tissue-of-origin remains the most important driver of tumor characteristics at
every scale and by every measure. Our observations contrast with this view, and
instead support an interpretation of publicly available data in which all tumors
manifest one of a limited number of phenotypes resulting from
disruptions of basic pathways.

We attempted to account for our clusters in terms of more trivial explanations.
For example, it could be that the clusters we identified are consistent with
disease progression. We were unable to identify any such trend in the number of
mutations, the relative staging or metastasis; and each cluster instead was
associated with unique combinations of pathways.

Some cancer types are unevenly
distributed among the clusters, though we could not identify any cancers that were exclusive to a
single cluster. Only cancer types with the fewest samples were found to be
absent from one or more clusters at all. Somewhat surprisingly to us, this
finding extends to histological subtypes of breast, head and neck cancers,
leukemias, _etc._ This result implies that histological
subtypes could reflect differences in cell-of-origin, rather than fundamental
differences in cancer phenotype. The four major subtypes of breast cancer
correspond to histological and molecular expression profiles that define them
and how they respond to experimental stimuli [@breast-subtypes2]. It has been
hypothesized that differences in the molecular regulators of development in the
precursor cell types present in the breast epithelium drive the various
histological phenotypes [@breast-subtypes; @breast-subtypes3]. Consistent with
this view we found that breast tumor samples of the Luminal A subtype were
heavily biased toward membership in clusters 1 and 7, and basal tumor samples
were biased toward cluster 6. However, both subtypes also contained samples in
every other cluster (without exception), and Luminal B and Her2 positive samples
are distributed across clusters. We interpret these data to mean that inherited
cell-of-origin signatures could predispose certain precursor cells
within the breast epithelium to forming tumors of one cluster or another but are
not determinative. This view is compatible with the previously stated hypothesis
but opens the way for a more granular view of individual tumors. It would be
surprising if we did not observe bias for some cancers and subtypes amongst our
classes, since some treatment regimens have greater efficacy for patients of a
given cancer or histological subtype [@breast-subtypes2]. Nonetheless, the basis
for some tumors being treatment-refractory in spite of receiving the standard of
clinical care for diagnostic markers remains elusive. Doubtless some of this is
due to chance events, as tumors can metastasize and remain dormant years before
they are detected at distal sites, or resistant clones may have already arisen
at undetectable levels [@hallmarks_II], but our analysis suggests the
possibility of identifying more informative molecular, histological or cellular
subtypes that could form a basis for future stratification of patients into
different precision treatment regimens.

### Tissue specific manifestation of pathway-centric disruptions

Our results illustrate how unique combinations of mutations in pan-cancer driver
genes with tissue-specific pathway disruptions result in common categories when
viewed at the level of the pathway knowledge-base. Top cancer driver genes (_e.g.
PIK3CA_ and _TP53_) are found in most of the clusters, in spite of the fact that
they contribute to many cluster-specific pathways. This can only be explained as
a result of other less common driver genes complementing the pathway disruption
of the unique combination of driver genes that are disrupted in each tumor, and
we speculate that many of these are sensitive to mutation in different tissues.

### Incompleteness of the pathway disruption data

Finally, we must remark on the limitations of our work exemplified in two of the
clusters, 1 and 10, for which we did not find many distinctive associations with
pathways. Cluster 1 had a relatively low proportion of mutated pathways,
although it is broadly enriched in many of the same tumor-promoting pathways
common to the other groups. In addition, our data show clearly that this cluster
is as likely to contain stage IV metastatic tumors as it is to contain those of
stage I. This cluster likely represents a group of tumors with aberrations in
methylation, copy number, or other structural variants. Consistent with this,
kidney chromophobe and thyroid cancers have high proportions of structural
variations _vs._ other variant types [@PCAWG-MAIN] and are heavily skewed to
cluster 1 membership. Likewise, cluster 10 represents a group of hyper-mutated
tumors that harbor so many mutations that virtually no pathway is unaffected. It
seems likely that a significant fraction of the “mutant” samples for each
pathway are burdened with excess passenger mutations. This could be addressed
with more sophisticated filtering of likely passenger mutations [_e.g._
@spliceAI;
@deep-learning-clinical-variants]. In the future, we hope to incorporate these
other data into a comprehensive pathway-centric analysis as we have done here
for point mutations and indels.

### Estimates of survival reveal pathway-dependent differences

By modeling CDC longevity data as a baseline risk function we showed that each
of our pathway disruption clusters exhibit cancer-type specific effects on
expected survival. However, considering that within each cancer type there are
different clinical standards of care, and even within classes of drugs the
preferred treatment can very between cancers, it makes sense that we observe
tissue-specific cluster effects. Contrast the situation with ovarian _vs._
breast cancer, which are both hormonally driven cancers, for example. Ovarian
cancer has but one main treatment axis, platinum, whereas breast cancer patients
have a variety of treatment regimens based on molecular subtype and other
factors. Unfortunately, given the diversity of drug classes and treatments, we
lack sufficient power to explore these variables in the TCGA data. It is our
hope that future studies will help to distinguish between treatment-specific
effects on survival given different pathway disruption clusters.

### Implications for the evolution of cancer

Our findings imply that there are separate processes in the etiology of cancer
that can be broadly thought of as general cancer promoting, cluster-specific
mutations and metastasis. General cancer promoting processes include genome
stability and immortality, as "enabling characteristics" of the cancer phenotype
[@hallmarks_II]. Such pathways are disrupted in most clusters and are frequently
the result of aberrations involving common driver genes such as BRCA1/2, MMR
genes, mitotic checkpoints, cohesion complexes, _etc._ Cluster-specific
evolution must involve the acquisition of disruptions to pathways that may
individually be harmful (_e.g._ highly proliferative cells are more likely to
senesce) but together produce more specialized cancer phenotype and increased
fitness. Importantly, our observations do not imply the order in which these
mutations should accumulate. This could be addressed in a future study by
evolutionary analysis of clonality, drawing inference from variant allele
frequencies as in @PCAWG-EVO. However, since many of the genes in the
non-cluster-specific pathways involve the known driver genes, it is reasonable
to surmise that these mutations promote or enable acquisition of
cluster-specific defects via random mutation and natural selection, thus
producing the clusters we observed. In support of this, the pan-cancer analysis
of whole genomes consortium (PCAWG) found that oncogenic driver mutations are
highly enriched in early arising clones, whereas later arising clones have much
greater diversity in driver mutations [@PCAWG-EVO]. Moreover, driver genes that
are known to be responsible for discrete mutation signatures such as APOBEC,
BRCA1 and BRCA2 produce mutational hotspots reflecting varying selective
pressures in different tissues [@PCAWG-MAIN].

One of the drawbacks of bulk tumor whole genome sequencing data is the problem
of tumor heterogeneity. Consortium samples are likely to contain contamination
from support tissue, stroma, inflammatory cells, immune cells of the innate and
adaptive immune systems, all potentially harboring cancer supporting mutations
[@bhowmick_review]. We think it will be instructive to explore these ideas in
the context of single cell experiments.

### On metastasis as a convergence of phenotypes

We report that enrichment in metastatic tumors across all clusters yielded
generally lower effect sizes and larger p-values than the cluster specific
analysis, suggesting that signal is diluted when clusters are pooled, and
supporting the view that metastasis has cluster-specific requirements. Since the
number of metastatic samples is relatively low, this part of our analysis is
likely underpowered and subject to expanded analysis with larger cohorts. The
fact that most metastatic enrichment is cluster-specific and has a tendency to
overlap with cluster-specific pathways from non-metastatic tumors of neighboring
clusters suggests that newly acquired mutations result in similar clusters
converging on one or more deadly phenotypes with critical features of end-stage
cancer. Thus, even at relatively low power, our analysis of metastasis uncovered
differences between tumors against the noisy backdrop of tissue specific
profiles. As a major caveat to our pan-cancer analysis, we acknowledge that many
therapies _do_ target highly-specific driver genes, markers, and signaling
pathways ( _e.g._ TP53, EGFR or HER2), but understanding the broader context of
the genetic background and pathway vulnerability of tumors containing such
markers may aid in creating smarter combination therapies. We submit that when
we discover the requirements of each cluster with respect to pathway disruptions
and metastasis we may be able to target them therapeutically and prevent further
adaptation.