### 2020-08-28 code/survival/survfig.Rmd

```{r, survfig, echo=FALSE, warning=FALSE, error=TRUE,message=FALSE,fig.height=12, fig.width=9}

library(ggridges)
library(patchwork)

source("./kmcurve.R")

## create example data for a fixed 30 year old without vs with a k=2
## cancer

patientsurvivalsbkd = replicate(100,simdeath(r20wf,ageratewf,30,1))/365
patientsurvivalsk2 = replicate(100,simdeath(r20wf,ageratewf,30,2))/365

mixedages = runif(100,30,70)
mixedsurvwfk1 = sapply(mixedages, function(x) simdeath(r20wf,ageratewf,x,1)/365)
mixedsurvwfk2 = sapply(mixedages, function(x) simdeath(r20wf,ageratewf,x,2)/365)

p1 = data.frame(
    r = patientsurvivalsbkd,
    c = patientsurvivalsk2,
    rm = mixedsurvwfk1,
    cm = mixedsurvwfk2,
    status = rep(1, length(patientsurvivalsbkd))
) %>%
    rownames_to_column(var = "index") %>%
    pivot_longer(-c(index, status), values_to = "time", names_to = "condition") %>%
    ggplot(aes(time = time, color = condition, status = status)) +
    geom_km() +
    scale_color_manual(
        values = c(r = "#56b4e9", c = "#e69f00", rm = "#0072b2", cm = "#d55e00"),
        breaks = c("r", "c", "rm", "cm"),
        labels = c("Random 30 year old women", "Random 30 year old women cancer k=2", "Random uniform(30,70) year old women", "Random uniform(30,70) year old women k=2")
    ) +
    theme_classic() +
    theme(legend.position = c(.8, .75))

pstad = plotcancertype(cancerdat,"STAD")
pgbm = plotcancertype(cancerdat,"GBM")
ppaad = plotcancertype(cancerdat,"PAAD")
thca = plotcancertype(cancerdat,"THCA")

load("./survmodel-ethbkd.stansave")

dfsamps = as.data.frame(sampseth)

kcancplot = function(dfsamps,tisnames){
	ktisnames = grep("ktis",names(dfsamps),value=TRUE)
	dfktis = dfsamps[,ktisnames]
	dfktis = stack(dfktis) ## convert to two column "values" and "ind" with "ind" being the name of the variable
	names(dfktis) = c("value","parname")

	dim(tisnames) = c(length(tisnames),1)

	dfktis$cancname = as.factor(sapply(dfktis$parname,function(x) tisnames[x]))
	ggplot(dfktis,aes_(quote(value),quote(cancname))) + geom_density_ridges()+ theme_minimal() + scale_fill_brewer(palette="Paired") + xlim(0,22) + labs(x=expression(paste("k"["tis"], " multiplier")),y="Cancer Type",title="Tissue Specific Effects")
}

kclassplot = function(dfsamps,tisnames){
	kclnames = grep("k\\[",names(dfsamps), value=TRUE)
	dfclas = dfsamps[,kclnames]
	dfclas = stack(dfclas)
	names(dfclas) = c("value","coefname")
	dfclas$class = as.factor(as.numeric(gsub("k\\[(.*),(.*)\\]","\\2",dfclas$coefname)))
	dfclas$tiss = tisnames[as.numeric(gsub("k\\[(.*),(.*)\\]","\\1",dfclas$coefname))]
	ggplot(dfclas,aes_(quote(value),quote(class))) + geom_density_ridges() +
		facet_wrap(. ~ tiss,nrow=3) + theme_minimal() + scale_fill_brewer(palette = "Paired") +
		xlim(0,3) + theme(legend.position = "none") + ggtitle("Class Specific Coefficients")
}

pcanc = kcancplot(dfsamps,levels(cancerdat$type))
pclass = kclassplot(dfsamps,levels(cancerdat$type))

mytheme = theme(
#    axis.text.x = element_blank(),
#    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
#    axis.ticks = element_blank(),
#    axis.line = element_blank(),
    strip.text.y.right = element_text(angle = 0),
    strip.background = element_blank(),
    legend.key.height = unit(0.4, "lines"),
    legend.title = element_blank(),
    legend.position = c(0.7, 0.8),
    #legend.position = "none",
    title = element_text(size=10)
  )

kmlayout = "
AAFF
AAFF
AAFF
BBFF
BBFF
BBFF
CCFF
CCFF
CCFF
DDFF
DDFF
DDFF
EE##
"

BE_legend <- get_legend(thca[[1]] + theme(legend.direction = "horizontal"))
kmlegend <- theme(legend.direction = "horizontal")

  # ((p1+mytheme)/((pstad[[1]]+mytheme + kmlegend)+(pgbm[[1]]+mytheme + kmlegend) +
  #                  (ppaad[[1]]+mytheme + kmlegend) + (thca[[1]] + mytheme) + (kmlegend) + (pcanc+mytheme) + guide_area() + plot_layout(design = kmlayout, guides = "collect"))) + plot_annotation(tag_levels = "A")
layout = "
AAAA
AAAA
AAAA
AAAA
AAAA
AAAA
BBFF
BBFF
CCFF
CCFF
DDFF
DDFF
EEFF
EEFF
"

turn.off.legend <- theme(legend.position = "none")

(p1+mytheme) + (pstad[[1]]+mytheme) + (pgbm[[1]]+mytheme + turn.off.legend) + (ppaad[[1]]+mytheme + turn.off.legend) + (thca[[1]] + mytheme + turn.off.legend) + (pcanc+mytheme) + plot_annotation(tag_levels = "A") + plot_layout(design = layout)

```