### 2020-08-28 text/supplemental_methods.Rmd

---
title: "Supplementary methods"
author: "Peter T. Nguyen, Simon G. Coetzee, Daniel L. Lakeland, and Dennis J. Hazelett"
date: "8/24/2020"
bibliography: references.bib
output:
  word_document:
    reference_docx: "../text/word-style-reference-ttmanu.docx"
always_allow_html: TRUE
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi=600, warning = FALSE, message = FALSE)
library(foreach)
library(sqldf)
library(tidyverse)
library(ggthemes)
library(data.table)
library(biomaRt)
library(dplyr)
library(GenomicRanges)
library(SummarizedExperiment)
library(reactome.db)
library(plyr)
library(cluster)
library(pheatmap)
library(clusterProfiler)
library(org.Hs.eg.db)
library(dbscan)
library(class)
library(TCGAbiolinks)
library(patchwork)
library(ggplotify)
library(ComplexHeatmap)
library(knitr)
library(kableExtra)
library(UpSetR)
library(cowplot)
library(doRNG)
library(uwot)
library(janitor)
library(FactoMineR)
library(seriation)
library(rtracklayer)
library(poibin)
library(flextable)
library(plot3D)
library(ggfortify)
#library(maftools)
library(survival)
library(ggalluvial)
```

## Supplementary Methods

### Clustering

In order to classify tumors using this dataset, we used multiple correspondence
analysis (MCA) [@JSSv025i01]. First, we determined the number of dimensions
containing useful information by selecting the eigenvalue with the most
explanatory power, using the average of 100 permutations of the data as baseline
(**figure SM1A**). We then chose the maximum eigenvalue for which the p-value
remained ≤ 0.05 (see cutoff in **figure SM1B**). Then we performed a UMAP
analysis [@McInnes2018], both in order to summarize the MCA graphically, and as
a preprocessing step to boost the performance of density based clustering. The
resulting map was notable for its lobed structure, with several reproducible
projections regardless of random seed setting. A representative version of this
3D UMAP is shown in **figure SM1C**, rotated to enhance the visibility of the
major features. Following this spatial mapping we attempted to define groupings
of similar tumors within the spatial map using HDBSCAN, which performs
hierarchical clustering and provides metrics of cluster stability and
probabilities of cluster membership for each node [@hdbscan]. However, HDBSCAN
is sensitive to several parameters; key for our analysis are the minimum number
of tumor samples in a cluster that capture the maximum number of tumor samples,
measured by probability of membership of ≥ 5% in at least one cluster. Thus, we
created a score metric as the fraction of classified tumors with max probability
< 5% in one cluster and chose a cluster size of 92 to minimize the score
function (**figure SM1D**). HDBSCAN with these settings resulted in ten distinct
high-density clusters which we then projected onto the UMAP (**figure SM1E**).
This classified 6,038 out of 7,607 tumors but still resulted in a significant
fraction of unclassified tumor samples. Since we ultimately wish to be able to
classify any tumor using this scheme, we performed k Nearest Neighbors (kNN)
analysis, which computes a similarity metric to every tumor in the set and then
lets the _k_ most similar tumors "vote" as to the identity of the query tumor
sample based on their cluster labels. We set _k_ to be the square root of the
number of tumor samples (87). Using this method, we assigned cluster membership
to the remaining tumors (**figure SM1F**).

```{r, results='asis'}
cat("\\newpage")
```

#### Figure SM1
```{r child='../code/figureSM1.Rmd'}
```

**Figure SM1: Method for tumor classification using pathway aberration data** 
A) MCA analysis was performed to determine the eigenvalue with the most
explanatory power from the categorical data. Average of 100 permutations of the
data was used as a baseline. B) The p-value of each eigenvalue was plotted and
the max eigenvalue with a pvalue < 0.05 was selected. C) The resulting MCA was
projected down to 3 dimensions for the purpose of density based clustering (only
the first 2 dimensions are shown). Each dot represents a tumor D) HDBSCAN was
used to perform hierarchical density based clustering tuned to minimize the
fraction of unassigned tumors. E) HDBSCAN clustered tumors were visualized on
the UMAP projection; tumors are colored by cluster membership (see legend). This
results in some tumors remaining unclassified (grey points). F) kNN was used to
assign membership to the remaining tumors.

```{r, results='asis'}
cat("\\newpage")
```

### Selection of pathways

To understand the molecular mechanism of cancer at a pathway level, we used
Reactome (https://reactome.org/), a knowledge-based pathway database. The
mapping files of ENSEMBL genes to pathways, pathway hierarchy relationships, and
complexes to top pathways were downloaded from
https://reactome.org/download-data. Using these data, we imposed pathway
criteria to define basic cellular processes and biochemical pathways: (1)
human-derived pathways (\it{“HSA”}) (2) limited to grandchild node for each
parent pathway (_e.g._ ‘Beta-catenin independent WNT signaling’ in ‘Signal
Transduction’) (3) exclusion of pathways in the parent pathway: "Disease",
"Muscle contraction", and "Reproduction" or pathway names that include any of
the following keywords: "disease", "defect", "cancer", "oncogenic", "disorder",
"mutant", "loss", "infection", "bacteria", or "listeria". While some of the
excluded pathways have been shown to play an important role in cancer, they are
highly specialized (_e.g._ "PI3K/AKT Signaling in Cancer"). Additionally, for
most of the excluded pathways, a neutral version pathway of the pathway exists
(_e.g._ "PIP3 activates Akt signaling"). Finally, we mapped the 18,577
Ensembl IDs from the TCGA dataset to the highly selected Reactome pathways. This
operation produced a lookup table that consisted of 377 pathways mapped to 8,940
genes.

### Filtering genes

We filtered likely erroneous mutations due to transcription coupled repair. Our
approach was to determine the status of each gene (_i.e._ expressed or not
expressed) in each tissue type in order to exclude gene mutations low expressed
enes. To do this, we obtained the TCGA RNA sequencing data adjusted for batch
effect dataset (https://pancanatlas.xenahubs.net). Using the data, we removed
the genes and tumor samples that were not included in our analysis, grouped the
tumor samples by tissue type and computed the mean expression value for each
gene. A minimum threshold of 10 transcripts per million was set for expressed
genes based on an inflection point observed when plotting the mean expression
values of genes ranked by expression in each tissue type. Genes that did not
meet this threshold were considered not expressed. This operation produced a
lookup table for gene expression status in each tumor sample for 18,127 genes.

### Class and stage-specific Enrichment Calculations

We calculated the enrichment of pathways in one set of tumors as the relative
fraction of tumors (with estimated uncertainty) with a mutated gene in the
pathway to all tumors, inclusive of the category of interest. To do this
operation, we use the beta distribution to permute a posterior distribution on
the fraction of tumors with a pathway mutated for each category based on the
observed set, and compared this to the posterior obtained from the full set of
tumors (all tumors) as the distribution of differences between all permuted
samples. We considered a pathway enriched if the 95% range of credible
differences thus obtained excludes 0 and the mean credible difference was
greater than or equal to 30% enrichment, which excludes a large number of small
differences that are not likely to be biologically relevant.

### Survival

We used Bayesian inference to explore the relative impact of cancer diagnosis on
survival at the time of diagnosis. Our model assumes that at birth each person
has a small initial probability per unit time to die, and that this probability
grows exponentially in time at some baseline rate that matches the observed CDC
data. The rate of growth in risk is assumed to jump to a higher level at
diagnosis, which we take as a proxy for onset. This represents an accelerated
aging type model. To fit the model, we constrain the baseline risks using the
available CDC life tables (ftp://ftp.cdc.gov/pub/Health_Statistics/NCHS/Publications/NVSR/68_07/). We also constrain the cancer specific multiplier
using the longevity data within the cancer tumor dataset. Our model estimates
the effect of cancer type and class-specific cancer effects independently,
resulting in a cancer and class specific estimate of the aging rate multiplier.

Most TCGA cases are diagnosed late in life, so we fit a model that emphasizes
accuracy in the right tail of the distribution. We modeled baseline longevity
using a risk rate function:

$$R(t) = R_{20}\exp(a(A-20) + D(t) k a (\frac{t}{365}))$$

Time is split into two components. $A$ going from 0 up to the age of the patient
at diagnosis and then remaining fixed, whereas $t$ starts at 0 upon diagnosis
and represents the days since diagnosis, a field recorded in the dataset. $D(t)$
is an indicator for whether diagnosis has occurred or not (before diagnosis
$D(t) = 0$, after $D(t) = 1$). The model assumes that the risk of death is
$R_{20}$ at age 20 years and increases exponentially with a constant rate $a$ up
to the age at diagnosis. Risk of death assumes a new rate $ka$ thereafter.
Though the model extends to earlier ages, relatively few patients were under age
20.

Given this risk per unit time, the probability of death at time $T+dT$ is the
probability to survive to time $T$ which is $(1-P(T))$ and then die in the
remaining interval which is $R(T)dT$

$$dP = (1-P(T)) R(T) dT$$

leading to the differential equation for the cumulative probability of death
$P(T)$

$$\frac{dP}{dT} = (1-P(T))R(T)$$

The solution of this ODE is:

$$P(T) = 1 - \exp(-\int_0^T R(t) dt)$$

And the density of deaths per unit time is the derivative with respect to T:

$$p(T) = R(T) \exp(-\int_0^T R(t)dt)$$

The $R(T)$ function has two important parameters, the $a$ value which represents
the background risk rates, and the $k$ value, which is a function of both cancer
tissue type, and the cancer cluster identity.

$$k = k_{tis} k_{cl}$$

The product $k_{tis}k_{cl}$ is obviously symmetric between the two factors. To
disambiguate the meaning of the two, the prior distribution of $k_{cl}$ has peak
probability at 1 and a relatively narrow width, due to the fact that it is a
tissue specific multiplicative modifying factor. This allows the overall
magnitude of $k$ to be primarily determined by the $k_{tis}$ value whose prior
is significantly less constrained so that there is a wider range of risk across
tissue types.

The $k$ in our model can be taken as, roughly, the aging rate relative to the
baseline rates. Thus, a value of $k=2$ means that a given cancer causes you to
experience the same risk in time $dt$ as a non-cancer patient would experience
at the same age, in time $2dt$.

Using this model, we split the data by Male/Female sex, and by White _vs._
Black race as these were the main well-identified categories available to us in
the dataset. Sex is known to be a risk factor for death, with males dying at
slightly higher rates for all ages of interest in our dataset. Black _vs._ White
race is an important category due to differences in access to care and other
socioeconomic and education related factors. However, due to the low number of
Black patient tumors, we use ethnicity only to identify the baseline risk
function, while the tissue- and cluster-specific multipliers are not ethnicity
specific. There were too few tumors from other races to obtain good inference so
we excluded them from our analysis.

**Supplementary figure SM2** gives a comparison between the post-fit
survival curves for men and women vs the CDC data it is based on. This
comparison does not distinguish on ethnicity for simplicity in
visualization.

```{r, results='asis'}
cat("\\newpage")
```

#### Figure SM2

```{r child='../code/figureSM2.Rmd'}
```

**Supplementary Figure SM2** Comparison between CDC data and ODE model fit for
males and females of European descent.

```{r, results='asis'}
cat("\\newpage")
```

As an aside, we did not explicitly account for stage at diagnosis in our model
for the following reasons. In the tumor sample data, stage at diagnosis is
confounded with cancer type because some cancers are screened aggressively (_e.g._ colorectal and prostate cancers) while others are diagnosed typically
after they become problematic for the patient's lifestyle (_e.g._ ovarian and
pancreatic cancers). Secondly, such a model specification would suffer from
added noise because the staging data are not well standardized across cancer
types, have different criteria, and because it is unclear what the relationship
between stage and advancement of disease is (for example some sub-stage 4 tumors
are metastatic). To compound this latter issue, our tumor set is vastly
under-powered given the uneven representation of stage across cancer types.

### References