### 2020-08-28 code/figure2.Rmd
Update text in manuscript and file names

```{r sankey-plot}
umap_cancertype <- fread("../data/umap_3d_coors.tsv") ## load data
load("../data/kelly.colours.rda")
load("../data/ditto.colours.rda")

sankey.data <- umap_cancertype %>% dplyr::select(sample_id, project_code, clust_knn) %>% ## select relevant columns
  gather(., key = "type", value = "value", -sample_id) ## wide to long format

## Cancer names in alphabetical order
cancerorder <- unique(umap_cancertype$project_code)
cancerorder <- cancerorder[order(cancerorder)]
## Cluster in numerical order
clustorder <- unique(umap_cancertype$clust_knn)
clustorder <- as.character(clustorder[order(clustorder)])

## Factor to force ordering
sankey.data$type <- factor(sankey.data$type, levels = c( "clust_knn", "project_code"))
sankey.data$value <- factor(sankey.data$value, levels= c(cancerorder, clustorder))

## Color scheme for sankey plot
sankey.colours <- c(ditto_colours[2:24],kelly.colours[3:12])
names(sankey.colours) <- c(cancerorder, clustorder)

## Sankey plot
sankey.plot <- ggplot(sankey.data,
       aes(x = type, stratum = value, alluvium = sample_id,
           fill = value, label = value)) +
  scale_fill_manual(values = sankey.colours) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback") +
  geom_stratum(alpha = 0.5) +
  geom_text(stat = "stratum", size = 3) +
  theme_classic() +
  xlab(label = "") +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        plot.margin = unit(c(-0.4, -3, -1, -2.5), "cm"),
        panel.background = element_rect(fill = "transparent"))
```

```{r, message = FALSE}

## Arranging mutation freq for each cluster
mut_order <- list() ## store data in list 
for(n in 1:max(as.numeric(as.character(umap_cancertype$clust_knn)))) { ## loop through all the clusters
  mut_order[[n]] <- umap_cancertype %>% filter(clust_knn == n) %>%  ## filter for specific class
    dplyr::select(sample_id, plot_x, plot_y, clust_knn, mut_count) %>% ## select relevant columns
    arrange(mut_count) ## arrange mutation in increasing order
}
mut_order <- mut_order %>% do.call(rbind, .) %>% as.data.frame() ## rbind list

## Convert sample_id and clust_knn into factor
mut_order$sample_id <- factor(mut_order$sample_id, levels = mut_order$sample_id)
mut_order$clust_knn <- factor(mut_order$clust_knn, levels = c(1:10))

## Determining the breaks positions for the x-axis
previous_class <- 0 ## starting position 1
position_of_previous_class <- 0 ## starting position 2
xaxis_breaks <- list() ## list for the for loop
for(n in 1) { ## loop through 1 to 10 clusters
  class_spec <- round(mut_order %>% filter(clust_knn == n) %>% nrow(.) / 2, digits = 0) ## calculate halfway mark of class 
  xaxis_breaks[[n]] <- class_spec + previous_class + position_of_previous_class ## adding up all variables
  previous_class <- class_spec ## storing class halfway mark
  for(n in 2:max(as.numeric(as.character(mut_order$clust_knn)))) {
    position_of_previous_class <- xaxis_breaks[[n-1]] ## getting the previous class position
    class_spec <- round(mut_order %>% filter(clust_knn == n) %>% nrow(.) / 2, digits = 0) ## calculate halfway mark of new class
    xaxis_breaks[[n]] <- class_spec + previous_class + position_of_previous_class ## adding up all variables
    previous_class <- class_spec ## storing class halfway mark
  }
}
xaxis_breaks <- as.data.frame(do.call(rbind, xaxis_breaks)) ## rbind list
xaxis_breaks <- mut_order[rownames(mut_order) %in% xaxis_breaks$V1,] ## find the tumor samples in these positions

total_n <- as.data.frame(table(mut_order$clust_knn)) ## getting total n in cluster

mut_order$log10mut <- log10(mut_order$mut_count + 1) ## mutation count log transformation + pseudo count

## Calculate median for each class
mut_median <- mut_order %>% 
  group_by(clust_knn) %>% 
  dplyr::summarize(mut_med = median(log10mut))

## Need to add 100 pseudo samples front and end of data so they don't touch the y-axis
mut_order <- mut_order[,c("sample_id", "clust_knn", "log10mut")] ## selecting relevant column

begin_NA <- data.frame(sample_id = paste0("sample", 1:100), 
                       clust_knn = min(as.numeric(mut_order$clust_knn)), ## put lowest cluster to prevent a new facet
                       log10mut = NA) ## put NA values so they don't get plotted
end_NA <- data.frame(sample_id = paste0("sample", 101:200), ## repeat the step for the ends
                       clust_knn = max(as.numeric(mut_order$clust_knn)),
                       log10mut = NA)
mut_order <- rbind(begin_NA, mut_order, end_NA) ## rbind the data in a specfic order

mut_order$sample_id <- factor(mut_order$sample_id, levels = mut_order$sample_id) ## factor samples
mut_order$clust_knn <- factor(mut_order$clust_knn, levels = c(1:10)) ## factor clusters

fig1G <- ggplot(mut_order, aes(x = sample_id, y = log10mut, color = clust_knn)) + 
  geom_point(size = 0.25) + 
  ylab("Log10(Mutation Count)") + 
  xlab("") +
  theme_classic() +
  facet_wrap(~ factor(clust_knn, levels = c(1:10)), ncol = 10, scales = "free_x") + ## facet by cluster
  geom_hline(data= mut_median, aes(yintercept=mut_med)) + ## add median line
  scale_y_continuous(sec.axis = sec_axis(~.)) + ## dup y-axis
  scale_x_discrete(breaks = c(as.character(mut_order$sample_id[1]), as.character(xaxis_breaks$sample_id)), labels = c("n =" ,total_n$Freq)) +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        axis.ticks.x = element_blank(),
        axis.line.x = element_line(colour = "black", size = 0.5),
        axis.line.x.top = element_line(colour = "black", size = 0.5),
        panel.spacing = unit(0.25, "lines"),
        strip.background = element_rect(size = 0.5),
        strip.placement = "outside",
        strip.text.x = element_text(size = 10),
        legend.position = "none") +
  scale_color_manual(values = kelly.colours[-c(1:2)])

## code to keep only the bottom border line of the facet
fig1G <- ggplotGrob(fig1G)

lg <- linesGrob(x=unit(c(0,1),"npc"), y=unit(c(0,0),"npc"), ## All lines off expect bottom line of facet
                   gp=gpar(col="black", lwd=4))

for (k in grep("strip-t",fig1G$layout$name)) { ## turn off 3 lines from the facet border text
  fig1G$grobs[[k]]$grobs[[1]]$children[[1]] <- lg
}

fig1G <- as.ggplot(fig1G)
```

```{r, fig.width= 6, fig.height = 8}
# start new page
plot.new()

# setup layout
gl <- grid.layout(nrow=11, ncol=3)

## Setup viewports - How figures will be laid out
vp.1 <- viewport(layout.pos.col=1:2, layout.pos.row=1:8)
vp.2 <- viewport(layout.pos.col=3, layout.pos.row=1:2) 
vp.3 <- viewport(layout.pos.col=3, layout.pos.row=3:4) 
vp.4 <- viewport(layout.pos.col=3, layout.pos.row=5:6) 
vp.5 <- viewport(layout.pos.col=3, layout.pos.row=7:8)
vp.6 <- viewport(layout.pos.col=1:3, layout.pos.row=9:11)

# init layout
pushViewport(viewport(layout=gl))

## vp.1 - Figure A - sankey plot
pushViewport(vp.1) ## Enter vp.1 viewpoint
print(sankey.plot, newpage = FALSE) ## Plot ggplot
mtext("A", side=3, line=2.6, cex=1, adj=-0.17)
popViewport() ## Exit vp.1 viewpoint

## vp.2 - Figure B - BRCA subtype
pushViewport(vp.2)
par(new=TRUE, fig=gridFIG())
par(mar = c(0,0,0,0), mai=c(0.1,0.1,0.1,0.1), cex.lab = 0.8, cex = 0.8) ## setting plot parameters

brca_subtype <- umap_cancertype %>% filter(project_code == "BRCA") %>% ## Filter for selected cancer type
  dplyr::select(plot_x, plot_y, plot_z, Subtype_Selected) ## Select relevant column 
brca_subtype[is.na(brca_subtype)] <- "NA" ## Convert NA to character "NA"

## Color scheme for subtype
B_name <- unique(brca_subtype$Subtype_Selected) ## What are the subtype names
B_name <- B_name[order(B_name,na.last=T)] ## Order the names ABC and have NA last
B_color <- ditto_colours[c(2:(length(B_name) + 1))] ## Selecting enough colors from ditto
names(B_color) <- B_name ## Assigning subtypes to color

with(brca_subtype,
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = B_color, pch = 21, cex = 0.8, lwd = 0.2, 
               theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
## Axis labels
legend("top", inset = c(0), legend=names(B_color[1:2]), col=B_color[1:2], pch = 20, bty = "n", horiz = T, cex = 0.8)
legend("top", inset = c(0.07), legend=names(B_color[3:4]), col=B_color[3:4], pch = 20, bty = "n", horiz = T, cex = 0.8)
legend("top", inset = c(0.14), legend=names(B_color[5:6]), col=B_color[5:6], pch = 20, bty = "n", horiz = T, cex = 0.8)
title(xlab="UMAP 2", line=-0.25)
mtext("B", side=3, line=-1.2, cex=1, adj=0)

popViewport() ## exit

## vp.3 - Figure C - BRCA somatic mutations
pushViewport(vp.3)
par(new=TRUE, fig=gridFIG())
par(mar = c(0,0,0,0), mai=c(0.1,0.1,0.1,0.1), cex.lab = 0.8, cex = 0.8) ## setting plot parameters

load("../data/gdc_gene_exp_adj.Rda") ## load data
load("../data/gene_data.Rda")

brca <- gene_data %>% .[.$external_gene_name %in% c("BRCA1", "BRCA2"),] %>% dplyr::select(ensembl_gene_id, external_gene_name) ## Getting BRCA gene information
brca <- gdc_gene_exp_adj %>% .[rownames(.) %in% brca$ensembl_gene_id, ] %>% ## Include BRCA gene
  rownames_to_column(var = "ensembl_gene_id") %>% ## Move ensembl id to column
  left_join(., brca, by = "ensembl_gene_id") %>% ## left join with brca df 
  remove_rownames()  %>% column_to_rownames(var = "external_gene_name") %>% ## replace row with gene symbol
  dplyr::select(-ensembl_gene_id) %>% ## get rid of ensembl_id col
  t() %>% as.data.frame() %>% ## correct orientation for umap_cancertype df
  rownames_to_column(var = "sample_id") %>% ## getting data ready for leftjoin 
  mutate(BRCA1 = case_when(BRCA1 == "1" ~ "BRCA1", ## Convert booleans to categorical variables
                           BRCA1 == "0" ~ "NA"),
         BRCA2 = case_when(BRCA2 == "1" ~ "BRCA2",
                           BRCA2 == "0" ~ "NA")) %>%
  mutate(mut_type = gsub('NA,|,NA', '', paste(.$BRCA1, .$BRCA2, sep = ","))) %>% ## combine mutation cause some patients can have both
  dplyr::select(-BRCA1, -BRCA2) %>% ## remove col we don't need
  left_join(., umap_cancertype[,c("sample_id", "plot_x", "plot_y", "plot_z", "project_code")], by = "sample_id") ## leftjoin by sample id

C_colors <- c('#33a02c', '#e31a1c', '#ff7f00') ## Color scheme 
names(C_colors) <- c("BRCA1", "BRCA2", "BRCA1,BRCA2") ## Assigning mutation type to color
with(brca %>% filter(mut_type != "NA"), ## Filter samples that do not have a mutation in BRCA1 or BRCA2
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = C_colors, pch = 21, cex = 0.8, lwd = 0.2,
               theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
## Axis labels
legend("top", inset = c(0), legend=names(C_colors[c(1:2)]), col=C_colors[c(1:2)], pch = 20, bty = "n", horiz = T, cex = 0.8)
legend("top", inset = c(0.07), legend=names(C_colors[c(3)]), col=C_colors[3], pch = 20, bty = "n", horiz = T, cex = 0.8)
title(xlab="UMAP 2", line=-0.25)
mtext("C", side=3, line=-1.2, cex=1, adj=0)

popViewport()

## vp.4 - Figure D - Stage
pushViewport(vp.4)
par(new=TRUE, fig=gridFIG())
par(mar = c(0,0,0,0), mai=c(0.1,0.1,0.1,0.1), cex.lab = 0.8, cex = 0.8) ## setting plot parameters

D_colors <- c('darkgoldenrod3','dodgerblue3','yellow3','tomato2') ## Color scheme
names(D_colors) <- c("Stage I", "Stage II", "Stage III", "Stage IV") ## Assign stage to color
with(umap_cancertype %>% filter(ajcc_pathologic_tumor_stage != "Not Available"), ## Filter samples that do not have stage info
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = D_colors, pch = 21, cex = 0.8, lwd = 0.2,
               theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
## Axis labels
legend("top", inset = c(0), legend=names(D_colors[1:2]), col=D_colors[1:2], pch = 20, bty = "n", horiz = T, cex = 0.8)
legend("top", inset = c(0.07), legend=names(D_colors[3:4]), col=D_colors[3:4], pch = 20, bty = "n", horiz = T, cex = 0.8)
title(xlab="UMAP 2", line=-0.25)
mtext("D", side=3, line=-1.2, cex=1, adj=0)

popViewport()

## vp.5 - Figure E - Metastasis
pushViewport(vp.5)
par(new=TRUE, fig=gridFIG())
par(mar = c(0,0,0,0), mai=c(0.1,0.1,0.1,0.1), cex.lab = 0.8, cex = 0.8) ## setting plot parameters

E_colors <- c('#a6cee3','#1f78b4') ## Color scheme
names(E_colors) <- c("M0", "M1") ## Assign metastasis to color
with(umap_cancertype %>% filter(ajcc_metastasis_pathologic_pm %in% c("M0", "M1")), ## Only M0 and M1 tumor samples
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = E_colors, pch = 21, cex = 0.8, lwd = 0.2,
               theta = 0, phi = 65, scale = F,
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
## Axis labels
legend("top", inset = c(0), legend=names(E_colors), col=E_colors, pch = 20, bty = "n", horiz = T, cex = 0.8)
title(xlab="UMAP 2", line=-0.25)
mtext("E", side=3, line=-1.2, cex=1, adj=0)

popViewport()

## vp.1 - Figure A - sankey plot
pushViewport(vp.6) ## Enter vp.1 viewpoint
print(fig1G, newpage = FALSE) ## Plot ggplot
mtext("F", side=3, line=-9.8, cex=1, adj=-2.4)
popViewport() ## Exit vp.1 viewpoint
```