### 2020-08-28 code/figure6.Rmd
Add SM3 to supplemental_methods

```{r, message = FALSE, fig.width = 9, fig.height= 9}
load("../data/kelly.colours.rda")
cluster.colours <- c("grey", kelly.colours[c(3:12)])
names(cluster.colours) <- c(0:10)

## Getting n_race for white and black
cancerdat <- read_csv("../data/survival/clinical_plus_cluster.csv")
cancer.clusttype <- cancerdat %>% dplyr::select(type, race, clust_knn) %>% ## select relevant columns
  filter(race %in% c("WHITE", "BLACK OR AFRICAN AMERICAN")) ## filter black and white only
cancer.race <- as.data.frame(table(cancer.clusttype$type)) ## table cancer type
cancer.race$facetname <- paste0(cancer.race$Var1, " (N = ", cancer.race$Freq, (")")) ## Create a new col with type and freq together
cancer.race$Freq <- NULL ## Remove Freq Column
names(cancer.race) <- c("tiss", "facetname") ## rename cols for left join in kclassplot below

## Getting n_cluster within cancer.type
cancer.nclustintype <- as.data.frame(table(cancer.clusttype$type, cancer.clusttype$clust_knn))
cancer.nclustintype <- cancer.nclustintype %>% mutate(alpha = case_when(Freq < 10 ~ "0.05",
                                                                        Freq >= 10 & Freq < 25 ~ "0.5",
                                                                        Freq >= 25 ~ "1.0"))
textlist <- list()
cancertype <- "BRCA"
for(n in cancertype) {
  df <- cancer.nclustintype %>% filter(Var1 == n)
  cancer.race$lab <- as.character(paste(df[1, "alpha"], "\n",df[2, "alpha"], "\n", df[3, "alpha"], "\n",df[4, "alpha"], "\n",df[5, "alpha"], "\n",df[6, "alpha"], "\n",df[7, "alpha"], "\n",df[8, "alpha"], "\n",df[9, "alpha"], "\n",df[10, "alpha"]))
}

textlist <- as.data.frame(do.call(rbind, textlist))

tisnames <- levels(cancerdat$type)

  ## Processing data
	kclnames = grep("k\\[",names(dfsamps), value=TRUE)
	dfclas = dfsamps[,kclnames]
	dfclas = stack(dfclas)
	names(dfclas) = c("value","coefname")
	dfclas$class = as.factor(as.numeric(gsub("k\\[(.*),(.*)\\]","\\2",dfclas$coefname)))
	dfclas$tiss = tisnames[as.numeric(gsub("k\\[(.*),(.*)\\]","\\1",dfclas$coefname))]
	dfclas$class <- fct_rev(dfclas$class)

	## 
	dfclas <- left_join(dfclas, cancer.race, by = "tiss")

	## Plotting
	ggplot(dfclas,aes_(x = quote(value),y = quote(class), fill = quote(class))) + geom_density_ridges() +
		facet_wrap(. ~ facetname,nrow=3) + theme_minimal() + scale_fill_brewer(palette = "Paired") +
		xlim(0,3) + theme(legend.position = "none") + ggtitle("Class Specific Coefficients") +
	  labs(x = expression(paste("k"["cl"])), y = "Cluster Identity") +
	  geom_text(data = dfclas, aes(label = lab), x = 2.5, y = 6) +
	  mytheme + turn.off.legend +scale_fill_manual(values = cluster.colours)
```