### 2020-08-31 code/figure1.Rmd
Add legend for fig1B

```{r, fig.width = 9, fig.height= 4}
load("../data/cancer_disrupt_heatmap.Rda") ## Load  in data processed in figure1.Rmd

fig1 <- ggplot(data = cancer_disrupt_heatmap, aes(x=sample_id, y=path_name, fill = membership)) + 
  geom_tile(size = 0.6) +
  scale_fill_manual(values = c("grey", "firebrick3"),
                    labels = c("Not Mutated", "Mutated")) +
  facet_grid(.~type, scales = "free", space = "free", switch = "x") + ## facet by cancer type; switch: move facet labels top to bottom
  xlab("") + ylab("377 Reactome Pathways") + labs(fill = "") +
  guides(fill = guide_legend(override.aes = list(size = 2))) + ## make fill legend shape smaller
  theme(axis.text.x = element_blank(), ## turn off x axis text
        axis.text.y = element_blank(), ## turn off y axis text
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks = element_blank(), ## remove tick marks
        strip.text.x = element_text(angle = 90), ## rotate x-axis
        strip.background = element_blank(),
        legend.position = "bottom", ## legend parameters
        legend.key.height = unit(1, "lines"),
        legend.text = element_text(size = 8),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,-10,0,-10),
        plot.margin = unit(c(0.6, -3, 0, 0), "cm"),
        panel.background = element_rect(fill = "transparent"))

## fixing the text cutoff in heatmap
 fig1 <- ggplotGrob(fig1)

 for(i in which(grepl("strip-b", fig1$layout$name))){ ## grep all text strips at the bottom of plot 
   fig1$grobs[[i]]$layout$clip <- "off" ## turn off clipping of text
 }

fig1 <- as.ggplot(fig1) ## turn back fig into 
```

```{r fig.width = 9, fig.height= 6}
# start new page
plot.new()

# setup layout
gl <- grid.layout(nrow=2, ncol=3)

## Setup viewports - How figures will be laid out
vp.1 <- viewport(layout.pos.col=1:3, layout.pos.row=1)
vp.2 <- viewport(layout.pos.col=1, layout.pos.row=2) 
vp.3 <- viewport(layout.pos.col=2, layout.pos.row=2) 
vp.4 <- viewport(layout.pos.col=3, layout.pos.row=2)

# init layout
pushViewport(viewport(layout=gl))

## vp.1 - Figure A - sankey plot
pushViewport(vp.1) ## Enter vp.1 viewpoint
print(fig1, newpage = FALSE) ## Plot ggplot
mtext("A", side=3, line=3.15, cex=1.2, adj=-0.095)
popViewport() ## Exit vp.1 viewpoint

## Data for different perspectives of umap
umap_cancertype <- data.table::fread("../data/umap_3d_coors.tsv") ## load data
load("../data/kelly.colours.rda")
cluster.colours <- c("grey", kelly.colours[c(3:12)]) ## Color scheme for cluster
names(cluster.colours) <- c(0:10) ## assigning cluster to the colors

theta <- c(0, 90, 270)
phi <- c(65, 15, 15)
## View 1
pushViewport(vp.2)
par(new=TRUE, fig=gridFIG())
par(mar = c(0,0,0,0), mai=c(0.1,0.1,0.1,0.1), cex.lab = 0.8, cex = 0.8) ## setting plot parameters
n <- 1
with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                  bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2, # clust_knn is knn result
                                  theta = theta[n], phi = phi[n], scale = T,
                                  xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                  colvar = NULL))
title(xlab="UMAP 2", line=-1.5)
legend("top", inset = c(0.1), legend=names(cluster.colours[2:6]), col=cluster.colours[2:6], pch = 20, bty = "n", horiz = T, cex = 0.8)
legend("top", inset = c(0.15), legend=names(cluster.colours[7:11]), col=cluster.colours[7:11], pch = 20, bty = "n", horiz = T, cex = 0.8)
mtext("B", side=3, line=-1.2, cex=1.2, adj=0)
popViewport()
## View 2
pushViewport(vp.3)
par(new=TRUE, fig=gridFIG())
par(mar = c(0,0,0,0), mai=c(0.1,0.1,0.1,0.1), cex.lab = 0.8, cex = 0.8) ## setting plot parameters
n <- 2
with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                  bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2, # clust_knn is knn result
                                  theta = theta[n], phi = phi[n], scale = T,
                                  xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                  colvar = NULL))

popViewport()
## View 3
pushViewport(vp.4)
par(new=TRUE, fig=gridFIG())
par(mar = c(0,0,0,0), mai=c(0.1,0.1,0.1,0.1), cex.lab = 0.8, cex = 0.8) ## setting plot parameters
n <- 3
with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                  bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2, # clust_knn is knn result
                                  theta = theta[n], phi = phi[n], scale = T,
                                  xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                  colvar = NULL))
popViewport()

```