### 2020-08-31 text/abstract.Rmd
reduce abstract from 250 to 150 words (Cell requirement)

Cancer affects cellular metabolism, biochemical processes, and the
microenvironment. However, despite some unifying principles, molecular
signatures remain largely indistinguishable from tissue-of-origin, presenting a
major barrier for precision health and individualized medicine. To address this,
we assign tumors into ten clusters based on pathways largely free of bias from
tissue-of-origin, mutational burden or stage. Germline risk genes are equally
distributed among clusters, contrary to the expectation that they determine
molecular subtype. We show that clusters vary in prognosis, modeling risk of
death by cancer type and cluster. We investigate metastasis, and show that most
differences are cluster-specific and complement other clusters, suggesting
convergence. Our observations suggest that driver genes combine with uncommon
genes in tumor-promoting pathways to produce distinct molecular phenotypes.
Thus, we present a coherent view of global tumor biology, and explain how common
cellular dysfunction might arise from tissue-specific mutations.