### 2020-08-31 code/figure4.Rmd
Modify Figure 4 to generate 2 panel figure (combining enriched / depleted into a single forest plot + heatmap)

```{r forest-heatmap-data-generation, eval = FALSE}
load("../data/kelly.colours.rda")

cluster.colors <- kelly.colours[-c(1)]
names(cluster.colors) <- 0:20
cluster_colors <- cluster.colors[2:11]
names(cluster_colors) <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10")

### zcat simple_somatic_mutation.open.tsv.gz | grep -Fw -f mutation_classes.txt > simple_somatic_mutation.open.selmuts.tsv
### cat <(zcat simple_somatic_mutation.open.tsv.gz | head -1) simple_somatic_mutation.open.selmuts.tsv | gzip -c - >simple_somatic_mutation.open.selmuts.tsv.gz
all_muts <- icgcSimpleMutationToMAF("../data/simple_somatic_mutation.open.selmuts.tsv.gz", removeDuplicatedVariants = FALSE, addHugoSymbol = TRUE)
all_muts <- distinct(.data = all_muts,
                     Chromosome, Start_Position, End_Position, 
                     Reference_Allele, Tumor_Seq_Allele2, consequence_type, 
                     mutation_type, ens_id, icgc_donor_id, 
                     submitted_sample_id, .keep_all = TRUE) %>% 
  rename(icgc_sample_id = Tumor_Sample_Barcode) %>% 
  mutate(Tumor_Sample_Barcode = stringr::str_trunc(submitted_sample_id, 16, ellipsis = ""))

# vep.export <- all_muts %>% select(Chromosome, Start_Position, End_Position, Reference_Allele, Tumor_Seq_Allele2, Strand, icgc_mutation_id)
# vep.export <- vep.export %>% arrange(Chromosome, Start_Position)
# vep.export <- distinct(vep.export)
# vep.export <- with(vep.export,
#                    paste(Chromosome, Start_Position, End_Position,
#                          paste(Reference_Allele, Tumor_Seq_Allele2, sep = "/"),
#                          Strand, icgc_mutation_id, sep = "\t"))
# writeLines(vep.export, con = "../data/vep.export.tsv")
# vep settings availible in "../data/VEP_settings.png"

clinical_plus_cluster <- readr::read_csv("../data/clinical_plus_cluster.csv")
patient_ids <- clinical_plus_cluster$patient_id
solid_tumor_studies <- unique(clinical_plus_cluster$type)
tumor_clusters <- readr::read_tsv("../data/umap_3d_coors.tsv", guess_max = 10000)

clinical_plus_cluster <- clinical_plus_cluster %>% rename(Tumor_Sample_Barcode = sample_id)
clinical_plus_cluster <- left_join(clinical_plus_cluster,
                                   tumor_clusters %>% 
                                     select(starts_with(c("plot", "cluster", "clust_knn", "sample_id"))),
                                   by = c("dna_mutation_id" = "sample_id"))
all_muts <- all_muts %>% filter(Tumor_Sample_Barcode %in% clinical_plus_cluster$Tumor_Sample_Barcode)

vep_output <- readr::read_tsv("../data/vep_output.basic.txt", na = c("", "NA", "-"), guess_max = 10000)

selected_mut <- c("3_prime_UTR_variant", "5_prime_UTR_premature_start_codon_gain_variant", "5_prime_UTR_variant", 
                  "frameshift_variant", "missense_variant", "splice_acceptor_variant", "splice_donor_variant",
                  "stop_lost", "start_lost", "stop_gained")
selected_mut_alt <- c("3'UTR", "5'UTR", 
                      "Frame_Shift_Del", "Frame_Shift_Ins",
                      "Missense_Mutation", 
                      "Splice_Site", "Translation_Start_Site", 
                      "Nonsense_Mutation", "Nonstop_Mutation")

vep_output <- vep_output %>% 
  rename(icgc_mutation_id = "#Uploaded_variation")

all_muts <- all_muts %>% 
  left_join(vep_output, 
            by = c("consequence_type" = "Consequence", 
                   "icgc_mutation_id" = "icgc_mutation_id",
                   "ens_id" = "Gene"))

clinical_plus_cluster$clust_knn_char <- factor(clinical_plus_cluster$clust_knn, 
                                               levels = c(1:10), 
                                               labels = c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10"))

all_muts <- left_join(all_muts, clinical_plus_cluster %>% select(Tumor_Sample_Barcode, clust_knn, clust_knn_char, type) %>% distinct())

for( cluster_c in levels(clinical_plus_cluster$clust_knn_char) ) {
  c_muts <- read.maf(maf = all_muts %>% filter(clust_knn_char == cluster_c),
                      clinicalData = clinical_plus_cluster %>% filter(clust_knn_char == cluster_c),
                      vc_nonSyn = selected_mut_alt, 
                      removeDuplicatedVariants = FALSE)
  other_muts <- read.maf(maf = all_muts %>% filter(clust_knn_char != cluster_c),
                         clinicalData = clinical_plus_cluster %>% filter(clust_knn_char != cluster_c),
                         vc_nonSyn = selected_mut_alt, 
                         removeDuplicatedVariants = FALSE)

  c_compare <- mafCompare( c_muts,
                           other_muts )
  c_compare <- c_compare$results %>% 
    as_tibble() %>% 
    filter(adjPval < 0.01) %>%
    arrange(or) %>% 
    mutate(Hugo_Symbol = factor(Hugo_Symbol, levels = Hugo_Symbol)) %>%
    mutate(ci.up = log(ci.up), ci.low = log(ci.low), or = log(or)) %>%
    mutate(cluster = cluster_c)

  if (cluster_c == "01") {
    all_compare <- c_compare
  } else {
    all_compare <- bind_rows(all_compare, c_compare)
  }
}

save(all_compare, all_muts, clinical_plus_cluster, cluster_colors, file = "../data/forest.heatmap.data.rda")
```

```{r forest-heatmap-data-plot, fig.align='center', fig.width= 9, fig.height= 8, message=FALSE}
#fig.width= 16, fig.height= 20,
load("../data/forest.heatmap.data.rda")

down_compare_plot <- all_compare %>%
  group_by(cluster) %>%
  filter(or < 0) %>%
  filter(adjPval < 0.01) %>%
  arrange(-or) %>%
  filter(is.finite(or)) %>%
  slice_max(order_by = or, n = 10) %>%
  arrange(cluster, or, .by.group = TRUE)
down_compare_plot$Hugo_Symbol <- with(down_compare_plot,
                                      factor(as.character(Hugo_Symbol),
                                             levels = unique(as.character(Hugo_Symbol))))

up_compare_plot <- all_compare %>%
  group_by(cluster) %>%
  filter(or > 0) %>%
  arrange(-or) %>%
  filter(is.finite(or)) %>%
  slice_max(order_by = or, n = 10) %>%
  arrange(cluster, or, .by.group = TRUE)
up_compare_plot$Hugo_Symbol <- with(up_compare_plot,
                                    factor(as.character(Hugo_Symbol),
                                           levels = unique(as.character(Hugo_Symbol))))

all_compare_plot <- bind_rows(down_compare_plot, up_compare_plot)
all_compare_plot %<>% arrange(or, .by_group = TRUE) %>% mutate(Hugo_Symbol = as.character(Hugo_Symbol)) %>% ungroup() %>% mutate(Row_Index = sprintf("%3i", row_number()))
Hugo_rownames <- all_compare_plot$Hugo_Symbol
names(Hugo_rownames) <- as.character(all_compare_plot$Row_Index)
all_p <- ggplot(all_compare_plot, 
               aes(x = or,
                   y = Row_Index,down_compare_plot <- all_compare %>%
  group_by(cluster) %>%
  filter(or < 0) %>%
  filter(adjPval < 0.01) %>%
  arrange(-or) %>%
  filter(is.finite(or)) %>%
  slice_max(order_by = or, n = 10) %>%
  arrange(cluster, or, .by.group = TRUE)
down_compare_plot$Hugo_Symbol <- with(down_compare_plot,
                                      factor(as.character(Hugo_Symbol),
                                             levels = unique(as.character(Hugo_Symbol))))

up_compare_plot <- all_compare %>%
  group_by(cluster) %>%
  filter(or > 0) %>%
  arrange(-or) %>%
  filter(is.finite(or)) %>%
  slice_max(order_by = or, n = 10) %>%
  arrange(cluster, or, .by.group = TRUE)
up_compare_plot$Hugo_Symbol <- with(up_compare_plot,
                                    factor(as.character(Hugo_Symbol),
                                           levels = unique(as.character(Hugo_Symbol))))

all_compare_plot <- bind_rows(down_compare_plot, up_compare_plot)
all_compare_plot %<>% arrange(or, .by_group = TRUE) %>% mutate(Hugo_Symbol = as.character(Hugo_Symbol)) %>% ungroup() %>% mutate(Row_Index = sprintf("%3i", row_number()))
Hugo_rownames <- all_compare_plot$Hugo_Symbol
names(Hugo_rownames) <- as.character(all_compare_plot$Row_Index)
all_p <- ggplot(all_compare_plot, 
               aes(x = or,
                   y = Row_Index,
                   color = cluster,
                   group = cluster)) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_pointrange(aes(xmin = ci.low, xmax = ci.up)) +
  scale_y_discrete(labels = Hugo_rownames) +
  scale_color_manual(values = cluster_colors) +
  theme_classic() +
  xlab("log odds ratio") +
  ylab("Gene") +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y", switch = "y") +
  theme(
    panel.grid.major.y = element_line(colour = "#f0f0f0"),
    axis.title.y = element_blank()
  )

tdat_all <- all_muts %>%
  filter(Hugo_Symbol %in% all_compare_plot$Hugo_Symbol) %>%
  distinct(Tumor_Sample_Barcode, Hugo_Symbol, clust_knn_char, type) %>%
  select(type, clust_knn_char, Hugo_Symbol) %>%
  table()
tdat_all <- left_join(
  as.data.frame(tdat_all),
  clinical_plus_cluster %>%
    select(Tumor_Sample_Barcode, type, clust_knn_char) %>%
    distinct() %>%
    group_by(type, clust_knn_char) %>%
    summarise(ntumors = n())) %>%
  group_by(type, clust_knn_char) %>%
  mutate(c_prop = Freq / ntumors) %>%
  mutate(c_prop = ifelse(is.na(c_prop), 0, c_prop)) %>%
  select(type, clust_knn_char, Hugo_Symbol, c_prop) %>%
  rename(cluster = clust_knn_char)
tdat_all <- all_compare_plot %>% left_join(tdat_all)
tdat_all_p <- tdat_all %>% ggplot(aes(type, Row_Index, fill = c_prop)) +
  geom_tile() +
  scale_fill_viridis_c(option = "E", "Proportion of \nTumor Type \nMutated Within Cluster", limits = c(0, 1)) +
  xlab("Tumor Type") +
  ylab("Gene Symbol") +
  scale_y_discrete(labels = Hugo_rownames) +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y") +
  theme_classic() +
  theme(
    strip.text = element_blank(),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(colour = "#f0f0f0")
  )

layout <- "ABB"
all_p + tdat_all_p + plot_layout(guides = 'collect', design = layout) & theme(legend.position = "bottom")
                   color = cluster,
                   group = cluster)) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_pointrange(aes(xmin = ci.low, xmax = ci.up)) +
  scale_y_discrete(labels = Hugo_rownames) +
  scale_color_manual(values = cluster_colors) +
  theme_classic() +
  xlab("log odds ratio") +
  ylab("Gene") +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y", switch = "y") +
  theme(
    panel.grid.major.y = element_line(colour = "#f0f0f0"),
    axis.title.y = element_blank()
  )

tdat_all <- all_muts %>%
  filter(Hugo_Symbol %in% all_compare_plot$Hugo_Symbol) %>%
  distinct(Tumor_Sample_Barcode, Hugo_Symbol, clust_knn_char, type) %>%
  select(type, clust_knn_char, Hugo_Symbol) %>%
  table()
tdat_all <- left_join(
  as.data.frame(tdat_all),
  clinical_plus_cluster %>%
    select(Tumor_Sample_Barcode, type, clust_knn_char) %>%
    distinct() %>%
    group_by(type, clust_knn_char) %>%
    summarise(ntumors = n())) %>%
  group_by(type, clust_knn_char) %>%
  mutate(c_prop = Freq / ntumors) %>%
  mutate(c_prop = ifelse(is.na(c_prop), 0, c_prop)) %>%
  select(type, clust_knn_char, Hugo_Symbol, c_prop) %>%
  rename(cluster = clust_knn_char)
tdat_all <- all_compare_plot %>% left_join(tdat_all)
tdat_all_p <- tdat_all %>% ggplot(aes(type, Row_Index, fill = c_prop)) +
  geom_tile() +
  scale_fill_viridis_c(option = "E", "Proportion of \nTumor Type \nMutated Within Cluster", limits = c(0, 1)) +
  xlab("Tumor Type") +
  ylab("Gene Symbol") +
  scale_y_discrete(labels = Hugo_rownames) +
  facet_grid(cluster ~ ., scales = "free_y", drop = TRUE, space = "free_y") +
  theme_classic() +
  theme(
    strip.text = element_blank(),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(colour = "#f0f0f0")
  )

layout <- "ABB"
all_p + tdat_all_p + plot_layout(guides = 'collect', design = layout) & theme(legend.position = "bottom")
```