### 2020-08-31 text/abstract.Rmd
incorporate Tim's suggestions for abstract

Cancer is an organism-level disease, impacting processes from cellular
metabolism and the microenvironment to systemic immune response. Nevertheless,
efforts to distinguish overarching mutational processes from interactions with
the cell of origin for a tumor have seen limited success, presenting a barrier
to individualized medicine. Here we present a novel, pathway-centric approach,
extracting somatic mutational profiles within and between tissues, largely
orthogonal to cell of origin, mutational burden, or stage. Known predisposition
variants are equally distributed among clusters, and largely independent of
molecular subtype. Prognosis and risk of death vary jointly by cancer type and
cluster. Analysis of metastatic tumors reveals that differences are largely
cluster-specific and complementary, implicating convergent mechanisms that
combine familiar driver genes with diverse low-frequency lesions in
tumor-promoting pathways, ultimately producing distinct molecular phenotypes.
The results shed new light on the interplay between organism-level dysfunction
and tissue-specific lesions.