### 2020-09-12 code/supplemental_v2.Rmd

---
output:
  bookdown::pdf_book:
    number_sections: FALSE
    toc: FALSE
    keep_tex: yes
always_allow_html: TRUE
editor_options:
  chunk_output_type: console
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi=600, warning = FALSE, message = FALSE)
suppressPackageStartupMessages(c(
  library(tidyverse),
  library(data.table),
  library(plot3D),
  library(flextable),
  library(knitr),
  library(kableExtra),
  library(xlsx)
))
```

\pagenumbering{gobble}

```{r supplementary-figure-s1, fig.width = 9, fig.height = 10}
umap_cancertype <- fread("../data/umap_3d_coors.tsv") ## load data
load("../data/kelly.colours.rda")
cluster.colours <- c(kelly.colours[c(3:12)]) ## Color scheme
names(cluster.colours) <- c(1:10) ## Assign cluster to color

cancer_type <- unique(umap_cancertype$project_code) ## Get cancer type in data set
cancer_type <- cancer_type[order(cancer_type)] ## Order alphabetically

par(mfrow = c(6,4), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(0,0,0,0))
for(n in cancer_type[1:23]) {
with(umap_cancertype %>% 
       filter(project_code == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F, axes = FALSE,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
legend("bottom", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2, adj = 0.25)
}
```

**Supplementary figure S1. Distribution of cancer type on UMAP clusters.** Each panel shows the identity of 
individual tumors by one of 23 cancer types projected onto the cluster UMAP identified 
in **Figure 1B of the main text**. The colors indicate the tumor's cluster identity.

\newpage

```{r supplementary-figure-s2-data-generation, fig.width = 9, fig.height = 10}
load("../data/ditto.colours.rda")

## Determining which cancer type has subtype information
cancer_w_subtypes <- list()
for(n in cancer_type) {
  spec_cancer <- umap_cancertype %>% 
    mutate(J_color = case_when(project_code == n ~ Subtype_Selected,
                               TRUE ~ "Other")) %>%
    filter(J_color != "Other")
  if(nrow(spec_cancer) == 0) {next
  }
  cancer_w_subtypes[[n]] <- n
}

cancer_w_subtypes <- names(cancer_w_subtypes)

par(mfrow = c(6,4), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(0,0,0,0))

for(n in cancer_w_subtypes[1:19]) {
  spec_cancer <- umap_cancertype %>% 
    mutate(J_color = case_when(project_code == n ~ Subtype_Selected,
                               TRUE ~ "Other")) %>%
    filter(J_color != "Other")
  if(nrow(spec_cancer) == 0) {next
    }
  subtypes <- unique(spec_cancer$J_color)
  J_colors <- c(ditto_colours[c(2:(length(subtypes) + 1))])#, 'grey')
  names(J_colors) <- subtypes

  with(spec_cancer,
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = J_colors[J_color], pch = 21, cex = 0.8, lwd = 0.2,
               theta = 0, phi = 65, scale = F, axes = FALSE,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
  legend("bottom", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2, adj = 0.25)
  legend("top", inset = c(0), legend=names(J_colors[1:2]), col=J_colors[1:2], pch = 20, bty = "n", horiz = T, cex = 1)
  legend("top", inset = c(0.08), legend=names(J_colors[3:4]), col=J_colors[3:4], pch = 20, bty = "n", horiz = T, cex = 1)
  legend("top", inset = c(0.16), legend=names(J_colors[5:6]), col=J_colors[5:6], pch = 20, bty = "n", horiz = T, cex = 1)
}
```

**Supplementary Figure S2. Histological subtypes.** Each panel shows the histological subtypes of one of the 23 
cancers projected onto the tumors of the UMAP from **Figure 1B of the main text**. The colors indicate the tumor's histological subtype identity.

\newpage

```{r supplementary-figure-s3-data-generation, eval = FALSE}
load("../data/gdc_tumor_coding_complete.Rdata")
load("../data/gene_data.Rda")

## Booleanize mutation count for each tumor sample
missense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$missense) 
splice <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$splice)
frameshdel <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshdel)
frameshins <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshins)
utr3 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr3)
utr5 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr5)
nonstop <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonstop)
nonsense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonsense)

gdc_mut_count_patient <- missense + splice + frameshdel + frameshins + utr3 + utr5 + nonstop + nonsense ## add up all the mutation count
rm(missense, splice, frameshdel, frameshins, utr3, utr5, nonstop, nonsense)
gc()
gdc_mut_count_patient[gdc_mut_count_patient > 0] <- 1 ## booleanize data, anything greater than 1 is equal to 1

## MMR genes
genes <- c("MSH2", "MSH6", "MLH1", "MLH3", "PMS1", "PMS2", "BRIP1", "RAD51", "CHEK2", "APC",
           "TET2", "TET3", "TERT", "TEP1", "DKC1")

selected_mut <- gene_data %>% dplyr::select(external_gene_name, ensembl_gene_id) %>% filter(external_gene_name %in% genes) ## getting ensembl_id for the mmr genes for filtering gdc data

selected_mut <- gdc_mut_count_patient %>% filter(rownames(.) %in% selected_mut$ensembl_gene_id) %>% ## filter by ensembl ids
  rownames_to_column(var = "ensembl_gene_id") %>% left_join(., selected_mut, by = "ensembl_gene_id") %>% ## replace ensembl ids with gene name 
  column_to_rownames(var = "external_gene_name") %>% dplyr::select(-ensembl_gene_id) %>%
  t() %>% as.data.frame() %>% rownames_to_column(var = "sample_id")  ## prep for left_join

selected_mut <- umap_cancertype %>% dplyr::select(sample_id, plot_x, plot_y, plot_z, clust_knn) %>% ## select relevant columns
  left_join(., selected_mut, by = "sample_id") %>%
  melt(data = ., id.vars = c("sample_id", "plot_x", "plot_y", "plot_z", "clust_knn")) %>%
  filter(value != 0)

save(selected_mut, file = "../data/selected_mut.Rda")
```

```{r supplementary-figure-s3, fig.width = 9, fig.height = 10}
load("../data/selected_mut.Rda")

genes <- c("MSH2", "MSH6", "MLH1", "MLH3", "PMS1", "PMS2", "BRIP1", "RAD51", "CHEK2", "APC", "TET2", "TET3", "TERT", "TEP1", "DKC1")

par(mfrow = c(6,4), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(0,0,0,0))
for(n in genes[1:15]) {
with(selected_mut %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F, axes = FALSE,
                                xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}
```

**Supplementary Figure S3. Distribution of mismatch repair genes on UMAP clusters.** Each panel shows one of 15 
genes involved in genome stability projected on the UMAP from **Figure 1B of the main text**. The colors indicate the tumor's cluster identity.

\newpage

```{r, supplemental-s4-data-generation, eval = FALSE}
load("../data/gdc_tumor_coding_complete.Rdata")
load("../data/gene_data.Rda")
umap_cancertype <- fread("../data/umap_3d_coors.tsv") ## load data

## Booleanize mutation count for each tumor sample
missense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$missense) 
splice <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$splice)
frameshdel <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshdel)
frameshins <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshins)
utr3 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr3)
utr5 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr5)
nonstop <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonstop)
nonsense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonsense)

gdc_mut_count_patient <- missense + splice + frameshdel + frameshins + utr3 + utr5 + nonstop + nonsense ## add up all the mutation count
rm(missense, splice, frameshdel, frameshins, utr3, utr5, nonstop, nonsense)
gc()
gdc_mut_count_patient[gdc_mut_count_patient > 0] <- 1 ## booleanize data, anything greater than 1 is equal to 1

## class 7 and 8 distinguishing genes
genes <- c("AKT2","BCL2","CTNNB1","ERBB2","KRAS","PIK3CA","CDKN2C","PTEN","RB1","SMARCA4","TP53","VHL")

selected_mut <- gene_data %>% dplyr::select(external_gene_name, ensembl_gene_id) %>% filter(external_gene_name %in% genes) ## getting ensembl_id for the genes for filtering gdc data

selected_mut <- gdc_mut_count_patient %>% filter(rownames(.) %in% selected_mut$ensembl_gene_id) %>% ## filter by ensembl ids
  rownames_to_column(var = "ensembl_gene_id") %>% left_join(., selected_mut, by = "ensembl_gene_id") %>% ## replace ensembl ids with gene name 
  column_to_rownames(var = "external_gene_name") %>% dplyr::select(-ensembl_gene_id) %>%
  t() %>% as.data.frame() %>% rownames_to_column(var = "sample_id")  ## prep for left_join

s5_data <- umap_cancertype %>% dplyr::select(sample_id, plot_x, plot_y, plot_z, clust_knn) %>% ## select relevant columns
  left_join(., selected_mut, by = "sample_id") %>%
  melt(data = ., id.vars = c("sample_id", "plot_x", "plot_y", "plot_z", "clust_knn")) %>%
  filter(value != 0)

save(s5_data, file = "../data/s5_data.Rda")
```

```{r supplemental-s4-plot, fig.width=9}
load("../data/s5_data.Rda")
genes <- c("AKT2","BCL2","CTNNB1","ERBB2","KRAS","PIK3CA","CDKN2C","PTEN","RB1","SMARCA4","TP53","VHL")

par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in genes[1:6]) {
with(s5_data %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}

par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in genes[7:12]) {
with(s5_data %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}

```

**Supplementary Figure S4. Distribution of oncogenes/tumor suppressors on UMAP clusters.** Each panel shows shows one of the twelve oncogenes/tumor suppressors projected onto the tumors of the UMAP from **Figure 1B of the main text**. The colors indicate the tumor's cluster identity.

\newpage

```{r supplemental-table-st1-data-generation, eval = FALSE}
load("../data/gdc_tumor_coding_complete.Rdata")
load("../data/gdc_reactome_path_analysis.Rda")
load("../data/gene2_level2_lookup.Rda")
load("../data/pathname2level2.Rda")
load("../data/gene_data.Rda")
load("../data/gdc_gene_exp_adj.Rda")

## Parameters ##
n_path <- 377 ## how many top pathways do you want to see. 1 = top path, 2 = top path and second most
mutated_gene_position <- c(1,2,3) ## what gene do you want to see. 1 = top gene, 2 = second most, 3 = third
show_perc <- TRUE ## show perc for mutated gene

toppath <- gdc_reactome_path_analysis %>% rowSums() %>% as.data.frame() %>% rownames_to_column(var = "path_name") %>% arrange(desc(.)) %>% .[1:n_path,] %>% .[, "path_name"] ## getting the top mutated pathways

id_to_names <- gene2level2_lookup %>% rownames_to_column(var = "ensembl_gene_id") %>% ## convert ensembl to gene name
  left_join(., gene_data[,c("ensembl_gene_id", "external_gene_name")], by = "ensembl_gene_id") %>%
  remove_rownames() %>% column_to_rownames(var = "external_gene_name") %>%
  dplyr::select(-ensembl_gene_id) %>%
  t() %>% as.data.frame() %>% rownames_to_column(var = "PATH_ID") %>% ## convert path_id to path_name
  left_join(., pathname2level2, by = "PATH_ID") %>%
  remove_rownames() %>% column_to_rownames(var = "PATH_NAME") %>%
  dplyr::select(-PATH_ID) %>%
  t() %>% as.data.frame()

gene_mutation <- gdc_gene_exp_adj %>% rownames_to_column(var = "ensembl_gene_id") %>% ## convert ensembl to gene name
  left_join(., gene_data[,c("ensembl_gene_id", "external_gene_name")], by = "ensembl_gene_id") %>%
  remove_rownames() %>% column_to_rownames(var = "external_gene_name") %>%
  dplyr::select(-ensembl_gene_id)

table_st1 <- list() ## list for forloop

for(n in toppath) {
  message(n)
  ## finding top mutated gene in pathway
  gene_in_path <- id_to_names %>% .[,which(names(.) == n)] %>% as.data.frame() %>% ## select pathway
    'rownames<-' (rownames(id_to_names)) %>% 
    filter(. == 1) %>% ## filter for genes that are in that pathway
    rownames(.)
  top_mut_gene <- gene_mutation %>% .[rownames(.) %in% gene_in_path, ] %>% ## only include genes in paths
      rowSums() %>% as.data.frame() %>% 'colnames<-' ("obs_count") %>% ## sum the pathway mutation
    rownames_to_column(var = "gene") %>%
      mutate(n_sample = length(gene_mutation),
             perc_mut = round((obs_count / length(gene_mutation)) * 100, digits = 1)) %>% 
       arrange(desc(obs_count))
  top_gene_in_path <- list()
  for(p in mutated_gene_position) {
    message(p)
    if(show_perc == TRUE) {
    top_gene_in_path[[p]] <- paste0(top_mut_gene$gene[p], " (", top_mut_gene$perc_mut[p],"%)")  
    }
    if(show_perc == FALSE) {
    top_gene_in_path[[p]] <- paste0(top_mut_gene$gene[p])  
    }
  }
  top_gene_in_path <- top_gene_in_path %>% do.call(rbind, .) %>% as.data.frame() %>% ## rbind list
   'rownames<-' (mutated_gene_position) %>% t() %>% as.data.frame()
  ## calculating the percent mutated samples
  spec_path <- data.frame(Pathway = n,
                          Percent_Mutated = round(gdc_reactome_path_analysis %>% .[n,] %>% sum() / length(gdc_reactome_path_analysis) * 100, digits = 1))
  table_st1[[n]] <- cbind(spec_path, top_gene_in_path)
}
table_st1 <- as.data.frame(do.call(rbind, table_st1))
save(table_st1, file = "../data/table_st1.Rda")  
rm(list = ls(all.names = T))  
gc()
```

```{r supplemental-table-st1, eval = FALSE}
load("../data/table_st1.Rda")
## for html only
# table_st1 %>% dplyr::select(-Pathway) %>% knitr::kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 12) %>%
#   column_spec(1, bold = T, color = "black")

## for word doc
# table_st1 %>% flextable() %>% 
#   set_caption(caption = "") %>%
#   align(part = "all") %>% 
#   width(., width = c(4,0.5,0.5,0.5,0.5)) %>% 
#   fontsize(size = 8, part = "body")
```

```{r st2, fig.width= 9, eval = FALSE}
##**Supplementary Table 2. Cluster-specific enriched pathways (effect size >= 0.30).**
## docx friendly
load("../data/moderate_enrich_prob.Rda")

table_st2 <- moderate_enrich_prob %>% filter(difference >= 0.30 & signif <= 2500) %>% ## filter effect size >= 0.20 and enrichment at a 95% CI
  dplyr::select(path, clust, signif, difference) %>% 'colnames<-' (c("Pathway", "Cluster", "Signif", "Enrichment_Score")) ## select relevant column and rename col

table_st2$log10pvalue <- round(-log10((table_st2$Signif + 1) / 1E5), digits = 2) ## Getting p value ## adding 1 for pseudo count.. 1E5 is the n simulations performed
table_st2$Enrichment_Score <- round(table_st2$Enrichment_Score, digits = 2)
table_st2$Signif <- NULL ## we don't need this anymore

# st1 <- st1 %>% flextable() %>% 
#   set_caption(caption = "") %>%
#   align(part = "all") %>% 
#   width(., width = c(4,0.75,1,1)) %>% 
#   fontsize(size = 10, part = "body")
#

# names(st1) <- c("Pathway", "Cluster", "Effect Size", "-log(p)")
#  st1 %>%  knitr::kable(format = "latex", longtable = TRUE, align = c("l", rep("c",3))) %>%
#    kable_styling(latex_options=c("scale_down", "repeat_header"), full_width = FALSE, font_size = 12) %>%
#    column_spec(1, color = "black", width = "4in")

## Saving st1 and st2 in excel

write.xlsx(table_st1, file="supplemental.table.xlsx", sheetName="ST1", row.names=FALSE)
write.xlsx(table_st2, file="supplemental.table.xlsx", sheetName="ST2", append=TRUE, row.names=FALSE)
```