### 2020-09-12 text/survival.Rmd

If our clusters represent biological states distinct from tissue of origin, they
may have different prognoses within cancer types or across all cancers. These
analyses are limited by confounding factors of age, stage at diagnosis, sex,
ethnicity, and tissue-specific disease progression. To explore these ideas, we
used Bayesian inference to test models of survival using public longevity data
from the CDC. The estimates the effect of cancer type and cluster-specific
cancer effects independently, resulting in a cancer and cluster-specific
estimate of the aging rate multiplier, $k$.

To illustrate the model, we simulate survival four related groups (**figure
5A**). We start with a cohort of women aged 30 and show expected survival (light
blue). By setting cancer rate multiplier to $k=2$ we simulate the effect of a
moderately deadly cancer (yellow). Compare this to randomly selected women
ranging from 30 to 70 years old (dark blue). Immediately, survival changes due
to the mixture of ages, without malignancy. Adding a $k=2$ malignancy further
reduces expected lifespan (red). Hence, due to the change in distribution of
age-at-diagnosis, we expect equally deadly cancers to have *different* survival
curves. The model takes this into account and cancer aggressiveness is estimated
accurately without the confounding effect of age.

```{r, results='asis'}
cat("\\newpage")
```

#### Figure 5
```{r child='../code/survival/figure5.Rmd'}
```

**Figure 5: Comparison of different cancers, simulated and actual results.**
Comparison of survival in four simulated cohorts. A) A group consisting of
randomly selected 30 year old women, unaffected (light blue) _vs._ cancer, 
($k = 2$, light orange), another group consisting of random 30-70 year-old
women, unaffecteds (dark blue) and cancer $k = 2$ (dark orange). B,C,D,E)
Compare random group with correct age distribution for the given cancer
(turqoise) to the model predicted survival for the given average k (blue) and
the actual survival for the patients in our dataset (orange). Differences with
actual data are likely due to a mixture of cluster identities which we exclude
from these simulations for simplicity. F) Effects are compared between cancer
type by showing the probability density for the posterior value of the $k_{tis}$
multiplier. Larger values correspond to decreased age-independent life
expectancy.

```{r, results='asis'}
cat("\\newpage")
```

We found that cancer types, as expected, have a range of prognoses relative to
the general population. In **figure 5B-E** we see three particularly deadly
cancers (Stomach:STAD, Glioblastoma:GBM, and Pancreatic:PAAD), and one cancer
where diagnosis decreases risk relative to background (Thyroid: THCA). Cancers
with posterior probability for relative risk of less than 1 such as THCA should
be interpreted carefully. This Bayesian model is for a state of information. The
information that a person is diagnosed with cancer may lead us to expect shorter
survival than the general population of matched age (k > 1), or longer survival
than the general population of matched age (k < 1). Shorter prognosis could
result from cancer aggressiveness, injuring the body and causing death. By
contrast, longer prognosis could result from the cancer being relatively mild,
and therefore diagnosis could be an informational signal that the patient is
health conscious, with the comparison group having more people whose cancers go
undiagnosed. It is important to note therefore that the diagnosis can increase
our expectation of life relative to the comparison group, even if it decreases
the expectation of life of the individual relative to the counterfactual where
they did not have cancer.

The degree to which cancer accelerates aging can be determined by the $k_{tis}$
multiplier (**figure 5F**) multiplied by the cluster modifier $k_{cl}$ (**figure
6**). Looking at the tissue specific multiplier, the least deadly was prostate
cancer (PRAD), and the most deadly was glioblastoma (GBM) which has a risk
multiplier of between 15 and 22 relative to the background risk of death in the
population. This undoubtedly is influenced by the fact that relatively young
patients are affected by GBM and that it is extremely deadly. Among the
deadliest cancers outside of GBM were stomach (STAD), melanoma (SKCM) and
pancreatic (PAAD) cancers. Apart from these top 4 cancers which ranged from 6-22
in relative risk of age-adjusted death, the remaining cancer types ranged from
about 1 to 5 in magnitude.

```{r, results='asis'}
cat("\\newpage")
```

#### Figure 6

```{r child='../code/figure6.Rmd'}
```

**Figure 6 Cluster specific $k_{cl}$ values for each cancer type.** The overall
$k$ value is the product of $k_{tis} \times k_{cl}$. The cluster-specific $k$
represents relative aggressiveness of each cluster within cancer type. Alpha
transparency (α) is set to reflect varying confidence in the posterior
distribution when the number of observations is N<10, α = 0.1; N<25, α = 0.5; N
≥ 25 α = 1.0.

```{r, results='asis'}
cat("\\newpage")
```

Our estimates of tissue-specific cluster effects, in contrast, ranged from less
than 1 up to about 3 or 4, reflecting that some clusters are either less deadly
or more deadly than other clusters within each cancer type (**figure 6**). A
cluster-specific rate of 1 represents the typical rate for this tissue type. For
several cancers (_e.g._ PRAD, kidney chromophobe (KICH), diffuse large B-cell
lymphoma (DBLC), thyroid cancer (THCA)) the posterior estimates are largely
indistinguishable from the prior, reflecting that either there were too few
mortalities in the data to make an estimate (as expected for PRAD and THCA) or
two few samples, period. We did not observe cluster-specific trends that held
true across cancer types, which could result from different cancers having
different standards of care for example. In support of this interpretation, we
also tested a factored model which considered both cancer type cluster
independently. Though we were able to successfully fit this model, it is a
special case of the more general model where cluster and tissue independently
affect longevity, and there is no reason to believe that cluster specific
effects would necessarily remain constant across tissue types given how widely
the patients vary across tissue types, drug and surgical treatments.