### 2021-06-29 code/survival/figure5.Rmd

```{r, survfig, echo=FALSE, warning=FALSE, error=TRUE,message=FALSE,fig.height=11, fig.width=9}

library(ggridges)
library(patchwork)

source("./kmcurve.R")

## create example data for a fixed 30 year old without vs with cancer (effective age)
## cancer

## Effective age calc for typical BRCA
effage0tiss <- 62
effage0 <- 1.1
effagert <- 0.35

patientsurvivalsbkd = replicate(1000,simdeath(r20wf,ageratewf,30,1))/365 ## Typical survival for 30yr with k = 1
## Stimulating survival for white female with cancer - effective age for 30 yr old
effage <- effage0tiss * effage0 + effagert * (30 - 20)
patientsurvivalseffk1 <- replicate(1000,simdeath(r20wf,ageratewf,effage,1))/365 ## Typical survival for 30yr with effage due to cancer k=1
patientsurvivalseff10k1 <- replicate(1000,simdeath(r20wf,ageratewf,effage+10,1))/365 ## effage + 10 but still k = 1
## Mixed ages
mixedages = runif(1000,30,70)
mixedsurvwfk1 = sapply(mixedages, function(x) simdeath(r20wf,ageratewf,x,1)/365) ## Typical survival for mixed ages
## Stimulating survival for white female with cancer - mixed age
effage.mixed <- effage0tiss * effage0 + effagert * (mixedages - 20)
mixedsurvwfeffk1 = sapply(effage.mixed, function(x) simdeath(r20wf,ageratewf,x,1)/365) ## Typical survival for mixed ages with effage due to k=1
mixedsurvwfeff10k1 <- sapply(effage.mixed+10, function(x) simdeath(r20wf,ageratewf,x,1)/365)

simdata = data.frame(
  r = patientsurvivalsbkd,
  ce = patientsurvivalseffk1,
  ce10 = patientsurvivalseff10k1,
  m = mixedsurvwfk1,
  me = mixedsurvwfeffk1,
  me10 = mixedsurvwfeff10k1,
  status = rep(1, length(patientsurvivalsbkd))
) %>%
  rownames_to_column(var = "index") %>%
  pivot_longer(-c(index, status), values_to = "time", names_to = "condition")
p1 <- ggplot(simdata, aes(time = time, color = condition, status = status)) +
  geom_km() +
  labs(x = "Survival Time (years past diagnosis)", y = "Probability") +
  scale_color_manual(
    values = c(r = "#56b4e9", ce = "#e69f00", ce10 = "#0072b2", m = "#009E73", me = "#CC79A7", me10 = "#D55E00")
  ) +
  theme_classic() +
  annotate("text", x = 44, y = .75, label = "30yo women", color = "#56b4e9", size = 3) +
  annotate("text", x = 27, y = .70, label = "30yo women cancer BRCA typ", color = "#e69f00", size = 3) +
  annotate("text", x = 25, y = .65, label = "30yo women cancer BRCA+10yr", color = "#0072b2", size = 3) +
  annotate("text", x = 42, y = .45, label = "Mixed age women", color = "#009E73", size = 3) +
  annotate("text", x = 29, y = .40, label = "Mixed age women BRCA typ", color = "#CC79A7", size = 3) +
  annotate("text", x = 27, y = .35, label = "Mixed age women BRCA+10yr", color = "#D55E00", size = 3) +
  theme(legend.position = "none")

load("./survmodel_effcl_knn_062821.stansave")
cancerdat = read_csv("../../data/survival/clinical_plus_cluster.csv");
cancerdat$type <- as.factor(cancerdat$type)
tisnames <- levels(cancerdat$type)
## tissue specific effective age

effage0tiss <- as.data.frame(samps) %>% select(starts_with("effage0tiss[")) %>%
  gather(data = ., key = "coefname", value = "effage0tiss") %>%
  mutate(tiss = tisnames[as.numeric(gsub("effage0tiss\\[(.*)\\]","\\1",.$coefname))]) ## Replace with tumor type

p2 <- ggplot(effage0tiss,aes(x = effage0tiss, y = tiss)) + geom_density_ridges() +
  theme_minimal() +
  scale_x_continuous(breaks=seq(20,100, by = 20), limits = c(20,100)) +
  labs(x = "Effective Age", y = "Cancer Type") +
  theme(legend.position = "none") +
  geom_vline(aes(xintercept=20), linetype = "dashed", color = "red")

## tissue specific effective age rate

effagert <- as.data.frame(samps) %>% select(starts_with("effagert")) %>%
  gather(data = ., key = "coefname", value = "effagert") %>%
  mutate(tiss = tisnames[as.numeric(gsub("effagert\\[(.*)\\]","\\1",.$coefname))]) ## Replace with tumor type

p3 <- ggplot(effagert,aes(x = effagert, y = tiss)) + geom_density_ridges() +
  theme_minimal() +
  labs(x = "Effective Age Rate", y = "") +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        plot.title= element_text(hjust = 0.5)) +
  geom_vline(aes(xintercept=1), linetype = "dashed", color = "red")

ktis <- as.data.frame(samps) %>% select(starts_with("ktis")) %>%
  gather(data = ., key = "coefname", value = "ktis") %>%
  mutate(tiss = tisnames[as.numeric(gsub("ktis\\[(.*)\\]","\\1",.$coefname))]) ## Replace with tumor type

p4 <- ggplot(ktis,aes(x = ktis, y = tiss)) + geom_density_ridges() +
  theme_minimal() +
  scale_x_continuous(breaks=seq(-1,3, by = 1), limits = c(-0.25,3)) +
  labs(x = expression(paste("k"["tis"])), y = "") +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        plot.title= element_text(hjust = 0.5)) +
  geom_vline(aes(xintercept=1), linetype = "dashed", color = "red")

## Plot survival by cancer type
asamps <- as.array(samps)

endpoints = sqldf("select case when death_days_to is NULL then last_contact_days_to else death_days_to end as finaltime, case when death_days_to is NULL then 0 else 1 end as finalstatus from cancerdat;")

cancerdat = cbind(cancerdat,endpoints);

cancerdat = cancerdat[!is.na(cancerdat$finaltime),]

cancertype <- c("GBM", "PAAD", "OV", "THCA")
for(ct in cancertype) {
  lev <- which(levels(cancerdat$type) == ct)[[1]] ## level of cancer type
  women <- cancerdat[cancerdat$type == ct & cancerdat$gender == "FEMALE",]
  agesw <- women$age

  ## Mean of various parameters
  k = round(mean(asamps[,,sprintf("ktis[%d]",lev)]), digits = 2)
  effage0tiss <- round(mean(asamps[,,sprintf("effage0tiss[%d]",lev)]), digits = 0)
  effagert <- round(mean(asamps[,,sprintf("effagert[%d]",lev)]), digits = 2)
  effage0 <- round(median(asamps[,,sprintf("effage0[%d,%d]",lev,1:10)]), digits = 2)

  ## Simulation of women no cancer
  survyrswfref = sapply(agesw,function(a) tryCatch(simdeath(r20wf,ageratewf,a,1)/365,error= function(e) {0}))
  ## Effective age due to cancer
  effage.agesw <- effage0tiss * effage0 + effagert * (agesw - 20)
  survyrswf = sapply(effage.agesw,function(a) tryCatch(simdeath(r20wf,ageratewf,a,kmean)/365,error= function(e) {0}))

  df.plot <- plotkmcomp(survyrswf, survyrswfref, women$finaltime / 365, women$finalstatus) +
          coord_cartesian(xlim = c(0, 80)) +
          labs(
              x = "Survival Time (years past diagnosis)",
              y = "Probability"
          ) +
          scale_color_manual(
              values = c(
                  rtime = "#66c2a5",
                  stime = "#8da0cb",
                  acttime = "#fc8d62"
              ),
              breaks = c("rtime", "stime", "acttime"),
              labels = c(
                  "Simulated General Population; k = 1",
                  "Simulated Patient Population",
                  "Actual Patient Population"
              )
          ) +
    annotate(geom = "text", x = 50, y = 1.0, label = paste0(ct, " simulated")) +
    theme(legend.position = "none")
  if(ct == "GBM") {
    df.plot <- df.plot + annotate(geom = "text", x = 40, y = 0.7, label = "Simulated General Population; k = 1", color = "#66c2a5", size = 3) +
      annotate(geom = "text", x = 32, y = 0.5, label = "Simulated Patient Population", color = "#8da0cb", size = 3) +
      annotate(geom = "text", x = 27, y = 0.3, label = "Actual Patient Population", color = "#fc8d62", size = 3)
  }
  assign(ct, df.plot, envir = .GlobalEnv)
}

layout = "
AAAEE
AAAEE
AAAEE
AAAFF
AAAFF
AAAFF
BCDGG
BCDGG
BCDGG
BCDHH
BCDHH
BCDHH
"

p1 + p2 + p3 + p4 + GBM + PAAD + OV + THCA + plot_spacer() + plot_annotation(tag_levels = "A") + plot_layout(design = layout)

```