### 2021-07-01 text/introduction.Rmd

Advances in technology have greatly expanded our view into the mechanisms
of cancer at the systems biology level. Next-generation sequencing technologies
have made it possible to study germline and somatic mutations, expression
profiles, DNA methylation, and copy number variations from the same tissue. To
take advantage of these tools, several large consortia, including the cancer
genome atlas (TCGA) and pan-cancer analysis of whole genomes (PCAWG) sequenced
large numbers of tumors and collected data from multiple assays to analyze together, with the goal of
integration and increased understanding of the mechanisms of cancer.

Considerable progress has been made analyzing these data. Statistical analyses
identify hundreds of global and tissue-specific cancer driver genes
[@deesMuSiC2012; @tamboreroOncodriveCLUST2013;
@lawrenceDiscoverySaturationAnalysis2014;
@kumarStatisticallyIdentifyingTumor2015; @tokheimEvaluatingEvaluationCancer2016;
@jiangWITER2019; @zhaoModelingPositive2019] using approaches aimed at detecting
genes mutated at a greater rate than expected due to chance. Dees et al. [@deesMuSiC2012]
developed a pipline to separate driver mutations from passenger mutations. 
Tamborero et al. [@tamboreroOncodriveCLUST2013] expanded on this idea to identify 
tumor mutation biases towards specific regions of a protein sequence. Lawrence et 
al. [@lawrenceDiscoverySaturationAnalysis2014] focused on identifying cancer driver 
genes with intermediate mutation frequencies using tumor-normal pairs. It has been
estimated that fewer than five mutations in key oncogenes and/or tumor
suppressors would be sufficient to transform a normal cell to a cancerous state
[@vogelsteinPathCancerThree2015; @Iranzo-Network]. Mutations in cancer driver
genes are commonplace in healthy tissue and correlate with age and environmental
exposures [@mutations-healthy-tissues].

Other studies provide a comprehensive view of mutations, gene expression and
genomic signatures, with the goal of understanding common themes of all cancer
independent of tissue of origin. Understanding cancer as a disease of the cell
has long been a goal of the field as characterized in essays by
[@weinberg2000; @hallmarks_II]. The first of these studies in genomics identified
11 subtypes from 12 cancer types, using integrative analysis with co-equal
weighting of gene expression, methylation, copy number and proteomics data
[@hoadleyMultiplatformAnalysis2014]. The principal finding was that
tissue-of-origin is the predominant driving factor, though ~10% could be
reclassified independent of tissue-of-origin. In a second study involving 33
cancer types and a much greater number of tumors, the authors identified 28
clusters that could be further subdivided into organ specific groups, including
pan-gastrointestinal, pan-gynecological, pan-squamous,
pan-gynecological/squamous and pan-kidney
[@hoadleyCellofOriginPatternsDominate2018].

More recently, pan-cancer mutations and other data have been looked at as
disruptions of normal pathway activity. Pathway-centric analysis of tumors
provides additional benefits over gene-centric analysis: (1) It is significantly 
less noisy because it aggregates molecular events across multiple genes in the 
same pathway and (2) the identification of a potential casual mechanism is easier 
to interpret as genetic aberrations are linked to molecular pathways. In a perspective, 
Creixell et al. [@Creixell-pathwaynetworkanalysis] described approaches of pathway 
and network analysis applicable to next-generation sequencing data. The most common 
approach being fixed-gene enrichment analysis to identify over-represented pathways. 
Sanchez-Vega et al. [@PCAWG-PATH] carried out an exhaustive
analysis of pathway enrichment using whole genome data. Focusing on ten frequently 
altered canonical signaling pathways in cancer, they explored the mutation frequencies 
of coding regions of genes in these pathways for each cancer type. Network approaches
address the sparseness of mutations, allowing genes to be influenced by
mutations in nearest network neighbors [@hotnet2;@hornNetSigNetworkbasedDiscovery2018;
@Iranzo-Network]. These methods are powerful especially when applied to smaller
numbers of tumors. One of the challenges in the field remains that there may be
thousands of moderate effect genes that occur at such low frequency that they
are impossible to detect using positive selection theory. Some researchers have
attempted to address this challenge by applying machine learning to cancer data
to discover groups of functionally related genes as they interact with larger
pathways and networks [@kimPancancerAnalysisSomatic2018;
@mourikis-ml-pathways; @moonlight]. These studies have proven very effective at
highlighting fundamental disease phenotypes at the pathway level across cancers
with different origins at the cellular and tissue level.

In this study, we attempt to understand how tissue-specific gene disruptions
create common cancer phenotypes by focusing on discrete molecular pathways as
the unit of disruption. Our approach strips all cell-type-specific information
from the mutation data and equates gene-level mutations to cell-biological
pathway disruptions. We use this heuristic to evaluate all cancers and show
that, surprisingly, tumors that exhibit tissue-specific gene mutation patterns
nonetheless fall into common categories of pathway disruption having unique
prognoses in each cancer type.