### 2021-07-02 text/survival.Rmd
Minor edits to text

If our clusters represent biological states distinct from tissue of origin, they
may have different prognoses within cancer types or across all cancers. These
analyses are limited by confounding factors of age, stage at diagnosis, sex,
ethnicity, and tissue-specific disease progression. To explore these ideas, we
used Bayesian inference to test models of survival using public longevity data
from the CDC. The model estimates the effect of cancer type and cluster-specific
cancer effects independently, resulting in a cancer and cluster-specific
estimate of an effective age function and the aging rate multiplier, $k$.

To illustrate the model, we simulate survival for six related groups (**Fig
5A**). We start with a cohort of women aged 30 randomly selected from the US
population, and show expected survival (light blue). We then use typical values
from the posterior samples to set the effective age of that cohort to the
effective age of a BRCA patient (yellow). Finally to illustrate how the model
varies by effective age we add an additional artificial 10 years to the BRCA
effective age (dark blue). Compare these results to a uniformly selected group of
women ranging from 30 to 70 years old at random (green). Immediately, survival changes due
to the mixture of ages, without malignancy. By computing the effective ages for
this mixture of patients, we see that malignancy further reduces expected
lifespan (pink). Adding an additional 10 years to effective age would reduce the
lifespan even further (dark orange). Hence, simply due to the change in
distribution of age-at-diagnosis, we expect two equally deadly cancers to have
*different* survival curves. The model takes this into account and cancer
aggressiveness is estimated accurately without the confounding effect of age
through the effective age and aging rate parameters.

We found that cancer types, as expected, have a range of prognoses relative to
the general population. In **Fig 5B-D** we show how effective age baseline, and
effective age rate (which modifies the baseline based on your actual age) as
well as the $k_{tis}$ which accelerates or decelerates further aging, varies
considerably from cancer to cancer. For example a 20 year old diagnosed with
THCA or PRAD would be treated similarly to a 40 year old based on the Effective
Age parameter. It's noticeable that for every year beyond 20 in the PRAD
diagnosis we add only an extra ~ 0.2 years (Effective Age Rate). The $k_{tis}$
parameter represents an acceleration (greater than 1) or deceleration (less) of
the further aging beyond diagnosis. For example, for BRCA patients each year past
diagnosis adds about 1 year of risk, whereas for LGG patients it adds around 2,
and for SARC patients it adds only 0.25 or so.

In **Fig 5E-H** we see three particularly deadly cancers (Glioblastoma:GBM,
Pancreatic:PAAD, Ovarian:OV), and one cancer where diagnosis hardly changes risk
relative to background (Thyroid: THCA). Cancers with posterior probability for
relative risk of less than 1 should be interpreted carefully. This Bayesian
model is for a state of information. The information that a person is diagnosed
with cancer may lead us to expect shorter survival than the general population
of matched age, or longer survival than the general population of matched
age. Shorter prognosis could result from cancer aggressiveness, injuring the
body and causing death. By contrast, longer prognosis could result from the
cancer being relatively mild, and therefore diagnosis could be an informational
signal that the patient is health conscious, with the comparison group having
more people whose cancers and other health issues go undiagnosed. It is
important to note therefore that the diagnosis can increase our expectation of
life relative to the comparison group, even if it decreases the expectation of
life of the individual relative to the counterfactual where they did not have
cancer.

Our estimates of tissue-specific cluster effects were for the Effective Age
parameter. These cluster effects were multiplied by the tissue specific
coefficient to form the full effective age. Priors for this parameter were
relatively peaked around 1 reflecting the purpose of this coefficient as a
multiplicative perturbation (**Fig 6**). A cluster-specific rate of 1 represents
the typical rate for this tissue type. For several cancers (_e.g._  PRAD, kidney
chromophobe (KICH), diffuse large B-cell lymphoma (DLBC), thyroid cancer (THCA))
the posterior estimates are largely indistinguishable from the prior, reflecting
that either there were too few mortalities in the data to make an estimate (as
expected for PRAD and THCA) or too few samples, period. We did not observe
cluster-specific trends that held true across cancer types, which could result
from different cancers having different standards of care for example. For some
cancers however, there is evidence that certain forms are more or less
aggressive. PAAD class 8 seems to raise your effective age somewhat relative to
other forms, as does LGG group 7, whereas BRCA group 5, COAD group 8, and LGG
group 5 all seem somewhat lower risk than the other classes within those
tissues.

To compare our pathway-based clustering to other published methods, we evaluated
the performance in predicting time to death (tdeath) with other published
classification models. We downloaded cluster membership from two studies
conducted within the TCGA consortium [@hoadleyMultiplatformAnalysis2014;
@hoadleyCellofOriginPatternsDominate2018].  We generated tdeath predictions from
uniform random p, by inverting the cumulative distribution function of our death
probability function. Using the actual tdeath from TCGA as the reference, we
show that our pathway-based clustering (referred to as kNN) had slightly better
performance compared to COCA [@hoadleyMultiplatformAnalysis2014], iCluster
[@hoadleyCellofOriginPatternsDominate2018] and tissue-of-origin (tis). (**S5A
Fig**). Additionally, we computed the entropy of the error distribution in
predictions (**S5B Fig**).  All models had similar performance with kNN
performing slightly better as indicated by a slightly smaller error entropy.