### 2021-07-20 code/figure6.Rmd
Edit manuscript based on editor comments

```{r, message = FALSE, fig.width = 9, fig.height= 9}
load("../data/kelly.colours.rda")
cluster.colours <- c("grey", kelly.colours[c(3:12)])
names(cluster.colours) <- c(0:10)

## Getting n_race for white and black
cancerinfo <- read_csv("../data/survival/clinical_plus_cluster.csv")
cancer.clusttype <- cancerinfo %>% dplyr::select(type, race, clust_knn) ## select relevant columns
n_cancer <- as.data.frame(table(cancer.clusttype$type)) ## table cancer type
n_cancer$facetname <- paste0(n_cancer$Var1, " (N = ", n_cancer$Freq, (")")) ## Create a new col with type and freq together
n_cancer$Freq <- NULL ## Remove Freq Column
names(n_cancer) <- c("tiss", "facetname") ## rename cols for left join in kclassplot below

## Getting n_cluster within cancer.type to calculate alpha
cancer.nclustintype <- as.data.frame(table(cancer.clusttype$type, cancer.clusttype$clust_knn)) ## Determine n_cancer in each cluster
cancer.nclustintype <- cancer.nclustintype %>% mutate(alpha = case_when(Freq < 10 ~ 0.1, ## Create alpha column for geom_ridges
                                                                        Freq >= 10 & Freq < 25 ~ 0.5,
                                                                        Freq >= 25 ~ 1.0),
                                                      alphatext = case_when(Freq < 10 ~ "0.1", ## Create alpha text for labeling
                                                                        Freq >= 10 & Freq < 25 ~ "0.5",
                                                                        Freq >= 25 ~ "1.0"))

cancer.alpha <- cancer.nclustintype %>% dplyr::select(Var1, Var2, alphatext) %>%
  'colnames<-' (c("tiss", "class","alphatext"))

tisnames <- levels(cancerdat$type) ## getting cancer type names

# Code to generate the text label for each alpha
# textlist <- list() ## create list for for loop
# for(n in tisnames) {
#   df <- cancer.nclustintype %>% filter(Var1 == n) ## Filter for specific cancer type
#   textlist[[n]] <- as.character(paste(df[1, "alphatext"], "\n",df[2, "alphatext"], "\n", df[3, "alphatext"], "\n",df[4, "alphatext"], "\n",df[5, "alphatext"], "\n",df[6, "alphatext"], "\n",df[7, "alphatext"], "\n",df[8, "alphatext"], "\n",df[9, "alphatext"], "\n",df[10, "alphatext"])) ## collect all the alpha values and insert new line after each alpha
# }
# textlist <- as.data.frame(do.call(rbind, textlist)) ## rbind list
# textlist <- textlist %>% 'colnames<-' ("lab") %>% ## rename col to lab
#   rownames_to_column(var = "tiss") ## cancer type to column for left join

## Processing data
load("./survival/survmodel_effcl_knn_070121.stansave")
dfsamps <- as.data.frame(samps)
clnames = grep("effage0\\[",names(dfsamps), value=TRUE)
dfclas = dfsamps[,clnames]
dfclas = stack(dfclas)
names(dfclas) = c("value","coefname")
dfclas$class = as.factor(as.numeric(gsub("effage0\\[(.*),(.*)\\]","\\2",dfclas$coefname)))
dfclas$tiss = tisnames[as.numeric(gsub("effage0\\[(.*),(.*)\\]","\\1",dfclas$coefname))]
dfclas$class <- fct_rev(dfclas$class)

## Left joining data created above
dfclas <- left_join(dfclas, n_cancer, by = "tiss") ## add facet names

# ## Code for adding alpha level text
# 	dfclas <- left_join(dfclas, textlist, by = "tiss")
# 	for(n in tisnames) {
# 	  dfclas[which(dfclas$tiss == n)[-1], "lab"] <- NA ## Remove the extra 29999 labels that will be generated on the ggplot
# 	}

## Code for setting alpha in geom_density ridges
alpha.setup <- c() ## create empty vector for loop
	for(n in tisnames) { ##loop through the cancer type in alphabetical order
	  df <- cancer.nclustintype %>% filter(Var1 == n) %>% ## Filter for specific cancer type	
	    arrange(desc(Var2)) ## Arrange in desc order cause thats how the alpha are set from 10 <- 1
	  for(p in 1:nrow(df)) { ## loop through the clusters starting from 10 --> 1
	    alpha.setup <- c(alpha.setup, rep(df[p, "alpha"], 512)) ## Get the alpha value for each cluster in each cancer type and rep 512 (necessary for geom ridges)
	  }
	}

normal_lines <- data.frame(facetname = unique(dfclas$facetname),
                         x0 = 1) ## Add dashed line at x = 1

	## Plotting
	ggplot(dfclas,aes_(x = quote(value),y = quote(class), fill = quote(class))) + geom_density_ridges(alpha = alpha.setup) +
		facet_wrap(. ~ facetname,nrow=3) + theme_minimal() + scale_fill_brewer(palette = "Paired") +
	  theme(legend.position = "none") +
	  labs(x = expression(paste("A"["cl"])), y = "Cluster Identity") +
	  theme(plot.title= element_text(hjust = 0.5)) +
	  scale_fill_manual(values = cluster.colours) +
	  scale_x_continuous(breaks=seq(0.5,1.5, by = 0.5), limits = c(0.5,1.5)) +
    geom_vline(aes(xintercept=1), linetype = "dashed", color = "red")
```