### 2021-07-20 code/supplemental_v2.Rmd

---
output:
  word_document:
    reference_docx: "../text/word-style-reference-ttmanu-iscb-iscience-v2.docx"
always_allow_html: TRUE
editor_options:
  chunk_output_type: console
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi=600, warning = FALSE, message = FALSE)
suppressPackageStartupMessages(c(
  library(tidyverse),
  library(data.table),
  library(plot3D),
  library(flextable),
  library(knitr),
  library(kableExtra),
  library(xlsx),
  library(patchwork)
))
```

\pagenumbering{gobble}

```{r supplementary-figure-s1, fig.width = 9, fig.height = 10}
umap_cancertype <- fread("../data/umap_3d_coors.tsv") ## load data
load("../data/kelly.colours.rda")
cluster.colours <- c(kelly.colours[c(3:12)]) ## Color scheme
names(cluster.colours) <- c(1:10) ## Assign cluster to color

cancer_type <- unique(umap_cancertype$project_code) ## Get cancer type in data set
cancer_type <- cancer_type[order(cancer_type)] ## Order alphabetically

par(mfrow = c(6,4), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(0,0,0,0))
for(n in cancer_type[1:23]) {
with(umap_cancertype %>% 
       filter(project_code == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F, axes = FALSE,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
legend("bottom", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2, adj = 0.25)
}
```

**Figure S1. Distribution of cancer type on UMAP clusters, Related to Figure 2.** Each panel shows the identity of 
individual tumors by one of 23 cancer types projected onto the cluster UMAP identified 
in **Figure 1B of the main text**. The colors indicate the tumor's cluster identity.

\newpage

```{r supplementary-figure-s2-data-generation, fig.width = 9, fig.height = 10}
load("../data/ditto.colours.rda")

## Determining which cancer type has subtype information
cancer_w_subtypes <- list()
for(n in cancer_type) {
  spec_cancer <- umap_cancertype %>% 
    mutate(J_color = case_when(project_code == n ~ Subtype_Selected,
                               TRUE ~ "Other")) %>%
    filter(J_color != "Other")
  if(nrow(spec_cancer) == 0) {next
  }
  cancer_w_subtypes[[n]] <- n
}

cancer_w_subtypes <- names(cancer_w_subtypes)

par(mfrow = c(6,4), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(0,0,0,0))

for(n in cancer_w_subtypes[1:19]) {
  spec_cancer <- umap_cancertype %>% 
    mutate(J_color = case_when(project_code == n ~ Subtype_Selected,
                               TRUE ~ "Other")) %>%
    filter(J_color != "Other")
  if(nrow(spec_cancer) == 0) {next
    }
  subtypes <- unique(spec_cancer$J_color)
  J_colors <- c(ditto_colours[c(2:(length(subtypes) + 1))])#, 'grey')
  names(J_colors) <- subtypes

  with(spec_cancer,
     scatter3D(plot_y, -plot_z, plot_x, 
               bg = J_colors[J_color], pch = 21, cex = 0.8, lwd = 0.2,
               theta = 0, phi = 65, scale = F, axes = FALSE,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
               xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
               colvar = NULL))
  legend("bottom", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2, adj = 0.25)
  legend("top", inset = c(0), legend=names(J_colors[1:2]), col=J_colors[1:2], pch = 20, bty = "n", horiz = T, cex = 1)
  legend("top", inset = c(0.08), legend=names(J_colors[3:4]), col=J_colors[3:4], pch = 20, bty = "n", horiz = T, cex = 1)
  legend("top", inset = c(0.16), legend=names(J_colors[5:6]), col=J_colors[5:6], pch = 20, bty = "n", horiz = T, cex = 1)
}
```

**Figure S2. Histological subtypes, Related to Figure 2.** Each panel shows the histological subtypes of one of the 23 
cancers projected onto the tumors of the UMAP from **Figure 1B of the main text**. The colors indicate the tumor's histological subtype identity.

\newpage

```{r supplementary-figure-s3-data-generation, eval = FALSE}
load("../data/gdc_tumor_coding_complete.Rdata")
load("../data/gene_data.Rda")

## Booleanize mutation count for each tumor sample
missense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$missense) 
splice <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$splice)
frameshdel <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshdel)
frameshins <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshins)
utr3 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr3)
utr5 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr5)
nonstop <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonstop)
nonsense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonsense)

gdc_mut_count_patient <- missense + splice + frameshdel + frameshins + utr3 + utr5 + nonstop + nonsense ## add up all the mutation count
rm(missense, splice, frameshdel, frameshins, utr3, utr5, nonstop, nonsense)
gc()
gdc_mut_count_patient[gdc_mut_count_patient > 0] <- 1 ## booleanize data, anything greater than 1 is equal to 1

## MMR genes
genes <- c("MSH2", "MSH6", "MLH1", "MLH3", "PMS1", "PMS2", "BRIP1", "RAD51", "CHEK2", "APC",
           "TET2", "TET3", "TERT", "TEP1", "DKC1")

selected_mut <- gene_data %>% dplyr::select(external_gene_name, ensembl_gene_id) %>% filter(external_gene_name %in% genes) ## getting ensembl_id for the mmr genes for filtering gdc data

selected_mut <- gdc_mut_count_patient %>% filter(rownames(.) %in% selected_mut$ensembl_gene_id) %>% ## filter by ensembl ids
  rownames_to_column(var = "ensembl_gene_id") %>% left_join(., selected_mut, by = "ensembl_gene_id") %>% ## replace ensembl ids with gene name 
  column_to_rownames(var = "external_gene_name") %>% dplyr::select(-ensembl_gene_id) %>%
  t() %>% as.data.frame() %>% rownames_to_column(var = "sample_id")  ## prep for left_join

selected_mut <- umap_cancertype %>% dplyr::select(sample_id, plot_x, plot_y, plot_z, clust_knn) %>% ## select relevant columns
  left_join(., selected_mut, by = "sample_id") %>%
  melt(data = ., id.vars = c("sample_id", "plot_x", "plot_y", "plot_z", "clust_knn")) %>%
  filter(value != 0)

save(selected_mut, file = "../data/selected_mut.Rda")
```

```{r supplementary-figure-s3, fig.width = 9, fig.height = 10}
load("../data/selected_mut.Rda")

genes <- c("MSH2", "MSH6", "MLH1", "MLH3", "PMS1", "PMS2", "BRIP1", "RAD51", "CHEK2", "APC", "TET2", "TET3", "TERT", "TEP1", "DKC1")

par(mfrow = c(6,4), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(0,0,0,0))
for(n in genes[1:15]) {
with(selected_mut %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F, axes = FALSE,
                                xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}
```

**Figure S3. Distribution of mismatch repair genes on UMAP clusters, Related to Figure 2.** Each panel shows one of 15 
genes involved in genome stability projected on the UMAP from **Figure 1B of the main text**. The colors indicate the tumor's cluster identity.

\newpage

```{r, supplemental-s4-data-generation, eval = FALSE}
load("../data/gdc_tumor_coding_complete.Rdata")
load("../data/gene_data.Rda")
umap_cancertype <- fread("../data/umap_3d_coors.tsv") ## load data

## Booleanize mutation count for each tumor sample
missense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$missense) 
splice <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$splice)
frameshdel <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshdel)
frameshins <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$frameshins)
utr3 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr3)
utr5 <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$utr5)
nonstop <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonstop)
nonsense <- as.data.frame(SummarizedExperiment::assays(gdc_tumor_coding_complete)$nonsense)

gdc_mut_count_patient <- missense + splice + frameshdel + frameshins + utr3 + utr5 + nonstop + nonsense ## add up all the mutation count
rm(missense, splice, frameshdel, frameshins, utr3, utr5, nonstop, nonsense)
gc()
gdc_mut_count_patient[gdc_mut_count_patient > 0] <- 1 ## booleanize data, anything greater than 1 is equal to 1

## class 7 and 8 distinguishing genes
genes <- c("AKT2","BCL2","CTNNB1","ERBB2","KRAS","PIK3CA","CDKN2C","PTEN","RB1","SMARCA4","TP53","VHL")

selected_mut <- gene_data %>% dplyr::select(external_gene_name, ensembl_gene_id) %>% filter(external_gene_name %in% genes) ## getting ensembl_id for the genes for filtering gdc data

selected_mut <- gdc_mut_count_patient %>% filter(rownames(.) %in% selected_mut$ensembl_gene_id) %>% ## filter by ensembl ids
  rownames_to_column(var = "ensembl_gene_id") %>% left_join(., selected_mut, by = "ensembl_gene_id") %>% ## replace ensembl ids with gene name 
  column_to_rownames(var = "external_gene_name") %>% dplyr::select(-ensembl_gene_id) %>%
  t() %>% as.data.frame() %>% rownames_to_column(var = "sample_id")  ## prep for left_join

s5_data <- umap_cancertype %>% dplyr::select(sample_id, plot_x, plot_y, plot_z, clust_knn) %>% ## select relevant columns
  left_join(., selected_mut, by = "sample_id") %>%
  melt(data = ., id.vars = c("sample_id", "plot_x", "plot_y", "plot_z", "clust_knn")) %>%
  filter(value != 0)

save(s5_data, file = "../data/s5_data.Rda")
```

```{r supplemental-s4-plot, fig.width=9}
load("../data/s5_data.Rda")
genes <- c("AKT2","BCL2","CTNNB1","ERBB2","KRAS","PIK3CA","CDKN2C","PTEN","RB1","SMARCA4","TP53","VHL")

par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in genes[1:6]) {
with(s5_data %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}

par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in genes[7:12]) {
with(s5_data %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}

```

**Figure S4. Distribution of oncogenes/tumor suppressors on UMAP clusters, Related to Figure 2.** Each panel shows shows one of the twelve oncogenes/tumor suppressors projected onto the tumors of the UMAP from **Figure 1B of the main text**. The colors indicate the tumor's cluster identity.

\newpage

```{r supplemental-s5-plot-data, eval = FALSE}
## Calculate error distribution

## break error into bins for entropy calculation
## Set number of bins - 1024 bins
binsize <- 1024 - 2 ## -2 bins for the less than -3650 and greater than 3650 values
evaluebreak <- c(-1e10,seq(from = -3650.00, to = 3650.00, by = 3650*2/binsize), 1e10)
elabel <- list()
counter <- 2
for(i in evaluebreak[c(-1,-(length(evaluebreak)-1))]) {
  if(i == -1e10) {
    elabel[[counter]] <- paste0(i, "+")
  } else if(i == 1e10) {
    elabel[[counter]] <- paste0(i, "+")
  } else {
    elabel[[counter]] <- paste0(i, "-", round(evaluebreak[[counter+1]], digits = 2))
  }
  counter <- counter + 1
}
elabel <- as.character(elabel)
elabel[1] <- "-3650+"
elabel[1024] <- "3650+"

load("../data/deadpat_all.rda")
deadpat <- deadpat %>% rownames_to_column(var = "stanid")
## convert negative tdeath to 0
deadpat[which(deadpat$tevent < 0), "tevent"] <- 0
model <- c("knn", "icluster", "coca", "tis")
tdeath.error <- list() ## store error values
entropy <- list() 
for(n in model) {
  ## Load correct model
  if(n == "knn") {
    load("./survival/survmodel_effcl_knn_070121.stansave")
  } else if(n == "icluster") {
    load("./survival/survmodel_effcl_icluster_070121.stansave")
  } else if (n == "coca") {
    load("./survival/survmodel_effcl_coca_070121.stansave")
  } else if (n == "tis") {
    load("./survival/survmodel_effcl_tis_070121.stansave")
  } else {
    print("Model does not exist")
  }
  ## Extract tdeath for "adeath" indivduals
  tdeath <- samps %>% as.data.frame(.) %>% .[, grep(pattern = "tdeath", x = names(.))]
  names(tdeath) <- gsub("tdeath\\[(.*)\\]","\\1", names(tdeath)) ## update colname with stanid
  tdeath <- tdeath %>% gather(data = ., key = "stanid", value = "predicted") %>%
    left_join(x = ., y = deadpat[,c("stanid", "tevent", "eventtype")], by = "stanid") %>%
    dplyr::filter(eventtype == "adeath")
  ## Calculate error for each sample
  tdeath$error <- tdeath$predicted - tdeath$tevent
  tdeath.error[[n]] <- tdeath$error
  ## Split error data into bins and tally up 
  tdeath <- tdeath %>% mutate(ebin = cut(x = error,
                                         breaks = evaluebreak,
                                         right = T,
                                         labels = elabel))
  eprob <- as.data.frame(table(tdeath$ebin))
  eprob <- eprob %>% dplyr::filter(Freq > 0) %>%
    mutate(prob = Freq / nrow(tdeath))
  ## Entropy calculation
  entropy[[n]] <- round(-sum(eprob$prob*log(eprob$prob))/log(2), digit = 3)
}
knn <- data.frame(error = tdeath.error$knn,
                  model = "knn")
icluster <- data.frame(error = tdeath.error$icluster,
                       model = "icluster")
coca <- data.frame(error = tdeath.error$coca,
                   model = "coca")
tis <- data.frame(error = tdeath.error$tis,
                   model = "tis")
tdeath.error <- rbind(knn, icluster, coca, tis); rm(knn, icluster, coca, tis); gc()

## empirical cumulative distribution function
## Model selection
model <- c("knn", "icluster", "coca", "tis")

for(n in model) {
   ## Load correct model
  if(n == "knn") {
    load("./survival/survmodel_effcl_knn_070121.stansave")
  } else if(n == "icluster") {
    load("./survival/survmodel_effcl_icluster_070121.stansave")
  } else if (n == "coca") {
    load("./survival/survmodel_effcl_coca_070121.stansave")
  } else if (n == "tis") {
    load("./survival/survmodel_effcl_tis_070121.stansave")
  }
  else {
    print("Model does not exist")
  }
  dfs <- as.data.frame(samps) %>% 
    select(starts_with("tdeath[")) %>%
    stack(.) %>%
    'colnames<-' (c("tdeath", "stanid")) %>%
    .[sample(x = nrow(.), size = 1e5),]
  assign(x = n, value = dfs)
}

save(entropy, tdeath.error, knn, icluster, coca, tis, file = "../data/s5_plot.rda")
```

```{r supplemental-s5-plot, fig.width=9}
load("../data/s5_plot.rda")
load("../data/deadpat_all.rda")

s5.1 <- ggplot(knn) + stat_ecdf(aes(x=tdeath),color="#00BFC4", size = 1) + 
  stat_ecdf(aes(x=tevent),data=subset(deadpat,eventtype=="adeath"),col="#00BA38", size = 1) +
  stat_ecdf(aes(x=tdeath),data=icluster, color="#7CAE00", size = 1) + 
  stat_ecdf(aes(x=tdeath),data=coca, color="#F8766D", size = 1) + 
  stat_ecdf(aes(x=tdeath),data=tis, color="#F564E3", size = 1) + 
  coord_cartesian(xlim=c(0,365*5)) + 
  labs(x = "Survival Time (years past diagnosis)", y = "Fn(Survival Time)") +
  annotate(geom = "text", x = 250, y = 0.5, label = "Actual Death", color = "#00BA38", size = 3) +
  theme_minimal()

s5.2 <- ggplot(data = tdeath.error, aes(x = error, group = model, fill = model)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  #scale_x_continuous(breaks=seq(-4000, 4000, by = 2000), limits = c(-3650,3650)) +
  labs(x = "Error (Predicted - Actual)", y = "Density")#, 
       #subtitle = paste0("Entropy - knn: ", entropy$knn, " ; icluster: ", entropy$icluster, "\ncoca: ", entropy$coca, " ; tis: ", entropy$tis))

layout = "
AABB
AABB
"

s5.1 + s5.2 + plot_annotation(tag_levels = "A") + plot_layout(design = layout)
```

**Figure S5. Performance comparison of different classification model - knn, COCA, iCluster, tissue-of-origin (tis), Related to Figure 5.** A) Empirical cumulative distribution function comparing time of death prediction of all classification models with actual time of death (dark green). B) Error distribution plot of predicted and actual time of death.

\newpage

```{r child='../code/figureSM1.Rmd'}
```

**Figure S6. Method for tumor classification using pathway aberration data, Related to Figure 1.** 
A) MCA analysis was performed to determine the eigenvalue with the most
explanatory power from the categorical data. Average of 100 permutations of the
data was used as a baseline. B) The p-value of each eigenvalue was plotted and
the max eigenvalue with a pvalue < 0.05 was selected. C) The resulting MCA was
projected down to 3 dimensions for the purpose of density based clustering (only
the first 2 dimensions are shown). Each dot represents a tumor D) HDBSCAN was
used to perform hierarchical density based clustering tuned to minimize the
fraction of unassigned tumors. E) HDBSCAN clustered tumors were visualized on
the UMAP projection; tumors are colored by cluster membership (see legend). This
results in some tumors remaining unclassified (grey points). F) kNN was used to
assign membership to the remaining tumors.

\newpage

```{r child='../code/figureSM2.Rmd'}
```

**Figure S7. Comparison between CDC data and ODE model fit for
males and females of European descent, Related to Figure 5.** This comparison does not distinguish on
ethnicity for simplicity in visualization.

\newpage

```{r sm3, fig.height=3, fig.width=6}
sm3.1 <- ggplot(data.frame(x=0:110),aes(x)) + 
  stat_function(fun=function(x) dgamma(x,30.0,29.0/65),geom="area", fill = alpha("black", 0.5)) +
  theme_classic() + 
  scale_x_continuous(limits = c(30, 110), breaks = c(seq(30,110, by =10))) +
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        panel.grid.major = element_line(size = 0.75),
        panel.grid.minor = element_line()) +
  labs(x= expression(paste("Prior for A"["0"])), y = "Probability Density")

sm3.2 <- ggplot(data.frame(x=0:3),aes(x)) + 
  stat_function(fun=function(x) dgamma(x,50.0,49.0/1.0),geom="area", fill = alpha("black", 0.5)) +
  theme_classic() + 
  scale_x_continuous(limits = c(0.5, 1.6), breaks = c(seq(0.6, 1.5, by = 0.2))) +
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        panel.grid.major = element_line(size = 0.75),
        panel.grid.minor = element_line()) +
  labs(x= expression(paste("Prior for A"["cl"])), y = "Probability Density")

layout = "
AABB
AABB
"

sm3.1 + sm3.2 + plot_annotation(tag_levels = "A") + plot_layout(design = layout)
```

**Figure S8. Priors for the tissue and class coefficients, Related to Figure 5 and 6.** A) The tissue
coefficient is chosen to admit ages between 40 and 100 years old as a baseline
age. B) Within a given tissue, the class parameter multiplies the tissue value and
is given a peak density at 1, and a credible range from near 0.6 to 1.4,
reflecting our expectation of a moderate multiplicative perturbation of
variation across the classes.