### 2021-07-20 code/survival/effage_stanmodel.Rmd

```{r load-libraries}
library(sqldf)
library(tidyverse)
library(rstan)
library(readxl)
library(ggkm)
library(gridExtra)
library(PerformanceAnalytics)
library(patchwork)
```

```{r add-other-classes}
setwd("D:/repo/TTmanu/data/survival/")

cancerdat = read_csv("clinical_plus_cluster.csv");
treatdays = read_csv("tcga_clinical_data.csv")

#####################################
## Other TCGA paper classification ##
#####################################
## Hoadley et al. 2018 PMID: 29625048
icluster <- read_excel(path = "NIHMS957746-supplement-6.xlsx", skip=1)
icluster <- icluster %>% 'colnames<-' (c("sample_id", "icluster"))
## Hoadley et al. 2015 PMID: 25109877
coca <- read_excel(path = "mmc2.xlsx", skip=1)
coca <- coca %>% dplyr::select(Sample, COCA) %>% 'colnames<-' (c("sample_id", "coca"))

## Append additional classification model to cancer data
cancerdat <- cancerdat %>% mutate(sample_id = substring(text = sample_id, first = 1, last = 12)) %>% 
  left_join(x = ., y = icluster, by = "sample_id") %>% 
  left_join(x = ., y = coca, by = "sample_id")

## Assign unclassified tumor samples in icluster and coca to the next cluster assignment number
cancerdat[which(is.na(cancerdat$icluster)), "icluster"] <- max(cancerdat$icluster, na.rm = TRUE) + 1
cancerdat[which(is.na(cancerdat$coca)), "coca"] <- max(cancerdat$coca, na.rm = TRUE) + 1
```

```{r prep-for-stan}
setwd("D:/repo/TTmanu/data/survival/")
#####################
## Processing Data ##
#####################
deadpat <- sqldf("select patient_id,
                 age_at_initial_pathologic_diagnosis as age, 
                 gender,
                 race, 
                 death_days_to,type,
                 clust_knn as clust, 
                 clust as origclust, 
                 icluster, 
                 coca,
                 `ajcc_pathologic_tumor_stage.y` as stage, 
                 last_contact_days_to as lastcont from cancerdat where (death_days_to is not null or lastcont is not null) and age is not null and age > 9")

## Clean up data
deadpat$patient_id <- as.factor(deadpat$patient_id)
deadpat$type <- as.factor(deadpat$type)
deadpat$tevent <- deadpat$death_days_to
deadpat[is.na(deadpat$tevent),"tevent"] <- deadpat[is.na(deadpat$tevent),"lastcont"]
deadpat[is.na(deadpat$death_days_to),"eventtype"] <- "bfollow"
deadpat[!is.na(deadpat$death_days_to),"eventtype"] = "adeath"
deadpat$eventtype = as.factor(deadpat$eventtype)
deadpat$gender = as.factor(deadpat$gender)

#############################
## mortality data from CDC ##
#############################
namesmort = c("agerng","pdead","nsurv","ndie","pyrs","pyrsabove","expectedlife")
menmort = read_csv("lifetables/Table02.csv",skip=2)
## drop the final row, it's just a text info statement
menmort = menmort[1:NROW(menmort)-1,] 
wommort = read_csv("lifetables/Table03.csv",skip=2)
wommort = wommort[1:NROW(wommort)-1,]

names(menmort) = namesmort
names(wommort) = namesmort
yrs=seq(0,100,by=1);
menmort$yrnum = yrs
wommort$yrnum = yrs

## Store original cancer data in new object for further analysis
#save(deadpat, file = "D:/repo/TTmanu/data/survival/deadpat_all.rda")
load("../../data/survival/deadpat_all.rda")
standata <- deadpat

## convert negative tdeath to 0
standata[which(standata$tevent < 0), "tevent"] <- 0
## There are currently 3 types of classification model in this analysis:
## knn (our classification), icluster, and coca

model <- "tis"

if(model == "knn") {
  standata$clust <- as.factor(standata$clust)
  colnames(standata)[which(names(standata) == "clust")] <- "modelofinterest"
} else if(model == "icluster") {
  standata$icluster <- as.factor(standata$icluster)
  colnames(standata)[which(names(standata) == "icluster")] <- "modelofinterest"
} else if(model == "coca") {
  standata$coca <- as.factor(standata$coca)
  colnames(standata)[which(names(standata) == "coca")] <- "modelofinterest"
} else if(model == "tis") {
  standata$modelofinterest <- standata$type ## Convert tissue type to modelofinterest
  standata$type <- "none" ## No tissue category
  standata$type <- as.factor(standata$type) ## Factor the no tissue category
} else {
  print("Incorrect model selection")
}

if(model %in% c("knn", "icluster", "coca")) {
  standata <- list(Np=NROW(standata), 
                Nt=NROW(levels(standata$type)), 
                Nc=NROW(levels(standata$modelofinterest)),
                ##patient=as.numeric(deadpat$patient_id),
                age=standata$age,
                tissue=as.numeric(standata$type),
                menccdf=menmort$nsurv/100000,
                womccdf=wommort$nsurv/100000,
                gender=as.numeric(standata$gender),
                tclass=as.numeric(standata$modelofinterest),
                tevent = standata$tevent,
                eventtype=as.numeric(standata$eventtype))
}

if(model == "tis") {
  standata <- list(Np=NROW(standata), 
                Nt=NROW(levels(standata$modelofinterest)), 
                age=standata$age,
                tissue=as.numeric(standata$modelofinterest),
                menccdf=menmort$nsurv/100000,
                womccdf=wommort$nsurv/100000,
                gender=as.numeric(standata$gender),
                tclass=as.numeric(standata$modelofinterest),
                tevent = standata$tevent,
                eventtype=as.numeric(standata$eventtype))
}

### Fit the survival model with one rate per tissue*class
rm(smodel, start, start2, start3); gc()
if(model %in% c("knn", "icluster", "coca") == TRUE) {
  smodel = stan_model(file="../../code/survival/survmodel.stan")
} else if(model %in% c("tis", "icluster.kcl", "coca.kcl") == TRUE) {
  smodel = stan_model(file="../../code/survival/survmodel_tisonly.stan")
} else {
  print("Incorrect model selection")
}

## optimization for initialization doesn't always work... 
start = optimizing(smodel,data=standata);
start2 = optimizing(smodel,data=standata);
start3 = optimizing(smodel,data=standata);

if(model == "knn") {
  tryCatch(load("survmodel_effcl_knn_070121.stansave"),error=function(e){
    samps = sampling(smodel,data=standata,chains=4,warmup=1000,iter=2500,cores=4,control = list(max_treedepth=9)) #,init=list(start,start2,start3));
    save("samps",file="survmodel_effcl_knn_070121.stansave")
  })
} else if(model == "tis") {
  tryCatch(load("survmodel_effcl_tis_070121.stansave"),error=function(e){
    samps = sampling(smodel,data=standata,chains=4,warmup=1000,iter=2500,cores=4,control=list(max_treedepth=9)) #,init=list(start,start2,start3));
    save("samps",file="survmodel_effcl_tis_070121.stansave")
  })
} else if(model == "icluster") {
  tryCatch(load("survmodel_effcl_icluster_070121.stansave"),error=function(e){
    samps = sampling(smodel,data=standata,chains=4,warmup=1000,iter=2500,cores=4,control=list(max_treedepth=9)) #,init=list(start,start2,start3));
    save("samps",file="survmodel_effcl_icluster_070121.stansave")
  })
} else if(model == "coca") {
  tryCatch(load("survmodel_effcl_coca_070121.stansave"),error=function(e){
    samps = sampling(smodel,data=standata,chains=4,warmup=1000,iter=2500,cores=4,control=list(max_treedepth=9)) #,init=list(start,start2,start3));
    save("samps",file="survmodel_effcl_coca_070121.stansave")
  })
} else {
  print("Incorrect model selection")
}

```